---
title: "TILFAR"
subtitle: Things I Learned From Advanced R
date: 2020-09-22
author: Tan Ho
output: 
  html_document:
    theme: darkly
    css: tan-styles.css
---

I set out reading this book hoping to understand the metaprogramming chapter, and ended up learning a bunch of other stuff along the way!
Here's a non-exhaustive list of things I learned from each of the sections of Advanced R:

### I - Foundations: 

- I notice that I use the switch function a ton now, where previously I used if-else nests.
- I use condition-handling to pass errors back to the user of my Shiny apps now.
- I now know what an Environment is and where and why R would look for something. (This helps both in function-writing and debugging!)
- I write a heck of a lot of functions (this ties into II: Functional Programming, of course)

### II - Functional Programming

- I map the crap out of everything now (if I didn't before, it's way worse now!), and have learned to use functions in a ton of different ways.
- I now understand what a function factory is, and what to do if a function somehow returns me a function instead (as opposed to panicking about it!)
- I now know about a series of modifiers to help adjust function behaviours (memoise, quietly, safely, possibly) - although I use tryCatch as much as I do these.

### III - OOP

- I wrote a package that uses S3! It's cool to minimize the user interface for multiple platforms and the mental load accompanying. 
- I still lean to using the explicit "switch" function a lot instead of writing out a useMethod call. May be a habit/pattern worth breaking in the future.
- I'm now comfortable USING an R6 package - I don't know that I'd reach for an R6 system now but the whole "function$method" thing doesn't scare me anymore
- S4? wot that

### IV Metaprogramming

- I now know how to pass expressions to tidyverse stuff!
- I now know where to look in the Metaprogramming and how to intepret the text.

### V Techniques

- Debugging: I now know how to use browser, and love it! Also debug()!
- Performance measure/improve: still don't do a ton of this, even though I'm a shiny dev. I feel like I should care more but I mostly work with smaller data (sub 1,000,000 rows) right now. 
- I ... still don't know C++ or Rcpp. `r emo::ji('shrug')`

## Toy Code
```{r groupvars, eval = FALSE, echo = FALSE}
library(dplyr)
library(rlang)

grouping_variables <- switch(
  params$group_type,
  "S" = exprs(CraftCode, starts_with("Product"), ModelCode, Elevation),
  "B" = exprs(starts_with("Rate")),
  "P" = exprs(starts_with("Part"))
)
  
null_bids <- bids %>%
  group_by(SupplierID, ProjectID, !!!grouping_variables) %>%
  slice_max(LastModDate) %>%
  ungroup() %>%
  select(!!!grouping_variables, ProjectID, SupplierAID, SupplierName, UnitPrice) %>%
  left_join(lowest_bids, by = as.character(grouping_variables))

```

Metaprogramming - how to pass expressions into select statements. Also example of switch. 
```{r echo = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(flair)})

flair::decorate_chunk("groupvars",eval = FALSE) %>% flair("grouping_variables")
```

