---
title: "Chapter 12: Base Types"
subtitle: "Advanced R Book Group, Cohort 3"
author: "Megan Stodel (@MeganStodel, www.meganstodel.com)"
date: "10 November, 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---
class: inverse, center, middle

# But first...

---
# Object-oriented programming (OOP)

In R, functional programming is much more important than OOP. This is because problems are decomposed into simple *functions* rather than simple *objects*. 

OOP can be more challenging in R:

* There are multiple OOP systems (3 are explored in this book).

* The different systems are not objectively set in relative importance so different communities use different systems.

* S3 and S4 use generic function OOP rather than encapsualted OOP, which is more common in other langauges, so it is hard to transfer existing OOP skills to R. 
---
# The three systems

* *S3:* Used throughout base R. Functions can return rich results with user-friendly display and programmer-friendly internals. R's first OOP system and relies on common conventions rather than ironclad guarantees. 

* *R6:* A way to escape copy-on-modify semantics; important if you model objects that exist independently of R - for e.g. data that comes from a web API. An implementation of encapsulated OOP. 

* *S4:* A rigorous system that requires careful design consideration, so suited for large systems that might have multiple contributions. Offers more guarantees and greater encapsulation than S3, but is more work upfront. 
---
# OOP glossary

* *polymorphism* - a developer can consider a function's interface separately from its implementation, so the same function form can be used for different types of output. 

* *encapsulation* - the user doesn't need to worry about an object's details because they are 'encapsulated' behind a standard interface.

* *class* - the type of an object.

* *method* - a implementation for a specific class.

* *fields* - defined by the class, this is the data possessed by all instances of that class.

* *method dispatch* - the process of finding the correct method for a given class. 
---
# But what does that mean?
## An example

Polymorphism allows `summary()` to produce different outputs for numeric and factor variables.

```{r}
diamonds <- ggplot2::diamonds
summary(diamonds$carat)
summary(diamonds$cut)
```

This is more flexible than an `if-else` statement series because any developer can extend the interface.
---
# The two paradigms

*Encapsulated* OOP has methods that belong to objects or classes; the object encapsulates both data (with fields) and behaviour (with methods). 

*Functional* OOP has methods that belong to generic functions. From the outside it looks like a regular function call, and internally the components are also functions. 

---
class: inverse, center, middle

# Base Types

---
# What is an object?

There are two uses of the word "object". 

So far we have talked very generally...

> "Everything that exists in R is an object." - John Chambers

But not everything is object-oriented. This is to do with the organic evolution of S, because base objects come from S when they were developed before it was considered to need an OOP system. So we distinguish here between _Base objects_ and _OO objects_. 
---
# What kind of object is it?

Use `is.object()` (TRUE only if OO object) or `sloop::otype()` (returns object type). The technical difference is whether the object has a 'class' attribute, so you can also test with `attr()`.

```{r}
num_range <- 1:10
is.object(num_range)
sloop::otype(num_range)
attr(num_range, 'class')
```
---
# What kind of object is it?

```{r}
is.object(diamonds)
sloop::otype(diamonds)
attr(diamonds, 'class')
```

---
# What kind of object is it?

`sloop::s3_class()` is aso useful; it returns the implicit class used to pick the methods by S3 and S4 systems (which gets missed by `class()`). 

```{r}
x <- matrix(1:4, nrow = 2)
class(x)
sloop::s3_class(x)
```

---
# So what is a base type?

Every object has a base type! 

```{r}
typeof(1:10)
typeof(diamonds)
```

If a function behaves differently for different base types, they are primarily written in C code using switch statements. This means only R-core can create new types and it is a real mission. Therefore new base types are not added often. 

(Remember: with OO objects, any developer can specify a new behaviour for a new class.)

---
# Yes, but what are some types?

There are 25 different base types. They can be grouped into...

*VECTORS*, including types `NULL`, `logical`, `integer`, `double`, `complex`, `character`, `list` and `raw`. 

*FUNCTIONS*, including `closure`, `special` and `builtin`. 

*ENVIRONMENTS*

*S4*, which is for S4 classes that don't inherit from an existing base type. 

*LANGUAGE COMPONENTS*, including `symbol`, `language` and `pairlist`. 

There are also some esoteric and rarely used types that mainly exist because they are important for C code. 
---
# Is there a numeric type? (1)

There are three different meanings of numeric type so be careful. 

Sometimes used to mean double type - for example `as.numeric()` is the same as `as.double()`.

---
# Is there a numeric type? (2)

In S3 and S4, a shorthand for either integer or double type, used when picking methods. 

```{r}
sloop::s3_class(1)
sloop::s3_class(1L)
```
---
# Is there a numeric type? (3)

It can also mean objects that behave like numbers. Factors have the type `integer` but don't behave like numbers (you wouldn't care about the mean for e.g.).

```{r}
typeof(factor("x"))
is.numeric(factor("x"))
```

Throughout we are using the second meaning - either integer or double. 

---
class: inverse, center, middle

# General discussion
---
# Some possible topics

* Why use OOP?

* Why are there different OOP systems in R? 

* What are some ways of finding out if an object is an OO object?
