---
title: "Advanced R"
subtitle: "Metaprogramming Review"
author: "Jon Harmon"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---

```{r libraries, include = FALSE}
library(xaringanthemer)
library(emo)
library(lobstr)
library(magrittr)
```

```{r xaringan-themer, include = FALSE}
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

# Learning Objectives: Ch. 17

- 17.2 Understand that code in R is data.
- 17.3 Recognize an **abstract syntax tree** as a representation of a function call.
- 17.4 Understand that R code can write code.
- 17.5 Remember that `base::eval` executes expressions in environments.
- 17.6 Understand that environments can customize evaluation.
- 17.7 Understand that `rlang::eval_tidy` lets you use data frames like environments.
- 17.8 Recall that **quosures** bind together an expression with an environment.

--
## `r emo::ji("heavy_check_mark")`

---

# Learning Objectives: Ch. 18

- 18.2 Translate between ASTs and R code.
- 18.3 Define the expression subtypes (other than pairlist).
- 18.4 Describe the grammar used by R's parser to translate code to expressions.
- 18.5 Write recursive functions to walk the AST.
- 18.6 Define the expression subtypes.

---

# Learning Objectives: Ch. 18

- 18.2 Translate between ASTs and R code.

```{r ast}
lobstr::ast({`+` = `-`; 1} + (100 + 1) + 1)
{`+` = `-`; 1} + (100 + 1) + 1
```

---

# Learning Objectives: Ch. 18

- 18.3 Define the expression subtypes (other than pairlist).
 - **Constants:** NULL & length-1 atomic vectors.
 - **Symbols:** Names of things, such as `x`, `mtcars`, or `mean`.
 - **Calls:** A captured function... call. Special type of list.
 - Symbols and calls are captured with `rlang::expr`.

---

# Learning Objectives: Ch. 18

- 18.4 Describe the grammar used by R's parser to translate code to expressions.
 - **Precedence:** Like math, use () when confusing.
 - **Associativity:** Pretend parentheses are around the left set, except for ^ and <-.
 - You can translate **(parse)** strings to code with rlang & base but you shouldn't.

---

# Learning Objectives: Ch. 18

- 18.5 Write recursive functions to walk the AST.
 - Copy `expr_type`, `switch_expr`, `flat_map_chr` from book.
 - Fill in a `switch_expr` skeleton depending on usecase.
 
 ```{r walk-ast}
 lobstr:::ast_tree
 ```

---

# Learning Objectives: Ch. 18

- 18.6 Define the (remaining) expression subtypes.
 - **Pairlists:** `mean %>% rlang::fn_fmls() %>% class()`
 - **Missing arguments:** Use `rlang::missing_arg()` and `rlang::maybe_missing` if you need to explicitly pass missing args.
 - **Expression vectors:** These exist but you probably don't need to care.

---

# Learning Objectives: Ch. 19

- 19.2 Define quasiquotation.
- 19.3 Quote expressions in rlang and base R.
- 19.4 Selectively unquote parts of an expression during quotation.
- 19.5 Describe the types of non-quoting used in base R.
- 19.6 Mix `...` and lists to supply arguments to functions.
- 19.7 Use rlang's quasiquotation functions to generate code.
- 19.8 Describe the history of quasiquotation.

---

# Learning Objectives: Ch. 19

- 19.2 Define quasiquotation.
 - Quotation with selective unquoting.

---

# Learning Objectives: Ch. 19

- 19.3 Quote expressions in rlang and base R.

**rlang quasiquoting functions**

|      | Command line | In functions   |
|------|--------------|----------------|
| One  | `expr()`     | `enexpr()`     |
| Many | `exprs()`    | `enexprs()`    |

**base R quoting functions**

|      | Command line | In functions                 |
|------|--------------|------------------------------|
| One  | `quote()`    | `substitute()`               |
| Many | `alist()`    | `as.list(substitute(...()))` |

---

# Learning Objectives: Ch. 19

- 19.4 Selectively unquote parts of an expression during quotation.
 - `!!` to unquote one thing
 - `!!!` to unquote many (often dots)

---

# Learning Objectives: Ch. 19

- 19.5 Describe the types of non-quoting used in base R.
 - `bquote` exists.
 - **Quoting/non-quoting function pairs:** `$` & `[`
 - **Quoting/non-quoting argument pairs:** `rm` quotes `...`, doesn't quote `list`
 - **Argument controls quoting:** `library(pkg)` vs `library(pkg, character.only = TRUE)`.
 - **Quoting if evaluation fails:** `help` tries to quote then unquotes if that fails.
 - **plotting functions:** Follow the "standard non-standard evaluation rules."

---

# Learning Objectives: Ch. 19

- 19.6 Mix `...` and lists to supply arguments to functions.
 - `rlang::exec` vs `base::do.call`

---

# Learning Objectives: Ch. 19

- 19.7 Use rlang's quasiquotation functions to generate code.

```{r summands}
intercept <- 10
coefs <- c(x1 = 5, x2 = -4)

summands <- c(intercept,
  purrr::map2(
    .x = rlang::syms(names(coefs)),
    .y = coefs,
    .f = ~ rlang::expr(!!.x * !!.y)
  )
)

eq <- purrr::reduce(.x = summands, .f = ~rlang::expr(!!.x + !!.y))
eq
rlang::eval_tidy(eq, list(x1 = 1:10, x2 = 10:1))
```

---

# Learning Objectives: Ch. 20

- 20.2 Use `base::eval` to process expressions.
- 20.3 Use `rlang::eval_tidy` to evaluate quosures.
- 20.4 Use data masks to provide variable definitions for tidy evaluation.
- 20.5 Avoid errors when working with functions that use tidy evaluation.
- 20.6 Remember the caveats of working with evaluation in base R.

---

# Learning Objectives: Ch. 21

- 21.2 Generate functions to translate another language into R functions.
- 21.3 Walk the AST to translate R expressions to another language.
