---
title: "Chapter 11 - Function Operators"
output: 
    learnr::tutorial:
        css: css/styles.css
        toc: true
        toc_depth: 2
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(purrr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

# 11 Function operators {-}
## Definition 

Last chapter we learned about __function factories__ which are functions
that creates functions. 

__Function operators__ are similar to __function factories__ in that they return 
functions as output, but a __function operator__ takes one or more functions 
as the input as well. 

## First Example 
Let's say we wanted to make a function that would square the results of a vector 
and also tell us which element in the vector it was working on with a `purr::map`. 

This is easy enough to write.

```{r}
square <- function(x) x ^ 2
talky <- function(x){
  print(glue::glue("Processing value: {x}"))
  
  res <- square(x)
  print(glue::glue("Returning results: {res}"))
  
  return(res)
}
s <- c(3, 2, 1)

purrr::map(s, talky)

```

But what if now I want a function that's going to speak the results for a cube?

Rather than re-writing `talky` I can write a __function operator__ which 
will take in any function as an input

```{r}
chatty <- function(f) {
  force(f)
  
  talkitive_function <- function(x, ...) {
    res <- f(x, ...)
    print(glue::glue("Processing value: {x}"))
    print(glue::glue("Returning results: {res}"))

    return(res)
  }
  
  return(talkitive_function)
}
```

```{r}
square <- function(x) x ^ 2
cube <- function(x) x ^ 3
screaming_strings <- function(x, n = 5) paste(rep(x,n),collapse = "")
s <- c(3, 2, 1)
v <- c("a","o","e")

purrr::map_dbl(s, chatty(square))
purrr::map_dbl(s, chatty(cube))
purrr::map_chr(v, chatty(screaming_strings))
purrr::map_chr(v, chatty(screaming_strings), n = 30)


```


## Pre-built function operators

Chapter 9 tried to convince us to stop using `for` loops for everything. 

But, one reason you might use a for loop, is that at least if one iteration fails, 
you get to keep all the results up to that point

Let's try to sum a list of vectors. 

```{r}
my_broken_list <- list(
  c(1, 2, 3),
  c(2, 2, 2),
  c("oops","I'm","not a double"),
  c(7, -2, 1))
```


```{r,error=TRUE}
#make something the length of my expectation
out <- rep(NA_real_, length(my_broken_list))
for (i in seq_along(my_broken_list)) {
  out[[i]] <- as.double(try(sum(my_broken_list[[i]])))
}
out
```

However if we tried to do the same thing using a __functional__

```{r,error=TRUE}
map_out <- map_dbl(my_broken_list, sum)
map_out
```

Fortunately, `purr` comes with a __function operator__ that allows us to get 
around this error. It's called `safely`

```{r,error=TRUE}

safe_map_out <- map(my_broken_list,purrr::safely(sum) )
safe_map_out %>% str()
```
## A brief aside about `transpose`

The book describes using `transpose` to make the `purrr::safely` output
more friendly. If we learned `transpose` before, I forgot. So a brief description 
of `purrr::transpose`

```{r showmap,exercise = T}
annoying_list <- rerun(5, x = runif(1), y = runif(5))
annoying_list %>% str()
```

`purrr::transpose` will turn this list-of-lists into a list of pairs

```{r}
annoying_list <- rerun(5, x = runif(1), y = runif(5))
annoying_list %>% transpose() %>% str()
```

## How to map `safely`

Back to trying to sum our broken_list without using for loops
```{r,error=TRUE}

safe_map_out_transposed <- transpose(map(my_broken_list,purrr::safely(sum) ))
safe_map_out_transposed %>% str()
```

Now we can easily ask which of our elements lead to problems

```{r surprise,exercise = T}
ok <- map_lgl(safe_map_out_transposed$error, is.null)
my_broken_list[!ok]
```

## A better way to try linear models!

Have a bunch of datasets where the models only converge sometimes?
Use safely!

```
fit_model <- function(df) {
  glm(y ~ x1 + x2 * x3, data = df)
}

models <- transpose(map(datasets, safely(fit_model)))
ok <- map_lgl(models$error, is.null)

# which data failed to converge?
datasets[!ok]

# which models were successful?
models[ok]
```

## Other ways to be safe

`possibly()` returns a default value

```{r}
map_dbl(my_broken_list,purrr::possibly(sum,otherwise = 0))
```

`auto_browser()`: automatically executes browser() inside the function when there’s an error.
```{r autoexample,exercise = TRUE}
my_sum <- function(v){
  final_sum = 0
  if(!every(v,is.numeric)){
    rlang::abort("Not everything is a number!")
  }
  for(i in v){
    final_sum = final_sum + i
  }
  return(final_sum)
}

map(s,auto_browse(my_sum)) %>% simplify()
map(v,auto_browse(my_sum)) %>% simplify()
```


## How to `memoise`

Memoization is a method to speed up a function by storing the results of 
slow function calls and returning the cached results. 

Memoized functions can run much faster, but they use more memory.

```{r memo}
library(microbenchmark)
library(ggplot2)

slow_function <- function(x) {
  Sys.sleep(1)
  x * 10 * runif(1)
}
fast_function <- memoise::memoise(slow_function)
fast_function(1)

mbm = microbenchmark(
  slow_function_run_time =  slow_function(1),
  fast_function_run_time_previously_seen = fast_function(1),
  fast_function_run_time_new_arg = fast_function(2),

  times = 10L
)
autoplot(mbm) + 
  scale_y_log10(labels = scales::comma_format()) + 
  ylab("Runtime") + 
  geom_boxplot()
```

## Calculating Fibonacci series with `memoise`
A more likely use of `memoise` might be calculating fibonacci series

```{r}
fib_naive <- function(n) {
  if (n < 2){
    return(1)
  }

  return(fib_naive(n - 2) + fib_naive(n - 1))
}

fib_meme <- memoise::memoise(function(n) {
  if (n < 2){
    return(1)
  }
  fib_meme(n - 2) + fib_meme(n - 1)
})

mbmfib = microbenchmark(
  fib_naive =  fib_naive(23),
  fib_meme = fib_meme(23)
)
autoplot(mbmfib) + 
  scale_y_log10(labels = scales::comma_format()) + 
  ylab("Runtime") 


system.time(fib_meme(24))
system.time(fib_naive(24))

```

## 11.2.3 Exercise 1

Base R provides a function operator in the form of Vectorize(). What does it do? When might you use it?
```
Description
Vectorize creates a function wrapper that vectorizes the action of its argument FUN.

Usage
Vectorize(FUN, vectorize.args = arg.names, SIMPLIFY = TRUE,
          USE.NAMES = TRUE)
Arguments
FUN	
function to apply, found via match.fun.

vectorize.args	
a character vector of arguments which should be vectorized. Defaults to all arguments of FUN.

SIMPLIFY	
logical or character string; attempt to reduce the result to a vector, matrix or higher dimensional array; see the simplify argument of sapply.

USE.NAMES	
logical; use names if the first ... argument has names, or if it is a character vector, use that character vector as the names.

Details
The arguments named in the vectorize.args argument to Vectorize are the arguments passed in the ... list to mapply. Only those that are actually passed will be vectorized; default values will not. See the examples.

Vectorize cannot be used with primitive functions as they do not have a value for formals.

It also cannot be used with functions that have arguments named FUN, vectorize.args, SIMPLIFY or USE.NAMES, as they will interfere with the Vectorize arguments. See the combn example below for a workaround.

Value
A function with the same arguments as FUN, wrapping a call to mapply.
```

## 11.2.3 Exercise 2
Read the source code for `possibly()`. How does it work?
```
possibly
#> function (.f, otherwise, quiet = TRUE) 
#> {
#>     .f <- as_mapper(.f)
#>     force(otherwise)
#>     function(...) {
#>         tryCatch(.f(...), error = function(e) {
#>             if (!quiet) 
#>                 message("Error: ", e$message)
#>             otherwise
#>         }, interrupt = function(e) {
#>             stop("Terminated by user", call. = FALSE)
#>         })
#>     }
#> }
#> <bytecode: 0x17ce470>
#> <environment: namespace:purrr>
```


## 11.2.3 Exercise 3
Read the source code for `safely()`. How does it work?
```
safely
#> function (.f, otherwise = NULL, quiet = TRUE) 
#> {
#>     .f <- as_mapper(.f)
#>     function(...) capture_error(.f(...), otherwise, quiet)
#> }
#> <bytecode: 0x4b1c0e0>
#> <environment: namespace:purrr>

purrr:::capture_error
#> function (code, otherwise = NULL, quiet = TRUE) 
#> {
#>     tryCatch(list(result = code, error = NULL), error = function(e) {
#>         if (!quiet) 
#>             message("Error: ", e$message)
#>         list(result = otherwise, error = e)
#>     }, interrupt = function(e) {
#>         stop("Terminated by user", call. = FALSE)
#>     })
#> }
#> <bytecode: 0x4b72eb8>
#> <environment: namespace:purrr>
```
## Case Study: Creating our own function operators

Let's say we have a list of urls we want to download and we want to store somewhere

```
urls <- c(
  "adv-r" = "https://adv-r.hadley.nz", 
  "r4ds" = "http://r4ds.had.co.nz/"
  # and many many more
)
path <- paste(tempdir(), names(urls), ".html")
```

We learned in Chapter 9 that we can use `walk2` to download these files
```
walk2(urls, path, download.file, quiet = TRUE)
```

But what if we wanted to download 10,000 urls? 
And this will take a while so we want to add a few extra features like a bit
of a delay to not overload the server and to show a dot every few URLS
so we know our function is still running.

We could do it in a `for` loop

```
for(i in seq_along(urls)) {
  Sys.sleep(0.1)
  if (i %% 10 == 0) cat(".")
  download.file(urls[[i]], paths[[i]])
}
```

But this makes things hard to reuse and everytime we need to change something
we have to go into the loop and edit

Let's try with __function operators__

## Case Study: Our first function operator

```
delay_by <- function(f, amount) {
  force(f)
  force(amount)
  
  function(...) {
    Sys.sleep(amount)
    f(...)
  }
}
```

We can then use the original `walk2` command

```
walk2(urls, path, delay_by(download.file, 0.1), quiet = TRUE)
```

## Case Study: Our second function operator

Then we can create a function to display an occasional dot using the
function factory trick of a stateful function and the super
assigment operator to rebind an index value in the enclosing environment.


```{r dots,exercise = TRUE}
dot_every <- function(f, n) {
  force(f)
  force(n)
  
  i <- 0
  function(...) {
    i <<- i + 1
    if (i %% n == 0) cat(".")
    f(...)
  }
}
walk(1:100, dot_every(runif, 10))
walk(1:100, dot_every(runif, 20))
```

I had questions about this behavior though

```{r dot2,exercise = TRUE}
dot_every_2 <- function(f, n) {
    force(f)
    force(n)
    print(glue::glue("This is n: {n}"))
    
    i <- 0
    function(...) {
        print(glue::glue("This is i: {i}"))
        i <<- i + 1
        if (i %% n == 0) cat(".")
        f(...)
    }
}
count_test <- dot_every_2(runif,1)
```


## Case Study: Putting it all together

Then we can express our original `walk2` like this

```
walk2(
  urls, path, 
  download.file %>% dot_every(10) %>% delay_by(0.1), 
  quiet = TRUE
)
```

## Case Study: Exercises


1.  Weigh the pros and cons of 
`download.file %>% dot_every(10) %>% delay_by(0.1)` vs 
`download.file %>% delay_by(0.1) %>% dot_every(10)`

2. Should you memoise file.download()? Why/why not?

