---
title: "Advanced R Chapter 20: Evaluation"
subtitle: "Daryn Ramsden"
author: "thisisdaryn at gmail dot com"
date: "last updated: `r Sys.Date()`"
institution: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: libs/switch-themer.css
    chakra: libs/remark.js
    nature:
      highlightLines: true
      countIncrementalSlides: false
    includes:
      after_body: libs/toggle.html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = "")
xaringanExtra::use_tile_view()
xaringanExtra::use_panelset()
xaringanExtra::use_webcam()
xaringanExtra::use_editable()
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         
  mute_unhighlighted_code = TRUE  
)

library(rlang)
```

## Evaluation: What is this chapter about?

  * **Evaluation**:  evaluating quoted expressions in custom environments to achieve specific goals.

  * The fact that we are customising the environments means all the forms of evaluation we go over are **non-standard**
  
  * Particular emphasis is placed on one type of *Non-standard Evaluation (NSE)*, **Tidy Evaluation**
  
  * To do *Tidy Evaluation*, we make use of functions in the `rlang` package  
  


---
## The Basics: `base::eval()`


`eval`: evaluates an input expression in an input environment. 

Arguments to `eval`:

  * *expr*: the expression you want to evaluate

  * *envir*: the environment you want to evaluate it in 
  
  
```{r}
x <- 10
eval(rlang::expr(x))
```
```{r}
y <- 2
eval(expr(x + y), env(x = 1000))
```


---
### Something to observe

The first argument of `eval` is evaluated not quoted. 


```{r}
eval(print(x + 1), env(x = 1000))
```

```{r}
eval(expr(print(x + 1)), env(x = 1000))
```

---
### Application: `local()`

`local()`: a base R function that allows you to carry out a series of steps in new environment.

  * common use case: carrying out a multi-step computation and disposing of intermediate data automatically

```{r}
not_foo <- local({
  x <- 10
  y <- 200
  x + y
})
```


---
### Replicating `local` using `eval`

We can replicate `local`: 

```{r}
local2 <- function(expr) {
  env <- env(caller_env())
  eval(enexpr(expr), env)
}

foo <- local2({
  x <- 10
  y <- 200
  x + y
})

foo
```


---
## Application: replicating `source`

Similarly, we can replicate `source`: 

.panelset[
.panel[.panel-name[Replicating `source`]

```{r eval = FALSE}
source2 <- function(path, env = caller_env()) {
  file <- paste(readLines(path, warn = FALSE), collapse = "\n")
  exprs <- parse_exprs(file)

  res <- NULL
  for (i in seq_along(exprs)) {
    res <- eval(exprs[[i]], env)
  }

  invisible(res)
}
```


]
.panel[.panel-name[More concisely]

```{r eval = FALSE}
source3 <- function(file, env = parent.frame()) {
  lines <- parse(file) # creates an expression vector
  res <- eval(lines, envir = env)
  invisible(res)
}
```


]
]



---
## Quosures: what are they? 

Encapsulation of:
  * expression
  
  * environment
  
Coupling is so important that `rlang` provides a composite structure 
  
---
### Quosures:  How do you make them?

`rlang` provides 3 ways:

--

1. `enquo()` and `enquos()` 
  
  * this is the best way
  
--

2. `rlang::quo()` and `rlang::quos()`

  * these exist to match `expr()` and `exprs()`. 
  * You probably don't need this

--

3. `new_quosure()`

  * useful for learning
  * You probably don't need this either

---
### Examples creating quosures

```{r}
quosure_create <- function(x) enquo(x)
quosure_create(a + b)
```

```{r}
quosures_create <- function(x)enquos(x)
quosures_create(list(x = x ^ 2, y = y ^ 3, z = z ^ 4))
```


---
### Creating quosures using `quo` and `new_quosure`


#### Using `quo`

```{r}
quo(x + y + z)
```

#### Using `new_quosure`

```{r}
new_quosure(expr(x + y), env(x = 1, y = 10))
```


---
### Quosures: Under the hood 


```{r}
super_quosure <- new_quosure(expr(x + y + z))
class(super_quosure) # are subclasses of formulas
```

```{r}
is_call(super_quosure) # are call objects
```


```{r}
attr(super_quosure, ".Environment") #have a .Environment attribute
```


```{r}
get_expr(super_quosure) # an expression can be extracted
get_env(super_quosure) # an environment can be extracted
```


---
## Tidy evaluation: What we really came for

A form of NSE utilizing 3 main features:

  * quasiquotation
    
    * talked about this last week


  * quosures

  
  * data masks
  
    * soon ... 

---
### `eval_tidy()`: the function that does the work

`eval_tidy` takes two arguments: 

  1. a quosure
  
  2. a data mask (data frame): first place to look for variable definitions
  
Example: using `eval_tidy` to find the largest penguin (mass) in `palmerpenguins::penguins`:  
  
```{r}
library(palmerpenguins)

penguin_quosure <- quosure_create(max(body_mass_g, na.rm = TRUE))

# Now use the penguins data frame as data mask
eval_tidy(penguin_quosure, penguins) 
```

---
### Example: replicating `with`

How `with` works:

```{r}
library(palmerpenguins)
with(penguins, mean(body_mass_g, na.rm = TRUE))
```


A new version of `with`:

```{r}
with2 <- function(data, expr) {
  expr <- enquo(expr)
  eval_tidy(expr, data)
}
```


```{r}
with2(penguins, mean(body_mass_g, na.rm = TRUE))
```

---
### Another example: replicating `subset`


```{r}
subset2 <- function(data, rows) {
  rows <- enquo(rows)
  rows_val <- eval_tidy(rows, data)
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}

df <- subset2(penguins, species == "Adelie")
table(df$species)
```


---
## Using pronouns to avoid ambiguity

The data mask provides two pronouns: .data and .env.

  * .data$x always refers to x in the data mask.
  * .env$x always refers to x in the environmen

```{r}
x <- 1
df <- data.frame(x = 2)
```


```{r}
with2(df, .data$x)
```


```{r}
with2(df, .env$x)
```

There's no reason that should work. But it does and can be used to avoid ambiguity.

---
### Why does that work? 


  * `.data` and `.env` are actually exported from `rlang`
  
    * `.data` retrieves data-variables from the data frame
    
    * `.env` retrieves env-variables from the enviroment
  
  * They are not real data frames: they just act like them sometimes 
    
    * you can't take do `names(.data)` or `map` over it. 


---
### When is tidy evaluation actually beneficial?

A practical example:

```{r}
resample <- function(df, n) {
  idx <- sample(nrow(df), n, replace = TRUE)
  df[idx, , drop = FALSE]
}
```

You want to create a new function that resamples and subsamples in a single step


---
### Example continued: 

An approach that does not work:


```{r error = TRUE}
subsample <- function(df, cond, n = nrow(df)) {
  df <- subset2(df, cond)
  resample(df, n)
}

df <- data.frame(x = c(1, 1, 1, 2, 2), y = 1:5)
subsample(df, x == 1)
```

`subsample()` doesn’t quote any arguments so cond is evaluated normally (not in a data mask), and we get an error when it tries to find a binding for x.

---
### Example continued 


An approach that does work


```{r}
subsample <- function(df, cond, n = nrow(df)) {
  cond <- enquo(cond)

  df <- subset2(df, !!cond)
  resample(df, n)
}

subsample(df, x == 1)
```

---
### Tidy evaluation handles ambiguity well 

Consider the function that is meant to find all the rows of df where *x* is at least some threshold value:

```{r}
threshold_x <- function(df, val) {
  subset2(df, x >= val)
}
```

How can this go wrong?

  * if `val` is in `df`
  
  * if `x` is in the calling environment but not in `df`
  
---
### A more robust implementation

Here's a better implementation:

```{r error = TRUE}
threshold_x <- function(df, val) {
  subset2(df, .data$x >= .env$val)
}
```


Case 1: `x` is in the calling environment but not in `df`

```{r error = TRUE}
no_x <- data.frame(y = 1:3)
x <- 10
threshold_x(no_x, 2)
```

Case 2: If `val` is in `df`

```{r error = TRUE}
has_val <- data.frame(x = 1:3, val = 9:11)
threshold_x(has_val, 2)
```





---
### NSE in base R 


Two common patterns for NSE in base R:

  1. `substitute()` and evaluation in the caller environment using `eval()`
  
  2. `match.call()`, call manipulation, and evaluation in the caller environment
  
  
---
### NSE Using `substitute`


`substitute` returns the parse tree for the (unevaluated) expression expr, substituting any variables bound in env.

`eval`: evaluates an R expression. Its arguments are:

  * *expr*: an object to be evaluated.

  * *envir*: the environment in which expr is to be evaluated. May also be NULL, a list, a data frame, a pairlist or an integer as specified to sys.call.

  * *enclos*: Relevant when envir is a (pair)list or a data frame. Specifies the enclosure, i.e., where R looks for objects not found in *envir*.
  
---
### NSE in base example:  


How `subset` is used: 

```{r}
sample_df <- data.frame(a = 1:5, b = 5:1, c = c(5, 3, 1, 4, 1))


subset(sample_df, a >= 4)
```

How `subset` is implemented in base 

```{r}
subset_base <- function(data, rows) {
  rows <- substitute(rows)
  rows_val <- eval(rows, data, caller_env())
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}
```

---
### Problems with `base::subset`

  1.  always evaluates rows in the calling environment, but if ... has been used, then the expression might need to be evaluated elsewhere.
  
    * this means you cannot reliably work with functionals like `map()` or `lapply()`
  
  2. Calling `subset()` from another function requires some care: you have to use substitute() to capture a call to subset() complete expression, and then evaluate
  
  3. `eval()` doesn’t provide any pronouns so there’s no way to require part of the expression to come from the data. 


---
### Alternative to `subset`:  Using tidy evaluation

```{r}
subset_tidy <- function(data, rows) {
  rows <- enquo(rows)
  rows_val <- eval_tidy(rows, data)
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}
```


---
### NSE with `match.call`: Background

`match.call`: returns a call in which all of the specified arguments are specified by their full names.

```{r}
my_func <- function(input1, input2){
  match.call()
}
```

```{r}
my_func(input1 = 1)
```

```{r}
my_func(input2 = 1, 2)
```

```{r error =TRUE}
my_func(input3 = 2)
```




---
## Using match.call in NSE 

Steps in using `match.call` to do NSE:

  1. Capture the complete call 
  
  2. Modify it
  
  3. Evaluate the results

---  
### Example: `write.csv` 


```{r}
write.csv <- function(...) {
  call <- match.call(write.table, expand.dots = TRUE)

  call[[1]] <- quote(write.table)
  call$sep <- ","
  call$dec <- "."

  eval(call, parent.frame())
}
```

---
### Alternate implementation of `write.csv`

It could have been done like this: 

```{r}
write.csv <- function(...) {
  write.table(..., sep = ",", dec = ".")
}
```



---
### Wrapping modeling functions

Simplest possible wrapper:

```{r}
lm2 <- function(formula, data) {
  lm(formula, data)
}
```


```{r}
lm2(bill_length_mm ~ body_mass_g, penguins)
```

---
### A better wrapper function

```{r}
lm3 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  data <- enexpr(data)

  lm_call <- expr(lm(!!formula, data = !!data))
  expr_print(lm_call)
  eval(lm_call, env)
}

lm3(bill_length_mm ~ body_mass_g, penguins)
```

---
### Things to note


There are 3 key steps: 

  1. capture the unevaluated arguments using `enexpr()`, and capture the caller environment using `caller_env()`.

  2. generate a new expression using `expr()` and unquoting.

  3. evaluate that expression in the caller environment.

Nice side-effect: Unquoting can be used to generate formulas 

```{r}
y <- expr(bill_length_mm)
x1 <- expr(body_mass_g)
x2 <- expr(species)
lm3(!!y ~ !!x1 + !!x2, penguins)
```

---
### Potential problem situation: Mingling objects 

Problem: What if you want a function that resamples before training the model?


Something that doesn't work:
```{r error = TRUE}
resample_lm0 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_call <- expr(lm(!!formula, data = resample_data))
  expr_print(lm_call)
  eval(lm_call, env)
}

df <- data.frame(x = 1:10, y = 5 + 3 * (1:10) + round(rnorm(10), 2))
resample_lm0(y ~ x, data = df)
```

`lm_call` and `resample_data` are in different environments. 



---
### Example: Approach 1


Unquote the data frame into the call:

```{r}
df <- data.frame(x = 1:10, y = 5 + 3 * (1:10) + round(rnorm(10), 2))

resample_lm1 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_call <- expr(lm(!!formula, data = !!resample_data))
  expr_print(lm_call)
  eval(lm_call, env)
}
resample_lm1(y ~ x, data = df)$call
```



---
### Example continued: a cleaner approach

A cleaner approach:

  1. create a new environment that inherits from the caller 
  
  2. bind variables that you’ve created inside the function to that environment.


```{r}
resample_lm2 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_env <- env(env, resample_data = resample_data)
  lm_call <- expr(lm(!!formula, data = resample_data))
  expr_print(lm_call)
  eval(lm_call, lm_env)
}
resample_lm2(y ~ x, data = df)
```


---
### Abrupt ending: Overall takeaways 


  * There are many ways to do non-standard evaluation
  
  * Tidy evaluation is a good framework for applying NSE


 