---
title: "Advanced R"
subtitle: "Chapter 14"
author: |
  | Stephen Holsenbeck
  | Template & Layout: Tyler Grant Smith
institute: '`r fontawesome::fa("github", fill = "#f0f042")`&nbsp;[yogat3ch](http://github.com/yogat3ch)<br>`r fontawesome::fa("twitter", fill = "#f0f042")`&nbsp;[ascensionasana](http://twitter.com/ascensionasana)<br>`r fontawesome::fa("linkedin", fill = "#f0f042")`&nbsp;[holsenbeck](http://linkedin.com/in/holsenbeck)<br>`r fontawesome::fa("globe", fill = "#f0f042")`&nbsp;[themindful.life](http://www.themindful.life)'
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "custom.css"]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ['left', 'middle']
    includes:
      after_body: ["insert-logo.html", "use-fontawesome.html"]
---

<style>
.hljs-github .hljs {
color: #333;
background: #f8f8f8
}
</style>


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```
```{r libraries, include=FALSE}
# devtools::install_github("ropenscilabs/icon")
suppressPackageStartupMessages({
  library(tidyverse)
  library(R6)
  library(glue)
  library(xaringanthemer)
  library(tidyverse)
  library(glue)
  library(rlang)
  library(emo)
  requireNamespace("pkgsearch")
  requireNamespace("testthat")
})
```

```{r xaringan-themer, include = FALSE}
r_icon <- fontawesome::fa("r-project")

style_duo_accent_inverse(
  primary_color = "#f0f042",
  title_slide_text_color = "#f0f042",
  secondary_color = "#eb9641",
  header_font_google = google_font("PT Sans"),
  text_font_google   = google_font("Varela Round", "300", "300i"),
  base_font_size     = "18px",
  code_font_google   = google_font("Fira Code"),
  title_slide_background_image = "images/r6logo.png"
)
```

# R6
<div class="show-logo"></div>

<div class='container'>
```{r package_details, results='asis', echo=FALSE}
info <- library(help = "R6")$info
cat(paste0(
  info[[1]] %>% 
    stringr::str_replace_all("^([^:]*):", "<b>\\1</b>:"),
  collapse = "<br>"))

```
</div>

---
# R6 in the wild

#### How many?
```{r 'how_many'}
purrr::map_dbl(c("Depends", "Imports", "Suggests"), ~attr(do.call(pkgsearch::advanced_search, rlang::list2(!!.x := "R6")), "metadata")$total) %>%
  sum() %>% 
  glue(" packages on CRAN utilize `R6`.")
```
---
#### Notable R6 Packages:
```{r 'notable_r6', results='asis'}
r6_pkgs <- purrr::map(c("callr", "processx", "plumber", "promises", "websocket", "keras", "roxygen2", "QuantTools", "testthat"), ~pkgsearch::advanced_search(Package = .x))

htmltools::tags$ol(do.call(htmltools::tagList, purrr::map(r6_pkgs, ~htmltools::tags$li(htmltools::tags$p(glue::glue("{htmltools::tags$code(.x$package)}: {.x$title}"))))))
```
---
## 14.1 [Introduction](https://adv-r.hadley.nz/r6.html#introduction-13)
Two special properties:

 - R6 uses the encapsulated OOP paradigm, which means that methods belong to objects, not generics, and you call them like `object$method()`.

 - R6 objects are mutable, which means that they are modified in place, and hence have reference semantics.

---
### [Outline](https://adv-r.hadley.nz/r6.html#outline-12)

 - 14.2 introduces `R6::R6Class()`, the one function that you need to know to create R6 classes. 

 - 14.3 discusses the access mechanisms of R6: private and active fields.

 - 14.4 explores the consequences of R6’s reference semantics.

 - 14.5 describes why Hadley covers R6, rather than the base RC system.

---
## 14.2 [Classes & Methods](https://adv-r.hadley.nz/r6.html#r6-classes)
 - `classname` argument with `UpperCamelCase` convention.
 - `public` argument supplies a list of methods (functions) and fields (anything else) that make up the public interface of the object.
```{r inside_r6}
callr::r_process
```
**Always assign the result of `R6Class()` into a variable with the same name as the class**

---
- Construct a new object from the class by calling `new()`
```{r 'new'}
px <- callr::r_process$new(
  callr::r_process_options(
    func = rlang::as_function(~ {
      purrr::reduce(1:10, ~{
        Sys.sleep(.1)
        .x * .y
        }, .init = 1)
    })
  )
)
```
- call methods and access fields with `$`
```{r 'poll'}
purrr::walk(vector(length = 2), ~{
  Sys.sleep(1.2)
  print(px$is_alive())
}) 
px$get_result()
```
---

### 14.2.1 [Method Chaining](https://adv-r.hadley.nz/r6.html#method-chaining)
 - Commonly used in languages like Python and JavaScript
 - Method chaining is deeply related to the pipe
```{r method_chaining}
Accumulator <- R6::R6Class("Accumulator", list(
  sum = 0,
  add = function(x = 1) {
    self$sum <- self$sum + x 
    invisible(self)
  })
)
a <- Accumulator$new()
a$
  add(4)$
  add(4)$
  sum

```
 
---
### 14.2.2 [Important Methods](https://adv-r.hadley.nz/r6.html#r6-important-methods)
 - `$initialize()`: overrides the default behaviour of `$new()`
  - add checks for input arguments
  - if more expensive validation is required, use `$validate` separately.
 - `$print()`: allows you to override the default printing behaviour.
  - `$print()` should return `invisible(self)`.
 - methods are bound to individual objects IE previously created objects do not get new methods 
```{r 'sample'}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) {
    self$name <- name
    self$age <- age
  },
  print = function(...) {
    cat("Person: \n")
    cat("  Name: ", self$name, "\n", sep = "")
    cat("  Age:  ", self$age, "\n", sep = "")
    invisible(self)
  }
))
```
---
### 14.2.3 [Adding methods after creation](https://adv-r.hadley.nz/r6.html#adding-methods-after-creation)
 - `$set()`: Add new elements to an existing class    
```{r 'add_methods'}
Accumulator <- R6Class("Accumulator")
Accumulator$set("public", "sum", 0)
Accumulator$set("public", "add", function(x = 1) {
  self$sum <- self$sum + x 
  invisible(self)
})
```
---
### 14.2.4 [Inheritance](https://adv-r.hadley.nz/r6.html#inheritance)
- `inherit` argument takes an R6 class object to inherit from
```{r 'chatty'}
AccumulatorChatty <- R6Class("AccumulatorChatty", 
  inherit = Accumulator,
  public = list(
    add = function(x = 1) {
      cat("Adding ", x, "\n", sep = "")
      super$add(x = x)
    }
  )
)

x2 <- AccumulatorChatty$new()
x2$add(10)$add(1)$sum
```
 - `$add()` overrides the superclass implementation, but we can still delegate to the superclass implementation by using `super$`. (This is analogous to NextMethod() in S3)
 
---
### 14.2.5 [Introspection](https://adv-r.hadley.nz/r6.html#introspection)
```{r 'all methods'}
names(px)
```

---
## 14.3 [Controlling Access](https://adv-r.hadley.nz/r6.html#r6-access)
### 14.3.1 [Privacy](https://adv-r.hadley.nz/r6.html#privacy)
`R6Class` has two other arguments that work similarly to public:

 - `private`: a list of methods & fields that are only available from within the class, not outside of it.
   - Fields and methods defined in private are available within the methods using private$ instead of self$. You cannot access private fields or methods outside of the class.
 - `active`: allows you to use accessor functions to define dynamic, or active, fields.

---
Private fields cannot be accessed from outside of the class:
```{r private_fields}
Person <- R6Class("Person", 
  public = list(
    initialize = function(name, age = NA) {
      private$name <- name
      private$age <- age
    },
    print = function(...) {
      cat("Person: \n")
      cat("  Name: ", private$name, "\n", sep = "")
      cat("  Age:  ", private$age, "\n", sep = "")
    }
  ),
  private = list(
    age = NA,
    name = NULL
  )
)

hadley3 <- Person$new("Hadley")
hadley3
hadley3$name
```
---
### 14.3.2 [Active Fields](https://adv-r.hadley.nz/r6.html#active-fields)
Active fields are implemented using active bindings (Section 7.2.6). Each active binding is a function that takes a single argument: `value`
```{r active_fields}
Rando <- R6::R6Class("Rando",
  private = list(
    .random = 1
  ),
  public = list(
    initialize = function(init = 1) {
      private$.random <- init
    }
  ),
  active = list(
    random = function(value) {
      if (missing(value)) {
        private$.random <- runif(1)
      } else {
        stop("`$random` is read-only", call. = FALSE)
      }
    }
  )
)
```
---
```{r 'active_fields_example', error = TRUE}
x <- Rando$new()
x$random
x$random <- 4
x$random(4)
```

---
## 14.4 [Reference Semantics](https://adv-r.hadley.nz/r6.html#r6-semantics)
 - objects are not copied when modified
 - if you want a copy, you’ll need to explicitly `$clone()` the object
  - $clone() does not recursively clone nested R6 objects. If you want that, you’ll need to use $clone(deep = TRUE)
 - It is harder to reason about code that uses R6 objects because you need to understand more context.
 - It makes sense to think about when an R6 object is deleted, and you can write a `$finalize()` to complement the `$initialize()`.
 - If one of the fields is an R6 object, you must create it inside `$initialize()`, not `R6Class()`.
 
---
### 14.4.1 [Reasoning](https://adv-r.hadley.nz/r6.html#reasoning)

`f()` calls methods of `x` or `y`, it might modify them as well as `z`
```{r, eval = F}
x <- List$new(a = 1)
y <- List$new(b = 2)

z <- f(x, y)
```
 - write functions that either return a value, or modify their R6 inputs, but not both 
 
---
### 14.4.2 [Finalizer](https://adv-r.hadley.nz/r6.html#finalizer)
Think of `$finalize()` as a complement to `$initialize()`. Finalizers usually play a similar role to `on.exit()`

```{r temporary_file}
TemporaryFile <- R6Class("TemporaryFile", list(
  path = NULL,
  initialize = function() {
    self$path <- tempfile()
  },
  finalize = function() {
    message("Cleaning up ", self$path)
    unlink(self$path)
  }
))
tf <- TemporaryFile$new()
rm(tf)
```

---
### 14.4.3 [R6 fields](https://adv-r.hadley.nz/r6.html#r6-fields)
If you use an R6 class as the default value of a field, it will be shared across all instances of the object! Take the following code: we want to create a temporary database every time we call `TemporaryDatabase$new()`, but the current code always uses the same path.


```{r r6_fields}
TemporaryDatabase <- R6Class("TemporaryDatabase", list(
  con = NULL,
  file = TemporaryFile$new(),
  initialize = function() {
    self$con <- DBI::dbConnect(RSQLite::SQLite(), path = file$path)
  },
  finalize = function() {
    DBI::dbDisconnect(self$con)
  }
))

db_a <- TemporaryDatabase$new()
db_b <- TemporaryDatabase$new()

db_a$file$path == db_b$file$path
```
---
To fix the problem, we need to make sure it’s called every time that `TemporaryDatabase$new()` is called, i.e. we need to put it in `$initialize()`
```{r r6_fields_cont, eval = F}
TemporaryDatabase <- R6Class("TemporaryDatabase", list(
  con = NULL,
  file = NULL,
  initialize = function() {
    self$file <- TemporaryFile$new()
    self$con <- DBI::dbConnect(RSQLite::SQLite(), path = file$path)
  },
  finalize = function() {
    DBI::dbDisconnect(self$con)
  }
))
```
 
---
## In use

Discussion



