---
title: "Advanced R"
subtitle: "Chapter 10"
author: "R4DS Reading Group"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---


```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
library(tidyverse)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Sans Mono"),
  header_h1_font_size  = "2.25rem"
)
```


# Overview

- What is a function factory?
  - Function factories and manufactured functions
  - Manufactured function environments
  - Promises and `force`
  - {factory}
- Why use a function factory?
  - Stateful functions
  - ggplot2
  - Expensive calculations
  
```{r packages}
library(rlang)
library(ggplot2)
library(scales)
# remotes::install_github("jonthegeek/factory")
library(factory)
```

---

class: inverse, hide-logo

# What is a function factory?

---

# Function factories and manufactured functions

```{r factory-basics}
# Function factory
power1 <- function(exponent) {
  function(x) {
    x ^ exponent
  }
}

# Manufactured functions
square1 <- power1(2)
cube1 <- power1(3)
square1(8)
```

---

# Manufactured function environments

```{r factory-environments}
square1
cube1
c(fn_env(square1)$exponent, fn_env(cube1)$exponent)
```

---

# Promise dangers

Lazy evaluation + factories = **danger**

```{r bad-factory}
my_exponent <- 2
square1b <- power1(my_exponent)
my_exponent <- 3
square1b(2)
```

---

# Forcing evaluation

`force` forces evaluation

```{r repaired-factory}
power2 <- function(exponent) {
  force(exponent)
  function(x) {
    x ^ exponent
  }
}
```

*(technically just `exponent` instead of `force(exponent)` does the same thing)*

---

# {factory}

- I created a [package](https://github.com/jonthegeek/factory) to handle some of the fancy stuff.
- Maybe discuss internals in Chapter 19?

```{r factory}
power3 <- factory::build_factory(
  function(x) {
    x ^ exponent
  },
  exponent
)
my_exponent <- 2
square3 <- power3(my_exponent)
my_exponent <- 3
square3(2)
square3
```

---

class: inverse, hide-logo

# Why use a function factory?

---

# Stateful functions

```{r guessing-game}
new_guessing_game <- function() {
  target <- sample(1:100, 1)
  previous_diff <- NA_integer_
  function(guess) {
    if (guess %in% 1:100) {
      if (guess == target) {
        message("Correct!")
        return(invisible(TRUE))
      }
      new_diff <- abs(target - guess)
      if (is.na(previous_diff) || new_diff == previous_diff) {
        message("Try again!")
      } else if (new_diff < previous_diff) message("Warmer!")
      else message("Colder!")
      previous_diff <<- new_diff
    } else stop("Your guess should be between 1 and 100.")
    return(invisible(FALSE))
  }
}
```

---

# Stateful functions (cont)

```{r guess-hack, include = FALSE}
# Hack to make sure 50 gets all three possible responses.
guess <- new_guessing_game()
rlang::fn_env(guess)$target <- 30
new_guessing_game <- function() guess
```

```{r guess}
guess <- new_guessing_game()
guess(50)
guess(75)
guess(50)
guess(25)
guess(50)
```

---

# {ggplot2}

*Lots* of ggplot2 functions accept functions as arguments

```{r geom-help}
?ggplot2::geom_histogram
```

> `binwidth` The width of the bins. Can be specified as a numeric value **or as a function that calculates width from unscaled x.** Here, "unscaled x" refers to the original x values in the data, before application of any scale transformation. When specifying a function along with a grouping structure, the function will be called once per group...

---

# {scales}

The {scales} package is full of function factories.

```{r scales}
scales::number_format
```

---

# Expensive calculations

```{r expensive-simulation}
boot_model <- function(df, formula) {
  # Pretend these calculations would be slow
  mod <- lm(formula, data = df)
  fitted_vals <- unname(fitted(mod))
  resid_vals <- unname(resid(mod))
  rm(mod) # Or use {factory} and this won't be necessary!
  function() {
    fitted_vals + sample(resid_vals)
  }
}
boot_mtcars1 <- boot_model(mtcars, mpg ~ wt)
head(boot_mtcars1())
head(boot_mtcars1())
```

---

# Expensive calculations (cont)

```{r expensive-cont}
boot_mtcars1
head(rlang::fn_env(boot_mtcars1)$fitted_vals)
head(rlang::fn_env(boot_mtcars1)$resid_vals)
```

---

class: inverse, hide-logo

# Questions?
