---
title: "Advanced R"
subtitle: "Chapter 19: Quasiquotation"
author: |
  | Stephen Holsenbeck
  | Template & Layout: Jon Harmon
institute: '`r fontawesome::fa("github", fill = "white")`&nbsp;[yogat3ch](http://github.com/yogat3ch)<br>`r fontawesome::fa("twitter", fill = "white")`&nbsp;[ascensionasana](http://twitter.com/ascensionasana)<br>`r fontawesome::fa("linkedin", fill = "white")`&nbsp;[holsenbeck](http://linkedin.com/in/holsenbeck)<br>`r fontawesome::fa("globe", fill = "white")`&nbsp;[themindful.life](http://www.themindful.life)'
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    nature:
      highlightLines: true
      highlightStyle: github
      countIncrementalSlides: false
      titleSlideClass: ['left', 'middle']
    lib_dir: libs
    css: [xaringan-themer.css, addendum.css]
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
sass::sass(sass::sass_file("libs/addendum.scss"), output = "addendum.css", cache = FALSE)
```

```{r libraries, include = FALSE}
library(rlang)
library(knitr)
library(purrr)
library(flair)
```

## 19.1: [Introduction](https://adv-r.hadley.nz/quasiquotation.html#introduction-18)

* **Quotation** prevents evaluation
* **Unquotation** re-allows evaluation
* **Quasiquotation** allows you to mix the two
  * Useful (?) for metaprogramming
  * Let's see if that's true!

---

### [19.2: Motivation](https://adv-r.hadley.nz/quasiquotation.html#quasi-motivation)

```{r cement, results='hide'}
cement <- function(...) {
  args <- rlang::ensyms(...)
  paste(
    purrr::map(args, rlang::as_string), 
    collapse = " "
  )
}
```
--
```{r cement-call}
canonical <- paste(
  letters[c(3, 15, 14, 20, 18, 9, 22, 5, 4)],
  collapse = ""
)
cement(
  This, example, is, !!canonical
)
```

---

### [19.2.1: Vocabulary](https://adv-r.hadley.nz/quasiquotation.html#vocabulary)

* **Evaluated** arguments obey R's normal rules.
* **Quoted** arguments are captured and processed in a special way.

--
```{r vocabulary}
# Evaluated
mean(1:5)
1:5
```
--
```{r vocabulary-2}
# Quoted
library(rlang)
```
```{r vocabulary-quoted, eval = FALSE}
rlang
#> Error: object 'rlang' not found
```

---

### 19.2.1: Cont

* **quoting function**
  * a function that quotes 1 or more arguments
* non-standard evaluation (NSE)
<p style="font-size:.75em"><strong>quoting function</strong> == function that quotes 1 or more arguments == NSE</p>

--

**Kinda.** Technically NSE is about arguments only.

---

### [19.2.2: Exercises: Base](https://adv-r.hadley.nz/quasiquotation.html#exercises-56)


```{r motivation-examples-noflair-flaired, echo = FALSE}
code <- "library(MASS)

mtcars2 <- subset(mtcars, cyl == 4)

with(mtcars2, sum(vs))
sum(mtcars2$am)"
flair::decorate(code) %>% 
  flair::knit_print.with_flair()
```
---
### [19.2.2: Exercises: Base](https://adv-r.hadley.nz/quasiquotation.html#exercises-56)
```{r motivation-examples-flair, echo = FALSE}
flair::decorate(code) %>% 
  flair::flair("MASS") %>% 
  flair::flair("cyl == 4") %>% 
  flair::flair("sum(vs)") %>% 
  flair::flair("am") %>% 
  flair::knit_print.with_flair()
```

---

### [19.2.2: Exercises: Tidyverse](https://adv-r.hadley.nz/quasiquotation.html#exercises-56)

```{r motivation-examples-t-noflair, fig.show='hide', message=FALSE, warnings = FALSE, echo = FALSE}



code <- '
suppressWarnings({
  library(dplyr, warn.conflicts = FALSE)
  library(ggplot2, warn.conflicts = FALSE)
})
byCyl <- mtcars %>%
  group_by(cyl) %>%
  summarise(mean = mean(mpg), .groups = "drop_last")

ggplot(byCyl, aes(cyl, mean)) + geom_point()'
flair::decorate(code) %>% 
  flair::knit_print.with_flair()
```

---

### [19.2.2: Exercises: Tidyverse](https://adv-r.hadley.nz/quasiquotation.html#exercises-56)

```{r motivation-examples-t-flair, echo = FALSE}
flair::decorate(code) %>% 
  flair::flair("dplyr") %>% 
  flair::flair("ggplot2") %>% 
  flair::flair("cyl") %>% 
  flair::flair_rx("(?<=\\()mpg") %>% 
  flair::flair_rx("mean(?=\\))") %>% 
  flair::knit_print.with_flair()
```


---

## 19.3: [Quoting](https://adv-r.hadley.nz/quasiquotation.html#quoting)

* "We'll need a pair of functions" (~6 pairs, including base)
  * **directly supplied** vs
  * **indirectly supplied**

---

### 19.3.1: [Capturing expressions](https://adv-r.hadley.nz/quasiquotation.html#capturing-expressions)
  
* `{rlang}` capturing expressions:
  * `expr`:  quotes expression directly
```{r expr}
rlang::expr(x)
(function(x) rlang::expr(x)) (1 + 2 + 3)
```
--  
  * `enexpr`: quote the expression from the user
```{r enexpr}
(function(x) rlang::enexpr(x)) (1 + 2 + 3)
```


---

### 19.3.1: [Capturing expressions](https://adv-r.hadley.nz/quasiquotation.html#capturing-expressions)
 * `exprs`: quotes multiple expressions directly
```{r exprs}
rlang::exprs(x = 1 ^ 2, y = 2 ^ 3, z = 3 ^ 4)
```
--
* `enexprs`: quotes multiple expressions from the user
```{r enexprs}
(function(...) rlang::enexprs(...)) (I(4 * 10e8), sd(c(4,5)))
```

---

### 19.3.2: [Capturing symbols](https://adv-r.hadley.nz/quasiquotation.html#capturing-symbols)

* `sym`: converts the character data in an object or character directly to a symbol
```{r sym}
rlang::sym("this")
chr <- "mpg"
sym_object <- rlang::sym(chr)
# sneak peak
dplyr::select(mtcars, !!sym_object) %>% head(1)
```
---
### 19.3.2: Cont.

* `ensym`: converts user specified character data in an object or character directly to a symbol
```{r ensym}
(function(data, symbol) dplyr::select(data, !!rlang::ensym(symbol))) (mtcars, "mpg") %>% head(1)

(function(data, symbol) dplyr::select(data, all_of(!!rlang::ensym(symbol)))) (mtcars, chr) %>% head(1)
```
---
### 19.3.2: [Capturing symbols](https://adv-r.hadley.nz/quasiquotation.html#capturing-symbols)

* `syms`: converts multiple characters in an object or character vector directly to a symbol
```{r syms}
rlang::syms(c("this", "that", "the other"))
chr <- c("mpg", "hp")
sym_object <- rlang::syms(chr)
dplyr::select(mtcars, !!!sym_object) %>% head(1)
```
---
### 19.3.2: Cont.

* `ensyms`: converts user specified character vector to symbols
```{r ensyms}
(function(data, symbols) dplyr::select(data, all_of(!!!rlang::ensyms(symbols)))) (mtcars, chr) %>% head(1)

(function(data, ...) dplyr::select(data, !!!rlang::ensyms(...))) (mtcars, "mpg", "drat", "hp") %>% head(1)
```

---

### 19.3.3: [With base R](https://adv-r.hadley.nz/quasiquotation.html#with-base-r)

* `expr` = `quote`
* `exprs` = `alist`
* The base function closest to `enexpr` is `substitute`

```{r 'substitute'}
(function(x) substitute(x ^ 2)) (x + y)
```

* The equivalent to `enexprs` is an undocumented feature of `substitute` 
```{r substitute_special}
(function(...) as.list(substitute(...()))) (x = 1, y = 10 * z)
```
  
 * `bquote` provides a limited form of quasiquotation, see [Section 19.5](https://adv-r.hadley.nz/quasiquotation.html#base-nonquote)
 * `~`, the formula, is a quoting function that also captures the environment. See [Section 20.3.4](https://adv-r.hadley.nz/evaluation.html#quosure-impl).
---

### 19.3.4: [Substitution](https://adv-r.hadley.nz/quasiquotation.html#substitution)

* `substitute` can be used to... substitute

```{r substitution}
(function(x) substitute(x * 2)) (a + b + c)

```

---

### 19.3.4: [Substitution](https://adv-r.hadley.nz/quasiquotation.html#substitution)

* `r htmltools::tags$span(style = "color:#0c5460;", "Recommendation:")` use the second argument to substitute to make it explicit

```{r substitution_with_list}
(function(a) substitute(x + a, list(x = 4))) (a = 10 * x)
```
--
`r htmltools::tags$span(style = "color:#856404;", "Warning:")` notice that it inactivates expression substitution when a list is supplied to substitute

```{r substitution2}
(function(a) substitute(x + a)) (a = 10 * x)

```

---

### 19.3: [Substitution - Summary ](https://adv-r.hadley.nz/quasiquotation.html#summary-1)
<div class="inline-table"><table class="table table-sm">
<caption>
<span id="tab:quoting-rlang">Table 19.1: </span> rlang quasiquoting functions</caption>
<thead><tr class="header">
<th></th>
<th>Developer</th>
<th>User</th>
</tr></thead>
<tbody>
<tr class="odd">
<td>One</td>
<td><code><a href="https://dplyr.tidyverse.org/reference/tidyeval-compat.html">expr()</a></code></td>
<td><code><a href="https://dplyr.tidyverse.org/reference/tidyeval-compat.html">enexpr()</a></code></td>
</tr>
<tr class="even">
<td>Many</td>
<td><code><a href="https://ggplot2.tidyverse.org/reference/tidyeval.html">exprs()</a></code></td>
<td><code><a href="https://dplyr.tidyverse.org/reference/tidyeval-compat.html">enexprs()</a></code></td>
</tr>
</tbody>
</table></div>
<hr>

--

<div class="inline-table"><table class="table table-sm">
<caption>
<span id="tab:quoting-base">Table 19.2: </span> base R quoting functions</caption>
<thead><tr class="header">
<th></th>
<th>Developer</th>
<th>User</th>
</tr></thead>
<tbody>
<tr class="odd">
<td>One</td>
<td><code><a href="https://rdrr.io/r/base/substitute.html">quote()</a></code></td>
<td><code><a href="https://rdrr.io/r/base/substitute.html">substitute()</a></code></td>
</tr>
<tr class="even">
<td>Many</td>
<td><code><a href="https://rdrr.io/r/base/list.html">alist()</a></code></td>
<td><code><a href="https://rdrr.io/r/base/list.html">as.list(substitute(...()))</a></code></td>
</tr>
</tbody>
</table></div>
---
## 19.4: [Unquoting](https://adv-r.hadley.nz/quasiquotation.html#unquoting)

* _quasiquotation_ means you can unquote selectively
* Base R doesn't allow you to selectively unquote (mostly)
* Unquoting means "replace the symbol with stored object itself"

---

### 19.4.1: [Unquoting one argument](https://adv-r.hadley.nz/quasiquotation.html#unquoting-one-argument)

* The confusingly beautiful heart of quasiquotation: **!!**
* Use it to unquote one thing
* `!!` preserves operator precedence because it works with expressions

```{r bang-bang}
x <- expr(-1)
y <- "a character"
rlang::expr(f(!!x, y))
rlang::expr(f(x, !!y))
rlang::expr(mean(1:3) + mean(4:6))
rlang::expr(!!mean(1:3) + !!mean(4:6))
```


---

### 19.4.2: [Unquoting a function](https://adv-r.hadley.nz/quasiquotation.html#unquoting-a-function)

* Now we start to move toward metaprogramming
* `expr(!!f(x, y))` unquotes the result of `f(x, y)`, so an extra pair of parentheses is needed

```{r replace-function}
replace_f <- function(func) {
  rlang::expr((!!rlang::enexpr(func))(x, y)) # Would be better if we used the actual formals...
}
replace_f(mean)
replace_f(rlang::enexpr)
```
---
### 19.4.2: [Unquoting a function Cont.](https://adv-r.hadley.nz/quasiquotation.html#unquoting-a-function)
* to avoid nested parentheses it can be clearer to use `rlang::call2`:
```{r 'call2'}
replace_f <- function(func) {
  call2(enexpr(func), expr(x), expr(y))
}
replace_f(mean)
```

---
### 19.4.3 [Unquoting a missing argument](https://adv-r.hadley.nz/quasiquotation.html#unquote-missing)
Will it work?
```{r 'missing', error = TRUE}
arg <- missing_arg()
expr(foo(!!arg, !!arg))
```

workaround: `rlang::maybe_missing`
```{r missing2}
expr(foo(!!maybe_missing(arg), !!maybe_missing(arg)))
```

---
### 19.4.4 [Unquoting in special forms](https://adv-r.hadley.nz/quasiquotation.html#unquoting-in-special-forms)
```{r special, error = TRUE}
x <- expr(x)
expr(df$!!x)
```
```{r special2}
(function(x) expr(`$`(df, !!enexpr(x)))) (y)
```

---
### 19.4.5: [Unquoting many arguments](https://adv-r.hadley.nz/quasiquotation.html#unquoting-many-arguments)

* `!!!` is to `!!` as `exprs` is to `expr`
* works with or without names
* **"unquote-splice"** because it **unquotes** the list then **splices** them in as if they were all separate arguments
* can be used in any rlang function that takes ... regardless of whether or not ... is quoted or evaluated.

```{r bang-bang-bang}
multi_arg <- list(a = 1, b = 2, c = "other")
rlang::expr(f(!!!multi_arg, another_arg))
call2("f", !!!multi_arg, expr(y))
```

---

### 19.4.6: [The polite fiction of !!](https://adv-r.hadley.nz/quasiquotation.html#the-polite-fiction-of)

* `!!` and `!!!` don't actually exist

```{r fictional-functions, error = TRUE}
rlang::`!!`
rlang::`!!!`
```
---
### 19.4.6: Cont.

 * it's actually the built-in
```{r negation, error = TRUE}
!TRUE
!!TRUE
!!!TRUE
```
---
### 19.4.6: Cont.
The biggest downside to using a fake operator is that you might get silent errors when misusing !! outside of quasiquoting functions:
* silently incorrect results when working with numeric values
```{r 'silent_error'}
df <- data.frame(x = 1:5)
y <- 100
with(df, x + !!y)
```
 * If you actually need double negation use `!(!x)`
 
---

### 19.4.7: [Non-standard ASTs](https://adv-r.hadley.nz/quasiquotation.html#non-standard-ast)
if you inline more complex objects, their attributes are not printed:
```{r 'Complex objects'}
(x1 <- expr(class(!!data.frame(x = 10))))
eval(x1)
```
-- 
The solution? Use `rlang::expr_print`
```{r expr_print}
expr_print(x1)
```

---
### 19.5 [Non-Quoting](https://adv-r.hadley.nz/quasiquotation.html#base-nonquote)

`bquote`

* does not provide an unquote-splice operator that allows you to unquote multiple expressions stored in a list
* lacks the ability to handle code accompanied by an environment

```{r 'bquote'}
xyz <- bquote((x + y + z))
bquote(-.(xyz) / 2)
```

---

### 19.5 Non-quoting techniques 1/4
There are four basic forms seen in base R:

1. A pair of quoting and non-quoting  _functions_
  * closely related to `$`: `subset`, `transform`, and `with`
  * `<-/assign` and `::/getExportedValue` work similarly

```{r 'quoting_non-quoting'}
x <- list(var = 1, y = 2)
var <- "y"
x$var
x[[var]]
```

---

### 19.5 Non-quoting techniques 2/4

2. A pair of quoting and non-quoting _arguments_
* `data` and `save` work similarly
```{r 'pair of quoting and unquoting'}
x <- 1
rm(x)
ls(pattern = "x")
y <- 2
vars <- c("y", "vars")
rm(list = vars)
ls(pattern = "^y$")
```

---

### 19.5 Non-quoting techniques 3/4

3. An argument that controls whether a different argument is quoting or non-quoting
```{r 'diff_arg'}
library(MASS)
pkg <- "MASS"
library(pkg, character.only = TRUE)
```


---
### 19.5 Non-quoting techniques 4/4

4. Quoting if evaluation fails
  * `ls`, `page`, and `match.fun`
```{r 'quote if eval fails', eval = FALSE}
# Shows help for var
help(var)

var <- "mean"
# Shows help for mean
help(var)

var <- 10
# Shows help for var
help(var)
```

---
## [19.6 ... (dot-dot-dot)](https://adv-r.hadley.nz/quasiquotation.html#tidy-dots)

* This section is the leftovers, it isn't about `...` per se
* `:=` "colon-equals" or "walrus"
  * Another fiction to trick R into working how we want
  * LHS of `=` can't be evaluated, so we trick R

```{r walrus_assignment}
var <- "my_var_name"
val <- 1:3
tibble::tibble(!!var := val)
```

---
### [19.6.1 - Examples](https://adv-r.hadley.nz/quasiquotation.html#examples)

Dots (...) can be converted to a usable list with `rlang::list2` or [`rlang::dots_list`]((https://adv-r.hadley.nz/quasiquotation.html#dots_list))
```{r using_dots}
set_attr <- function(.x, ...) {
  attr <- rlang::list2(...)
  attributes(.x) <- attr
  .x
}

attrs <- list(x = 1, y = 2)
attr_name <- "z"

1:10 %>%
  set_attr(w = 0, !!!attrs, !!attr_name := 3) %>% 
  str()
```

---

### [19.6.2: exec()](https://adv-r.hadley.nz/quasiquotation.html#exec)

* `rlang::exec` is similar to `base::do.call`
* Allows you to use unquoting to do fancy things

```{r exec1}
# Directly
exec("mean", x = 1:10, na.rm = TRUE, trim = 0.1)

# Indirectly
args <- list(x = 1:10, na.rm = TRUE, trim = 0.1)
exec("mean", !!!args)

# Mixed
params <- list(na.rm = TRUE, trim = 0.1)
exec("mean", x = 1:10, !!!params)
```

---
### [19.6.2: exec() - Cont.](https://adv-r.hadley.nz/quasiquotation.html#exec)
```{r exec2}
# Indirect arguments
arg_name <- "na.rm"
arg_val <- TRUE
exec("mean", 1:10, !!arg_name := arg_val)

# Functionalization
x <- c(runif(10), NA)
funs <- c("mean", "median", "sd")
purrr::map_dbl(funs, exec, x, na.rm = TRUE)
```

---
### [19.6.3: dots_list()](https://adv-r.hadley.nz/quasiquotation.html#dots_list)

 * `.ignore_empty` allows you to control exactly which arguments are ignored.
 * `.homonyms` controls what happens if multiple arguments use the same name
 * `.preserve_empty` controls what to do with them empty arguments that are not ignored
 
---
### [19.6.4: With base R](https://adv-r.hadley.nz/quasiquotation.html#do-call)


`do.call`: The first argument, what, gives a function to call. The second argument, args, is a list of arguments to pass to that function.

```{r do.call}
do.tibble <- function(...) {
  do.call(tibble::tibble, rlang::dots_list(...))
}
do.tibble(a = 1:5)
```


---
## 19.7: Case studies

### [19.7.1: lobstr::ast()](https://adv-r.hadley.nz/quasiquotation.html#lobstrast)

```{r ast}
z <- expr(foo(x, y))
lobstr::ast(z)
lobstr::ast(!!z)
```

---
### [19.7.2: Map-reduce to generate code](https://adv-r.hadley.nz/quasiquotation.html#map-reduce-to-generate-code)

---
### [19.7.3: Slicing an array](https://adv-r.hadley.nz/quasiquotation.html#slicing-an-array)
```{r 'slicing an array'}
slice <- function(x, along, index) {
  stopifnot(length(along) == 1)
  stopifnot(length(index) == 1)
    
  nd <- length(dim(x))
  indices <- rep(list(missing_arg()), nd)
  indices[[along]] <- index
  
  expr(x[!!!indices])
}
x <- array(sample(30), c(5, 2, 3))
slice(x, 1, 3)
slice(x, 2, 2)
slice(x, 3, 1)

```
---
### [19.7.4: Creating functions](https://adv-r.hadley.nz/quasiquotation.html#new-function)
`rlang::new_function`: a function that creates a function from its three components: arguments, body, and (optionally) an environment:

```{r 'creating functions'}
power <- function(exponent) {
  new_function(
    exprs(x = ), 
    expr({
      x ^ !!exponent
    }), 
    caller_env()
  )
}
power(0.5)
```
---
### [19.7.5: Exercises](https://adv-r.hadley.nz/quasiquotation.html#exercises-60)

2. Re-implement the Box-Cox transform defined below using unquoting and `new_function`:

```{r bc}
bc <- function(lambda) {
  if (lambda == 0) {
    function(x) log(x)
  } else {
    function(x) (x ^ lambda - 1) / lambda
  }
}
```
---
### [19.7.5: Exercises Cont. ](https://adv-r.hadley.nz/quasiquotation.html#exercises-60)

```{r bc2}
bc2 <- function(lambda) {
  rlang::new_function(
    rlang::exprs(lambda = ),
    purrr::when(lambda, 
         . == 0 ~ expr(log(x)),
          ~ expr((x ^ lambda - 1) / lambda))
  )
}
bc2(0)
bc2(2)
```
---
## [19.8 History](https://adv-r.hadley.nz/quasiquotation.html#history)

Quasiquotation was first used in a programming language, Lisp, in the mid-1970s.103 Lisp has one quoting function `` ` ``, and uses `,` for unquoting. Most languages with a Lisp heritage behave similarly. For example, Racket (`` ` `` and @), Clojure (`` ` `` and ~), and Julia (: and @) all have quasiquotation tools that differ only slightly from Lisp. 
---
### [19.8 History Cont](https://adv-r.hadley.nz/quasiquotation.html#history)

My attempt to resolve these limitations led to the lazyeval package (2014-2015). Unfortunately, my analysis of the problem was incomplete and while lazyeval solved some problems, it created others. It was not until I started working with Lionel Henry on the problem that all the pieces finally fell into place and we created the full tidy evaluation framework (2017).
 - Hadley Wickham
 
---
.center[# Thank you!]

