---
title: 'Chapter 11: Function Operators'
subtitle: 'Tony ElHabr'
author: 'R4DS Reading Group'
output:
  xaringan::moon_reader:
    nature:
      highlightLines: true
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  # cache = TRUE,
  include = TRUE,
  message = FALSE, 
  warning = FALSE
)
```

```{r xaringan-themer, eval=T, echo=F, include=F}
# xaringan::inf_mr()
library(xaringanthemer)
library(tidyverse)
style_mono_light(
  base_color = '#3092FF',
  header_font_google = google_font('Josefin Slab'),
  text_font_google = google_font('Raleway', '300', '300i')
)
```

<style>
img {
  display: block;
  margin: 0 auto;
}
hide {
  display: none;
}
.remark-slide-content h1 {
  font-size: 45px;
}
.remark-code, .remark-inline-code {
  font-family: var(--code-font-family);
}
h1 {
  font-size: 2em;
  margin-block-start: 0.67em;
  margin-block-end: 0.67em;
}
.remark-slide-content {
  font-size: 16px
}
.remark-code {
  font-size: 14px;
}
code.r {
  font-size: 14px;
}
pre {
  margin-top: 0px;
  margin-bottom: 0px;
}
.red {
  color: #FF0000;
}

.footnote {
  color: #800020;
  font-size: 9px;
}

</style>

# What are function operators (FO)?

Chapter 9 is about functionals. Chapter 10 is about function factories. What makes function operators different?

--

```{r eval=T, echo=F, include=T}
functional_programming_summary <-
  tibble(
    Term = c('Functionals', 'Function Factory', 'Function Operator'),
    `Required Input` = c('Function', '', 'Function'),
    `Optional Input` = c('Vector', 'Vector, Function', 'Vector'),
    Output = c('Vector', 'Function', 'Function')
  ) %>% 
  gt::gt() %>% 
  gt::tab_options(
    # column_labels.vlines.width = px(3),
    # column_labels.border.bottom.width = 
    column_labels.font.weight = 'bold'
  )
functional_programming_summary
```

```{r eval=F, echo=F, include=F}
gt::gtsave(functional_programming_summary, filename = here::here('Presentations', 'Week1R1', 'functional_programming_summary.png'))
```

--

FOs are probably best studied by thinking about how they operate on functions

+ __Behavioral FO__: Changes the behavior of a function, e.g. logging, running a function only when necessary

  + `memoise::memoise()`

+ __Output FO__: Manipulates the output of a function

  + `purrr::possibly()`, `purrr::safely()`, `purrr::quietly()`

+ __Input FO__: Maniuplates the input of a function

  + `purrr::partial()`


<hide>
`purrr::safely()` docs: ... "They are all adverbs because they modify the action of a verb (a function)."
</hide>

---

# Behavioral FO Example #1

Passing in only function `f`

```{r }
slowly <- function(f){
  force(f)
  function(n, ...){
    cat(
      glue::glue('Sleeping for {n} seconds.'), 
      sep = '\n'
    )
    Sys.sleep(n)
    # Need to do this to prevent extra printing.
    res <- utils::capture.output(f(n, ...))
  }
}

purrr::walk(
  seq(0.4, 0.2, by = -0.2), 
  slowly(cat) #<<
)
```

.footnote[
https://gist.github.com/ColinFay/d32cf4c9c5fb8d849f12a4e98d6c0549
]

---

# Behavior FO Example #1

Now with an additional input, vector `n`

```{r }
slowly <- function(f, n){
  force(f)
  force(n)
  function(...){
    stopifnot(is.numeric(n))
    cat(
      glue::glue('Sleeping for {n} seconds.'), 
      sep = '\n'
    )
    Sys.sleep(n)
    f(...)
  }
}

slowly_cat <- slowly(cat, 0.5)
slowly_cat('hello world')
```

---

# Behavioral FO Example #2

```{r }
twice <- function(f){
  force(f)
  function(...){
    f(...)
    f(...)
  }
}

purrr::walk(
  c('hello', 'world'), 
  twice(cat), #<<
  sep = '\n' # Passed to `cat()` via `...`
)
```

.footnote[
Inspiration: https://realpython.com/primer-on-python-decorators/
]

---

# Behavioral FO Example #2

With `python` `r fontawesome::fa('python')`

```{r setup-python, include=F}
library(reticulate)
config <- reticulate::py_config()
reticulate::use_python(config$python)
```

```{python }
def do_twice(f):
    def wrapper(*args, **kwargs):
        f(*args, **kwargs)
        f(*args, **kwargs)
    return wrapper
    
@do_twice
def say(x):
    print(x)

```

```{python }
list(map(say, ['hello', 'world']))
```

```{r }
```

---

# Behavioral FO Example #3

```{r }
download_beers <- function(name, verbose = TRUE) {
  base_url <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/'
  url <- glue::glue('{base_url}{name}.csv')
  if(verbose) {
    cat(glue::glue('Downloading {name}.csv'), sep = '\n')
  }
  readr::read_csv(url)
}
```

```{r eval=F, echo=F, include=F}
# Only really need this while developing this presentation.
memoise::forget(download_beers_quickly)
```

Using `memoise::memoise()` for caching

```{r echo=F}
download_beers_quickly <- memoise::memoise(download_beers)

bench::mark(
  download_beers('brewer_size', verbose = FALSE),
  download_beers_quickly('brewer_size', verbose = FALSE)
) %>% 
  dplyr::select(expression, min, median)
```

<hide>
This is useful for scraping... and also a tutorial, where there are repeated actions across code chunks :P.
Should memoisization be used when scraping? It depends. If the data really won't ever change, then it should be fine.
</hide>

---

# Behavioral FO Example #4

Testing the speed of `memoise::memoise()`

```{r }
# Forgive the contrived function.
slow_function <- function(x) {
  Sys.sleep(0.2)
  x * runif(1)
}
fast_function <- memoise::memoise(slow_function)
```

```{r eval=F, echo=F, include=F}
# memoise::forget(fast_function)
```

```{r }
system.time(slow_function(1))
system.time(slow_function(1)) #<<
```

```{r }
system.time(fast_function(1))
system.time(fast_function(1)) #<<
```

---

# Behavioral FO Example #4

Even if you've changed the inputs since the most recent call, it will still be fast.

```{r }
system.time(fast_function(2))
system.time(fast_function(3))
system.time(fast_function(2)) #<<
```
--

In fact, it remembers everything from the same session (assuming you haven't used `memoise::forget()`).

```{r }
system.time(fast_function(1))
system.time(fast_function(2))
system.time(fast_function(3))
```

---

# Behavioral FO Example #5

How does memoisization work with a factory?


```{r }
fast_function_factory <- function(x) {
    memoise::memoise(function(...) {
    slow_function(...)
  })
}
```

<br />

```{r }
system.time(fast_function(42))
system.time(fast_function(42)) #<<
```

```{r }
system.time(fast_function_factory(-1))
system.time(fast_function_factory(-1)) #<<
```

<hide>
TBH this seems silly. I was just curious about it.
</hide>

---

# Input FO Example #1

```{r }
stat_robust <- function(f, ...) {
  function(...) {
    f(..., na.rm = TRUE)
  }
}
mean_robust <- stat_robust(mean)
min_robust <- stat_robust(min)
quantile_robust <- stat_robust(quantile)
```

<br />

--

.pull-left[
```{r }
x1 <- 1L:10L
mean_robust(x1)
min_robust(x1)
quantile_robust(x1, 0.25)
```
]

.pull-right[
```{r }
x2 <- x1; x2[1] <- NA
mean_robust(x2)
min_robust(x2)
quantile_robust(x2, 0.25)
```
]

---

# Input FO Example #1

Using `purrr::partial()`

```{r }
mean_partial <- partial(mean, na.rm = TRUE)
min_partial <- partial(min, na.rm = TRUE)
quantile_partial <- partial(quantile, na.rm = TRUE, ... = )
```

--

Without `purrr::partial()`

```{r }
mean_wrapper <- function(...) {
  mean(..., na.rm = TRUE)
}
```

---

# Input FO Example #2

Using the `brewer_size` data set

```{r eval=T, echo=F, include=F}
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
```

```{r }
brewer_size %>% 
  summarize_at(
    vars(total_barrels, total_shipped),
    list(mean = mean, mean_robust = mean_robust)
  ) %>% 
  mutate_all(
    ~scales::number(., scale = 1e-3, big.mark = ',', suffix = ' k')
  ) %>% 
  glimpse()
```

---

# Output FO Example #1

Using `purrr::safely()`


```{r }
download_beers_safely <- purrr::safely(download_beers)
```

```{r }
brewing_material <- download_beers_safely('brewing_material') # Oops!
brewing_material

brewing_materials <- download_beers_safely('brewing_materials') # Good
brewing_materials$result %>% head(5)
```

<hide>
In a real-world example, it is more likely that the error is with something unexpected on the web-page, not in the user's code.
</hide>

---

# Output FO Example #2

Using `purrr::possibly()`


```{r }
download_beers_possibly <- purrr::possibly(download_beers, otherwise = tibble())
```

```{r }
brewing_material <- download_beers_possibly('brewing_material') # Oops!
brewing_material
```

<hide>
`purrr::safely()` also has an `otherwise` argument, but I think it makes more sense to use in conjuction with `possibly()`.
</hide>

---

# Output FO Example #3

Using `purrr::quietly()`


```{r }
download_beers_quietly <- purrr::quietly(download_beers)
```

```{r }
brewing_materials <- download_beers_quietly('brewing_materials') # Oops!
names(brewing_materials)
brewing_materials$result %>% head(5)
```

---

# Combining FOs Example

```{r }
nms <- c('woops', 'brewing_materials', 'beer_taxed', 'brewer_size', 'beer_states') %>% 
  setNames(., .)
```

```{r }
download_beers_nicely <- slowly(download_beers_safely, 0.1)
beers <- nms %>%
  map(., 
      ~download_beers_nicely(..1) %>% 
        purrr::pluck('result')
    )

beers %>% map(dim) %>% str()
```
---

# Combining FOs Example

And a real-world use-case for `purrr::reduce()`!

```{r }
beers %>% 
  purrr::discard(is.null) %>% 
  purrr::reduce(dplyr::left_join) %>% 
  dim()
```

---

# FOs in the Wild

+ `{scales}` and `{ggplot2}`'s `scale_(color|fill)_*()`

+ `{glue}` with it's transformers

+ Sparingly in `{styler}` and `{lintr}`

--

+ `{plumber}` uses R6 `r emo::ji('surprised')`

---

# FIN
