---
title: "Advanced R Exercises"
output:
  html_document:
    toc: true
    toc_float: true
    css: styles.css
---


A companion to [Advanced R](https://adv-r.hadley.nz/rcpp.html#acknowledgments) and supplement to [Advanced R Solutions](https://advanced-r-solutions.rbind.io/names-and-values.html#copy-on-modify). 

Writing up solutions for those missing or incomplete within the solutions manual. __Maybe with time and collaboration we can even [submit our answers as a PR](https://github.com/Tazinho/Advanced-R-Solutions)!__

:::question
Solutions in a question box are solved in the manual and copied over here in an effort to edit and clarify.

Otherwise we've copied over questions from the solutions manual that have been marked as `started` __[and have an orange A]{.started}__ in an effort to complete them, and we will be attempting ot answer questions that have been unanswered entirely.
:::

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("r4ds.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; 
                        right:0;
                        padding:0;
                        width: 15%;
               height 15%')
```

# Functions

## Function fundamentals
  
1. __[Q]{.Q}__: Given a name, like `"mean"`, `match.fun()` lets you find a function. Given a function, can you find its name? Why doesn't that make sense in R?

    __[A]{.started}__: A name can only point to a single object, but an object can be pointed to by 0, 1, or many names. What are names of the functions in the following block?

    ```{r}
  function(x) sd(x) / mean(x)
    
  f1 <- function(x) (x - min(x)) / (max(x) - min(x))
  f2 <- f1
  f3 <- f1
    ```


## `...`  (dot-dot-dot)

3. __[Q]{.Q}__: How does the `chdir` parameter of `source()` compare to `in_dir()`? Why might you prefer one approach to the other?

    The `in_dir()` approach was given in the book as
    ```{r, eval = FALSE}
    in_dir <- function(dir, code) {
      old <- setwd(dir)
      on.exit(setwd(old))
      
      force(code)
    }
    ```
    
   __[A]{.started}__:`in_dir()` takes a path to a working directory as an argument. First the working directory is changed accordingly. `on.exit()` ensures that the modification to the working directory are reset to the initial value when the function exits.
    
   In `source()` the `chdir` argument specifies if the working directory should be changed during the evaluation of the `file` argument (which in this case has to be a pathname). 
   
<!-- HW: I think I'm more intersted in supplying a path vs. a logical value here -->

## Function forms

3. __[Q]{.Q}__: Explain why the following code fails:
    ```{r, eval = FALSE}
    modify(get("x"), 1) <- 10
    #> Error: target of assignment expands to non-language object
    ```
    
   __[A]{.started}__: First, let's define `x` and recall the definition of `modify()` from the textbook:
    
    ```{r}
    x <- 1:3
    
    `modify<-` <- function(x, position, value) {
      x[position] <- value
      x
    }
    ```

