---
title: "Functions"
subtitle: "Advanced R"
author: "Meenakshi Kushwaha"
date: "2020/09/02 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader
---
class:middle, center
# The Basics
---
# Three components
- ###Argument | formals
  - list of arguments
- ###Body
  - code inside the function
- ###Environment
  - data structure that determines **how the function finds the values** associated with the names
  - specified implicitly based on *where* you define the function
  
.footnote[The Basics]  
---
# Exception - Primitive functions

- Exist primarily in C so their `formals()`, `body()`, and `environment()` are NULL
- Only found in base R
- Examples - `sum()`, `[`

```r 
typeof(sum)
#> [1] "builtin"
typeof(`[`)
#> [1] "special"
```
.footnote[The Basics]  
---
# First-class functions
- R functions are objects in their own right
- No special syntax for defining and naming a function
- Typically, create a function object using `function` and bind it to a name with `<-`
  - Binding step is not necessary
```{r eval=FALSE}
lapply(mtcars, function(x) length(unique(x)))
Filter(function(x) !is.numeric(x), mtcars)
integrate(function(x) sin(x) ^ 2, 0, pi)
```
.footnote[The Basics]  
---
# Invoking a function
Use `do.call` if arguments are already in a data structure
```{r, eval=FALSE}
args <- list(1:10, na.rm = TRUE)
do.call(mean, args)
> [1] 5.5
```
.footnote[The Basics]  
---
class:middle, center
# Function composition

---
# Nesting

```{r eval=FALSE}
x <- runif(100)
sqrt(mean(square(deviation(x))))
#> [1] 0.274
```

$\checkmark$ Concise, well suited for short sequences  
**x**  Can be hard to read

.footnote[Function composition]
???
Note- code highlight differs if r is not within parentheses
.footnote[Function composition]
---
# Intermediate Objects

```{r eval=FALSE}
out <- deviation(x)
out <- square(out)
out <- mean(out)
out <- sqrt(out)
out
#> [1] 0.274
```

**x**  intermediate names required  
$\checkmark$ This is useful only when intermediate objects are important  

.footnote[Function composition]
---
# Piping 

```{r eval=FALSE}  
x %>%
  deviation() %>%
  square() %>%
  mean() %>%
  sqrt()
#> [1] 0.274
```
$\checkmark$ More readable code  
**x**  Requires third party package   
.footnote[Function composition]
---
class: center, middle
# Lexical scoping
### How does R find the value associated with a name?

**"It looks up the values of names based on how a function is defined, not how it is called"**

---
# Four rules of lexical scoping

- ### Name masking
- ### Functions versus variables
- ### A fresh start
- ### Dynamic lookup
.footnote[Lexical scoping]
---
## Name masking
.footnote[Lexical scoping]
.pull-left[Names defined within the function mask names outside the function 
```{r eval=FALSE}
x <- 10
y <- 20
g02 <- function() {
  x <- 1
  y <- 2
  c(x, y)
}
g02()
#> [1] 1 2
```
]
--
.pull-right[If a name isn’t defined inside a function, R looks one level up.       
```{r eval=FALSE}
x <- 2
g03 <- function() {
  y <- 1
  c(x, y)
}
g03()
#> [1] 2 1
y
#> [1] 20
```
]     

--
The same rules apply if a function is defined inside another function. 

---
#Functions vs variables
- Since functions are objects, scoping rules also apply to functions
- What happens if you use the same name for a function and non-function object?
  - R ignores non-function objects when looking for that value

```{r eval=FALSE}
g09 <- function(x) x + 100
g10 <- function() {
  g09 <- 10
  g09(g09)
}
g10()
#> [1] 110
```
 
 !!! Using same name for different things is a bad idea!!!   
.footnote[Lexical scoping]
???
g09 takes two different values
best to not use same names (caution from Hadley) but knowing this can help us troubleshoot

---
# A fresh start
- Every time a function is called a new environment is created to host its execution. 
- Each invocation is completely independent
```{r eval=FALSE}
g11 <- function() {
  if (!exists("a")) {
    a <- 1
  } else {
    a <- a + 1
  }
  a
}

g11()
g11()
```


.footnote[Lexical scoping]
---
#Dynamic Look up
.footnote[Lexical scoping]
- R looks for values when the function is run, not when the function is created. 


```{r eval=FALSE}
g12 <- function() x + 1
x <- 15
g12()
#> [1] 16

x <- 20
g12()
#> [1] 21
```
--
- If there is a a spelling mistake in the code, you won’t get an error message when you create the function
- Use `codetools` to detect the external dependencies
--
```{r eval=FALSE}
codetools::findGlobals(g12)
#> [1] "+" "x"
```

---
class: middle, center
# Lazy evaluation
In R, function arguments are **lazily evaluated**: they’re only evaluated if accessed.
.left[
```{r eval=FALSE}
h01 <- function(x) {
  10
}
h01(stop("This is an error!"))
#> [1] 10
```
]

---
# Promises
#### Lazy evaluation is powered by "promise": a data structure 
A promise has three components
- **An expression**, like `x + y`, which gives rise to the delayed computation
- **An environment** where the expression should be evaluated
- **A value**, which is computed and cached the first time a promise is accessed when the expression is evaluated in the specified environment
```{r eval=FALSE}
double <- function(x) { 
  message("Calculating...")
  x * 2
}
h03 <- function(x) {
  c(x, x)
}
h03(double(20))
> Calculating...
> [1] 40 40
```

.footnote[Lazy evaluation]
---
# Default arguments
Default arguments can be defined in terms of
- other arguments
- variables defined later in the function  
  - !!! Not recommended because
      - makes the code harder to understand
      - need to know the exact order of evaluation

.footnote[Lazy evaluation]
---
# Missing arguments
you can use `missing()` to see if the argument's value comes from the user or default
```{r eval=FALSE}
h06 <- function(x = 10) {
  list(missing(x), x)
}
str(h06())
> List of 2
>  $ : logi TRUE
>  $ : num 10
str(h06(10))
#> List of 2
#>  $ : logi FALSE
#>  $ : num 10
```

.footnote[Lazy evaluation]
???
First example uses default and returns TRUE
second example uses user supplied so logi is FALSE
---
# ... (dot dot dot) | ellipsis

- function can take any number of named and unnamed arguments
- Pass arbitrary no. of functions to your function

```{r eval=FALSE}
red.plot <- function(x, y, ...) {
  plot(x, y, col="red", ...)
}
red.plot(1:10, 1:10, xlab="My x axis", ylab="My y axis")

```

.footnote[Lazy evaluation. Example from https://nicercode.github.io/guides/functions/]
---
class: middle, center
# Exiting a function
???
functions exit in two ways, they either return a value, indicating success, or they throw an error
---
# Implicit vs explicit return
```{r eval=FALSE}
j01 <- function(x) {
  if (x < 10) {
    0
  } else {
    10
  }
}
j01(5)
#> [1] 0
j01(15)
#> [1] 10
```

```{r eval=FALSE}
j02 <- function(x) {
  if (x < 10) {
    return(0)
  } else {
    return(10)
  }
}
```

.footnote[Exiting a function]
---
# Invisible values
- The most common function that returns invisibly is `<-`
- Can make invisible by applying `invisible()` to the last value


```{r eval=FALSE}
j04 <- function() invisible(1)
j04()
```
- Verify if the value exists by using `print()` or wrapping in `()`

```{r eval=FALSE}
print(j04())
#> [1] 1

(j04())
#> [1] 1
```

.footnote[Exiting a function]
---
# Errors
If a function cannot complete its assigned task, it should throw an error with `stop()` indicating the termination of function execution

```{r eval=FALSE}
j05 <- function() {
  stop("I'm an error")
  return(10)
}
j05()
#> Error in j05(): I'm an error
```
.footnote[Exiting a function]
---
# Exit handlers
Undo any changes made to the global environment
```{r eval=FALSE}
j06 <- function(x) {
  cat("Hello\n")
  on.exit(cat("Goodbye!\n"), add = TRUE)
  if (x) {
    return(10)
  } else {
    stop("Error")
  }
}
j06(TRUE)
> Hello
> Goodbye!
> [1] 10
j06(FALSE)
> Hello
> Error in j06(FALSE): Error
> Goodbye!
```

.footnote[Exiting a function]
???
can use it to switch back directories on exit

---
class: middle, center
# Function forms

---
#Prefix form
- Any function can be written in prefix form

- Function name comes before the arguments eg. `foofy(a, b, c)`

$\checkmark$ Helps better understand the structure of the language

```{r eval=FALSE}
x + y
`+`(x, y)

names(df) <- c("x", "y", "z")
`names<-`(df, c("x", "y", "z")) 

for(i in 1:10) print(i)
`for`(i, 1:10, print(i))
```
.footnote[Function forms]
---
# Using prefix form
- By position, like help(mean).
- Partial matching, like help(top = mean).
- By name, like help(topic = mean)  
!!! Use positional matching only for the first one or two arguments. Avoid using positional matching for less commonly used arguments, and **never use partial matching**.!!!



.footnote[Function forms]
???
Use positional matching only for the first one or two arguments; they will be the most commonly used, and most readers will know what they are.
---
# Infix form
- the function name comes in between its arguments, like `x + y`
- User defined functions begin with `%`

```{r eval=FALSE}
`%+%` <- function(a, b) paste0(a, b)
"new " %+% "string"
#> [1] "new string"

```
- names can contain any sequence of characters except `%`


.footnote[Function forms]

---
# Replacement
- Functions that replace values by assignment, like `names(df) <- c("a", "b", "c")`
- They have the special name `xxx<-`, must have arguments named `x` and value, and must return the modified object

```{r eval=FALSE}
`second<-` <- function(x, value) {
  x[2] <- value
  x
}
```


.footnote[Function forms]
---
# Special forms
- They don't have a consistent form like `[[`, `if`, and `for`
- Knowing the name of the function that underlies a special form is useful for getting documentation
- All special forms are implemented as primitive functions; this means printing these functions is not informative

```{r eval=FALSE}
`for`
> .Primitive("for")
```

.footnote[Function forms]
---
class: middle, center
#Thank you
.footnote[kmeena@uw.edu, twitter: @envhealthspeak]
