---
title: 'Chapter 21: Translating R Code'
subtitle: 'Pavitra Chakravarty'
author: 'R4DS Reading Group'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
---

<style>
 .remark-code-line {
   font-size: 11px;
 }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  cache.lazy = FALSE,
  include = TRUE,
  message = FALSE, 
  warning = FALSE
)
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
library(methods)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

<style>
hide {
  display: none;
}
.remark-slide-content h1 {
  font-size: 45px;
}
h1 {
  font-size: 2em;
  margin-block-start: 0.67em;
  margin-block-end: 0.67em;
}
.remark-slide-content {
  font-size: 16px
}
.remark-code {
  font-size: 14px;
}
code.r {
  font-size: 14px;
}
pre {
  margin-top: 0px;
  margin-bottom: 0px;
}
.red {
  color: #FF0000;
}

.footnote {
  color: #800020;
  font-size: 9px;
}

</style>

# Translating R code {#translation}

```{r, include = FALSE}
library(dbplyr)
library(methods)
library(rlang)
library(purrr)
source("dsl-html-attributes.r")
```

## Introduction

The combination of first-class environments, lexical scoping, and metaprogramming gives us a powerful toolkit for translating R code into other languages. One fully-fledged example of this idea is dbplyr, which powers the database backends for dplyr, allowing you to express data manipulation in R and automatically translate it into SQL. You can see the key idea in  `translate_sql()` which takes R code and returns the equivalent SQL:

```{r}
translate_sql(x ^ 2)
translate_sql(x < 5 & !is.na(x))
translate_sql(!first %in% c("John", "Roger", "Robert"))
translate_sql(select == 7)
```
---

## S3 CLASSES

The easiest way to do this is to create an S3 class that distinguishes between regular text (that needs escaping) and HTML (that doesn't).

```{r escape}
html <- function(x) structure(x, class = "advr_html")

print.advr_html <- function(x, ...) {
  out <- paste0("<HTML> ", x)
  cat(paste(strwrap(out), collapse = "\n"), "\n", sep = "")
}

html("1")
```

---


We then write an escape generic. It has two important methods:

* `escape.character()` takes a regular character vector and returns an HTML
  vector with special characters (`&`, `<`, `>`) escaped.

* `escape.advr_html()` leaves already escaped HTML alone.

```{r}
escape <- function(x) UseMethod("escape")

escape.character <- function(x) {
  x <- gsub("&", "&amp;", x)
  x <- gsub("<", "&lt;", x)
  x <- gsub(">", "&gt;", x)

  html(x)
}

escape.advr_html <- function(x) x
```

Some examples

```{r}
escape("This is some text.")
escape("x > 1 & y < 2")
escape(escape("This is some text. 1 > 2"))
escape(html("<hr />"))
```

---

## Basic tag functions

Now, lets look at the other tags in a html doc

`<p>`- HTML tag can have both attributes (e.g., id or class) and children (like `<b>` or `<i>`). Given that attributes are named and children are not, it seems natural to use named and unnamed arguments for them respectively. For example, a call to `p()` might look like:

```{r, eval = FALSE}
p("Some text. ", b(i("some bold italic text")), class = "mypara")
```

We could list all the possible attributes of the `<p>` tag in the function definition, but that's hard because there are many. Instead, we'll use `...` and separate the components based on whether or not they are named. With this in mind, we create a helper function that wraps around `rlang::list2()`

---

## Lets refer back to 19.6 ... (dot-dot-dot)

One place we could use `list2()` is to create a wrapper around `attributes()` that allows us to set attributes flexibly:

```{r}
numeric <- function(...) {
  dots <- list2(...)
  num <- as.numeric(dots)
  set_names(num, names(dots))
}


x<-numeric(a=1,b=2,c=3,d=4)
x

test <-  "q"
numeric( 1,!!!x,!!test:=9)
```

---

## Getting back to getting attributes

```{r named}
dots_partition <- function(...) {
  dots <- list2(...)
  
 if (is.null(names(dots))) {
  is_named <- rep(FALSE, length(dots))
} else {
  is_named <- names(dots) != ""
}
  
  list(
    named = dots[is_named],
    unnamed = dots[!is_named]
  )
}

str(dots_partition(a = 1, 2, b = 3, 4))
```

```{r p}

p <- function(...) {
  dots <- dots_partition(...)
  attribs <- html_attributes(dots$named)
  children <- map_chr(dots$unnamed, escape)

  html(paste0(
    "<p", attribs, ">",
    paste(children, collapse = ""),
    "</p>"
  ))
}

p("Some text", id = "myid")
p("Some text", class = "important", `data-value` = 10)
```

---

## Now for the children tags 

```{r}
tag <- function(tag) {
  new_function(
    exprs(... = ),
    expr({
      dots <- dots_partition(...)
      attribs <- html_attributes(dots$named)
      children <- map_chr(dots$unnamed, escape)

      html(paste0(
        !!paste0("<", tag), attribs, ">",
        paste(children, collapse = ""),
        !!paste0("</", tag, ">")
      ))
    }),
    caller_env()
  )
}
tag("b")
```

---

## Void Tags

```{r}
p <- tag("p")
b <- tag("b")
i <- tag("i")
p("Some text. ", b(i("some bold italic text")), class = "mypara")
```
```{r}
void_tag <- function(tag) {
  new_function(
    exprs(... = ),
    expr({
      dots <- dots_partition(...)
      if (length(dots$unnamed) > 0) {
        abort(!!paste0("<", tag, "> must not have unnamed arguments"))
      }
      attribs <- html_attributes(dots$named)

      html(paste0(!!paste0("<", tag), attribs, " />"))
    }),
    caller_env()
  )
}

img <- void_tag("img")
img
img(src = "myimage.png", width = 100, height = 100)
```
---

## Lets pull in all tags

```{r}
tags <- c("a", "abbr", "address", "article", "aside", "audio",
  "b","bdi", "bdo", "blockquote", "body", "button", "canvas",
  "caption","cite", "code", "colgroup", "data", "datalist",
  "dd", "del","details", "dfn", "div", "dl", "dt", "em",
  "eventsource","fieldset", "figcaption", "figure", "footer",
  "form", "h1", "h2", "h3", "h4", "h5", "h6", "head", "header",
  "hgroup", "html", "i","iframe", "ins", "kbd", "label",
  "legend", "li", "mark", "map","menu", "meter", "nav",
  "noscript", "object", "ol", "optgroup", "option", "output",
  "p", "pre", "progress", "q", "ruby", "rp","rt", "s", "samp",
  "script", "section", "select", "small", "span", "strong",
  "style", "sub", "summary", "sup", "table", "tbody", "td",
  "textarea", "tfoot", "th", "thead", "time", "title", "tr",
  "u", "ul", "var", "video"
)

void_tags <- c("area", "base", "br", "col", "command", "embed",
  "hr", "img", "input", "keygen", "link", "meta", "param",
  "source", "track", "wbr"
)
```

---

## A lot of these tags are the same as base R function

So do not include in global environment or in a package 
```{r}
html_tags <- c(
  tags %>% set_names() %>% map(tag),
  void_tags %>% set_names() %>% map(void_tag)
)

```
---

## Finally the DSL

```{r}
with_html <- function(code) {
  code <- enquo(code)
  eval_tidy(code, html_tags)
}
```
Lets execute our DSL

```{r}
with_html(
  body(
    h1("A heading", id = "first"),
    p("Some text &", b("some bold text.")),
    img(src = "myimg.png", width = 100, height = 100)
  )
)
```

```{r}
with_html(
  body(
    title("facebook", lang = "en", dir = "ltr"),
    p("This is &", b("Facebook")),
    img(src = "myimg.png", width = 100, height = 100)
  )
)
```