---
title: "Advanced R by Hadley Wickham"
subtitle: "Chapter 14: R6"
author: "Ezra Porter"
date: "Nov 5 2020"
output: 
  xaringan::moon_reader:
    lib_dir: libs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{css, echo=FALSE}
.remark-code, .remark-inline-code {
    background: #F0F0F0;
}

.show-only-last-code-result pre + pre:not(:last-of-type) code[class="remark-code"] {
    display: none;
}
```

# Big Picture

- In functional OOP, like S3, methods belong to functions.

- In **encapsulated OOP**, like R6, methods belong to objects.

<br><br>

- R6 objects are always modified in place and never copied on modify

  - Powerful for abstracting complex objects with lots of self-contained components you might want to update
  
  - Can produce spooky results and spookier code if you're not careful 

<br>

All we'll need is

```{r}
library(R6)
```


---

# Creating

Create R6 classes by calling `R6::R6Class()` and passing a list of methods and fields

Using exercise 14.2.6.1 as an example:

```{r}
BankAccount <- R6Class("BankAccount", list(
  balance = 0,
  deposit = function(x) {
    self$balance <- self$balance + x 
    invisible(self)
  },
  withdraw = function(x) {
    self$balance <- self$balance - x 
    invisible(self)
  })
)
```

`self$` lets methods reference other fields or methods internal to the object

---

# Using

Create a new instance of a class with the `$new()` method

```{r}
checking <- BankAccount$new()
```

Access fields and methods with `$`

```{r}
checking$balance

checking$deposit(10)

checking$balance
```

---

# Using cnt'd
## 14.2.1 Method Chaining

Methods called for their side-effects (like setting internal values) can be chained together

```{r}
checking$withdraw(10)$withdraw(10)

checking$balance
```

<br>

This is powered by having side-effect methods return the object invisibly

```r
withdraw = function(x) {
    self$balance <- self$balance - x 
    invisible(self)
  }
```

---

# 14.2.2 Important Methods

Some methods affect the behavior of objects in special ways

`$initialize()` overrides the default behavior of `$new()`

```{r}
BankAccount <- R6Class("BankAccount", list(
  balance = 0,
  pwd = NULL,
  initialize = function(pwd) {
    self$pwd <- pwd
  },
  deposit = function(x, pwd) {
    stopifnot(pwd == self$pwd)
    self$balance <- self$balance + x 
    invisible(self)
  },
  withdraw = function(x, pwd) {
    stopifnot(pwd == self$pwd)
    self$balance <- self$balance - x 
    invisible(self)
  })
)
```

---

# 14.2.2 Important Methods

Some methods affect the behavior of objects in special ways

`$print()` determines how the object is printed

```{r, include = FALSE}
BankAccount$set("public", "print", function(...) {
  cat("Balance:", scales::dollar(self$balance))
  invisible(self)
})
```

```r
BankAccount <- R6Class("BankAccount", list(
...
  print = function(...) {
  cat("Balance:", scales::dollar(self$balance))
  invisible(self)
  }
...
```

```{r}
savings <- BankAccount$new(pwd = "dont-tell")
try(savings$deposit(10, "password123"))
(savings$deposit(10, "dont-tell"))
```

---

# 14.2.2 Important Methods

**Beware!** Objects encapsulate methods so our old `BankAccount` objects don't retroactively get newly created methods

```{r}
checking
checking <- BankAccount$new(pwd = "dont-tell")
checking
```

Make sure you rebuild objects when you alter a class during interactive use

---

# 14.2.4 Inheritance

R6 classes can be subclasses of other R6 classes. Define that relationship using the `inherit` argument to `R6Class()`

```{r, include=FALSE}
SocialistBankAccount <- R6Class("SocialistBankAccount",
  inherit = BankAccount,
  public = list(
    check_balance = function() {
      if (self$balance > 100000) {
        cat("From each according to their ability!")
        self$balance <- 100000
      } else if (self$balance < 0) {
        cat("To each according to their need!")
        self$balance <- 100
      }
    }, deposit = function(x, pwd) {
      super$deposit(x, pwd)
      self$check_balance()
    }, withdraw = function(x, pwd) {
      super$withdraw(x, pwd)
      self$check_balance()
    })
)
```

```r
SocialistBankAccount <- R6Class("SocialistBankAccount",
  inherit = BankAccount,
  public = list(
    check_balance = function() {
      if (self$balance > 100000) {
        cat("From each according to their ability!")
        self$balance <- 100000
      } else if (self$balance < 0) {
        cat("To each according to their need!")
        self$balance <- 100
      }
    },
    deposit = function(x, pwd) {
      super$deposit(x, pwd)
      self$check_balance()
    },
    withdraw = function(x, pwd) {
    ...
```

---

# 14.2.4 Inheritance

Our subclass inherits the methods and fields we don't explicitly overwrite from its super class

```{r}
common_fund <- SocialistBankAccount$new(pwd = "dont-tell")
common_fund
```

--

`super$` allows us to refer to superclass methods and thereby "delegate" like with `NextMethod()` in S3

```r
deposit = function(x, pwd) {
      super$deposit(x, pwd)
      self$check_balance()
    }
```

---

# 14.2.4 Inheritance

```r
deposit = function(x, pwd) {
      super$deposit(x, pwd)
      self$check_balance()
    }
```

```{r}
common_fund$withdraw(10, "dont-tell")
common_fund
```

--

R6 objects also get S3 classes which automatically reproduce the sub/superclass relationships

```{r}
class(common_fund)
```

---

# 14.3 Controlling Access

Right now users have full access to internal elements of our objects

```{r}
checking$pwd
```

We can use the `private` argument of `R6Class()` to set components for internal use

```{r, include = FALSE}
SecureBankAccount <- R6Class("SecureBankAccount",
  public = list(
    balance = 0,
    initialize = function(pwd) {
      private$pwd <- pwd
    },
    deposit = function(x, pwd) {
      stopifnot(pwd == private$pwd)
      self$balance <- self$balance + x 
      invisible(self)
    }
  ),
  private = list(pwd = NULL)
)
```

```r
SecureBankAccount <- R6Class("SecureBankAccount",
  public = list(
    balance = 0,
    initialize = function(pwd) {
      private$pwd <- pwd
    },
... More methods ...
  ),
  private = list(pwd = NULL)
)
```

---

# 14.3 Controlling Access

```r
SecureBankAccount <- R6Class("SecureBankAccount",
  public = list(
    balance = 0,
    initialize = function(pwd) {
      private$pwd <- pwd
    },
... More methods ...
  ),
  private = list(pwd = NULL)
)
```

```{r}
secure_checking <- SecureBankAccount$new("dont-tell")
secure_checking$pwd
```

Just reference `private$` in methods rather than `self$`

---

# 14.4 Reference Semantics

R6 objects are always modified in place. To get a copy you can use the `$clone()` method.

--

The fact that methods of an object can change the object itself makes code harder to reason about.

Hadley's example:

What can we say about the effect of this line of code on `x` and `y` given that they're base objects? Given that they're R6 objects?

```r
z <- f(x, y)
```

--

<br>

If `f` calls methods of `x` and `y` it might change them. In our `BankAccount` example the only thing our methods did was change internal values.

---

# 14.4 Reference Semantics

But therein lies the power:

<blockquote>
14.6.2.3
Why can’t you model a bank account or a deck of cards with an S3 class?
</blockquote>

--

<br>

S3 objects are copied when they're changed so the best you could do is have a generic function return a modified version of the object

---

# 14.4.3 R6 Fields (a cautionary tale)

R6 objects behave unintuitively when the default value of a field is another R6 object

```{r}
Number <- R6Class("Number", list(
  value = 0,
  increment = function() {
    self$value <- self$value + 1
  })
  )

NumberPointer <- R6Class("NumberPointer", list(
  number = Number$new()
  ))
```

<br>

The instance of `Number` will be shared across **all** instances of `NumberPointer`

---

class: show-only-last-code-result

# 14.4.3 R6 Fields (a cautionary tale)

```{r}
x <- NumberPointer$new()
y <- NumberPointer$new()
```

--

```{r}
x$number$value
```

--

```{r}
y$number$value
```

--

```{r}
x$number$increment()
```

--

```{r}
x$number$value
```

--

```{r}
y$number$value
```

--

Avoid this by making sure objects are initialized **within** a method so you get a new instance every time

---

# How have I used R6?

I haven't!

--

But this reminded me of how some machine learning and optimization algorithms are implemented in Python