---
title: "Advanced R"
subtitle: "Chapter 3: Vectors"
author: "Vajresh Balaji"
institute: "@bvajresh"
date: " `r Sys.Date()` "
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
---


```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

# Outline

- 3.2  Atomic Vectors
- 3.3 Attributes
- 3.4 S3 Atomic Vectors
- 3.5 Lists
- 3.6 Dataframes and Tibbles
- 3.7 NULL

---
## Vectors
 - 2 types of vectors
  - Atomic
  - List

---
## Atomic Vectors
- 4 common types
```{r}
    a <- c(1,2,3,4) #Integer
    b <- c(TRUE, FALSE, T, F) #Logical
    c <- c(1.2, 2.3, 5.0) #Double
    d <- c("apple", "banana") #Character
``` 
- Rare types of Atomic Vectors: Raw and Complex

---
## Missing Values

- R uses `NA` to represent missing values.
```{r}
x <- c(NA, 5, NA, 10)
x == NA
```

- Use `is.na()` to check for missing values
```{r}
is.na(x)
```

---
## Testing
- Type of vectors can be tested with `is.*()` function.
- `is.logical()`, `is.integer()`, `is.double()`, and `is.character()`
- Avoid using `is.vector()`, `is.atomic()`, and `is.numeric()`

---
## Coercion
- For atomic vectors, type is a property of the entire vector.
- When attempting to combine different types of elements, they will be *coerced* in a fixed order.
- character -> double -> integer -> logical

```{r}
str(c("a", 1))
```
- Coercion happens automatically. 
```{r}
x <- c(FALSE, FALSE, TRUE)
as.numeric(x)

sum(x) #Total number of TRUEs
```
- Using `as.*()` allows us to deliberately coerce. 
  - `as.logical()`,`as.integer()`, `as.double()` and `as.character()`

---
## Attributes
- Attributes can be added to atomic vectors to build data structures like Arrays, Matrices, Factors or date-times.
- They can be individually set and retrieved using `attr()`
```{r}
car = "CR-V"
attr(car,'manufacturer') <- 'Honda'
attr(car, 'manufacturer')
```
- Set multiple attributes using
```{r}
car2 <- structure("Model S", manufacturer = "Tesla", year = 2020)
```

- You can retrieve multiple attributes by using
```{r}
attributes(car2)
```
---

## Names

```{r}
x <- c(apple = 'a', banana = 'b') # 1
x
y <- c('a', 'b')
names(y) <- c('apple', 'banana') # 2
y
setNames(y, c('apple', 'banana')) # 3
```

<small>Source: TonyElHabr Chapter 3 slide 8</small>
---
## Dimensions
- Allows vector to behave like a matrix or an array.
```{r}
a <- matrix(1:6, nrow = 2, ncol = 3)
a
b <- matrix(1:6, c(1, 3, 2))
b
```

---
## Unusual Behaviour
- Vectors without `dim` are thought of as 1-dimensionals but are `NULL`
- Matrices with 1 row or col and 1-dimensional arrays print the same but behave differently
- Use `str()` to reveal the differences.

---
## S3 Atomic Vectors
- Objects that have a `class` attribute. 
- 4 types of S3 vectors
  - factor (categorical)
  - Date (date)
  - POSIXct (date-time)
  - duration (difftime)

---
## Factors
- Vectors that only contain pre-defined values
- Used for Categorical Data
```{r}
sex_char <- c("m", "m", "m")
sex_factor <- factor(sex_char, levels = c("m", "f"))
table(sex_factor)


grade <- ordered(c("b", "b", "a", "c"), levels = c("c", "b", "a"))
grade
```
- Many base R functions automatically convert character vectors into factors.
- Use `stringsAsFactors = FALSE` to supress this behaviour.

---

### Dates, POSIXct & Duration
- All built on top of double vectors.
- Dates
```{r}
today <- Sys.Date()
typeof(today)
attributes(today)
```
- Date-times
  - Two types of storing date-time: POSIXct & POSIXlt
  - Underlying value represents number of seconds since Jan 1, 1970.
  - `tzone` attribute

---

- Duration
  - Amount of time between date/date-time pairs. 
  - Stored in `difftimes`
  - `units` attribute to determine how integer should be interpreted.
```{r}
one_week_1 <- as.difftime(1, units = "weeks")
one_week_1
typeof(one_week_1)
attributes(one_week_1)
```
---

## Lists
- Can be of any atomic type or contain other lists.
```{r}
    e <- list(1, TRUE, 1.2, "apple", list(2, 4, 6))
```
- Elements of a list are references.
- `c()` combines several lists into one if there given a combination of atomic vectors and lists.
```{r}
l4 <- list(list(1, 2), c(3, 4))
str(l4)
```

---
## Data Frames
- S3 Vectors that are built on top of lists. 
```{r}
df1 <- data.frame(x = 1:3, y = letters[1:3])
typeof(df1)
attributes(df1)
```
- Constraint
  - Length of each vector must be the same

---
## Tibble
- Share the same structure as data frames
- class vectors are longer
- Default behaviour: `stringsAsFactors = FALSE`
- Discourages rownames
- "Nicer" Printing

---

## NULL
- NULL has a unique type
```{r}
typeof(NULL)
length(NULL)
```
- Common uses
  - Represent an empty vector
```{r}
c()
```
  - Represent an absent vector
