---
title: "Advanced R by Hadley Wickham"
subtitle: "Chapter 4: Subsetting"
author: "Mike Jeziorski"
date: "20 Aug 2020"
output: 
  xaringan::moon_reader:
    lib_dir: libs
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter contents

- 4.2 Selecting multiple elements

- 4.3 Selecting a single element

- 4.4 Subsetting and assignment

- 4.5 Applications
---
## 4.2 Selecting multiple elements

- `[ ]` are used to select any number of elements from a vector

### 4.2.1 Atomic vectors
```{r, eval=FALSE}
x <- c(2.1, 4.2, 3.3, 5.4)
```

#### Six ways to select elements from an atomic vector

- Positive integers: select indicated elements
```{r, eval=FALSE}
x[c(1, 3)] # Don't forget to use c() for multiple indices
# [1] 2.1 3.3
```



- Negative integers: remove indicated elements
```{r, eval=FALSE}
x[-2] # Negative and positive integers cannot be combined
# [1] 2.1 3.3 5.4
```


---
- Logical vectors: select elements that match `condition = TRUE`
```{r, eval=FALSE}
x[c(TRUE, TRUE, FALSE, FALSE)]
# [1] 2.1 4.2

x[x > 3]
# [1] 4.2 3.3 5.4

# Note the recycling behavior of R, which can cause problems

x[c(TRUE, FALSE)] == x[c(TRUE, FALSE, TRUE, FALSE)]

# NA in the index returns NA

x[c(TRUE, FALSE, NA, TRUE)]
# [1] 2.1  NA 5.4
```
---
- Empty index: returns the original vector
      - Can be useful for multidimensional objects
```{r, eval=FALSE}
x[ ]
# [1] 2.1 4.2 3.3 5.4
```
- Zero: returns an empty vector
```{r, eval=FALSE}
x[0]
# numeric(0)
```
- Character vector (used for a named vector)
```{r, eval=FALSE}
(y <- setNames(x, letters[1:4]))
#   a   b   c   d 
# 2.1 4.2 3.3 5.4 
```
Indices can be reused
```{r, eval=FALSE}
y[c("a", "a", "a")]
#   a   a   a 
# 2.1 2.1 2.1
```
Partial matching of strings is not allowed with `[ ]`  
Avoid using a factor as an index, because R will use the underlying integer
---
### 4.2.2 Lists

- `[ ]` subsets a list from a list

- `[[ ]]` extracts an element from a list

- `$` also extracts an element from a list

### 4.2.3 Matrices and arrays

- Three ways to subset matrices and arrays
      1. Multiple vectors
      
      2. Single vector
      
      3. Matrix
---
```{r, eval=FALSE}
a <- matrix(1:9, nrow = 3)
colnames(a) <- c("A", "B", "C")

a[1:2, ]
#      A B C
# [1,] 1 4 7
# [2,] 2 5 8

a[c(TRUE, FALSE, TRUE), c("B", "A")]
#      B A
# [1,] 4 1
# [2,] 6 3

a[0, -2]
#      A C

dim(a[0, -2])
# [1] 0 2

a[1, ]
# A B C 
# 1 4 7

dim(a[1, ])
# NULL
```
---
Matrices and arrays are reconfigured vectors, so they can be indexed with single values (indexing down each column in turn)
```{r, eval=FALSE}
vals <- outer(1:5, 1:5, FUN = "paste", sep = ",")
vals
#      [,1]  [,2]  [,3]  [,4]  [,5] 
# [1,] "1,1" "1,2" "1,3" "1,4" "1,5"
# [2,] "2,1" "2,2" "2,3" "2,4" "2,5"
# [3,] "3,1" "3,2" "3,3" "3,4" "3,5"
# [4,] "4,1" "4,2" "4,3" "4,4" "4,5"
# [5,] "5,1" "5,2" "5,3" "5,4" "5,5"

vals[c(4, 15)]
# [1] "4,1" "5,3"
```
---
### 4.2.4 Data frames and tibbles
- Data frames act like both matrices and lists
      - A single index extracts element(s) (columns) from the list
```{r, eval=FALSE}
mtcars[1:2]
#                      mpg cyl
# Mazda RX4           21.0   6
# Mazda RX4 Wag       21.0   6
# Datsun 710          22.8   4
# Hornet 4 Drive      21.4   6
# Hornet Sportabout   18.7   8
# Valiant             18.1   6
# ...

mtcars[c("mpg", "cyl")]
#                      mpg cyl
# Mazda RX4           21.0   6
# Mazda RX4 Wag       21.0   6
# Datsun 710          22.8   4
# Hornet 4 Drive      21.4   6
# Hornet Sportabout   18.7   8
# Valiant             18.1   6
# ...
```
---
      - Two indices extract a two-dimensional object
```{r}
mtcars[1:2, 1:4]
#               mpg cyl disp  hp
# Mazda RX4      21   6  160 110
# Mazda RX4 Wag  21   6  160 110
```
      
- When subset with `[ ]`, a tibble returns a tibble
---
### 4.2.5 Preserving dimensionality

- An important difference between data frames and tibbles is that if a data frame is subset with a single dimension index, it will return a single dimension object.

- Why does this cause problems?  Because the class of object returned depends on the subset indices, so code can break in unexpected ways.

- Two ways to prevent this:

      1. Include the `drop = FALSE` argument when subsetting a data frame
      
      2. Use tibbles, which do not drop dimensions
---
## 4.3 Selecting a single element

- For lists, `[ ]` subsets a list, while `[[ ]]` and `$` extract an element

### 4.3.1 `[[ ]]`

- A single value must be used inside `[[ ]]`

      - `x[[c(1, 2)]]` is equivalent to `x[[1]][[2]]`
      
      - Hadley recommends using `[[ ]]` to select single values from atomic vectors as well, because it is more explicit about what is returned
---
### 4.3.2 `$`

- `x$y` is equivalent to `x[["y"]]`

- The `$` operator does not work with stored variables

```{r, eval=FALSE}
var <- "cyl"
mtcars$var
# NULL

mtcars[[var]]
# [1] 6 6 4 6 8 6 8 4 4 6 6 8 8 8 8 8 8 4 4 4 4 8 8 8 8 4 4 4 8 6 8 4
```

- Unlike `[[ ]]`, `$` allows partial matching, which can be dangerous

      - Ways to avoid this
      
            - Set `options(warnPartialMatchDollar = TRUE)`
            
            - Use tibbles, which do not allow partial matching
---
### 4.3.3 Missing and out-of-bound indices
- Due to the inconsistency of how R handles such indices, `purrr::pluck()` and `purrr::chuck()` are recommended
```{r, eval=FALSE}
x <- list(
  a = list(1, 2, 3),
  b = list(3, 4, 5)
)

purrr::pluck(x, "a", 1)
# [1] 1

purrr::pluck(x, "c", 1)
# NULL

purrr::pluck(x, "c", 1, .default = NA)
# [1] NA
```


---
### 4.3.4 `@` and `slot()`

- `@` is `$` for S4 objects (to be revisited in Chapter 15)

- `slot()` is `[[ ]]` for S4 objects
---
## 4.4 Subsetting and assignment

- Subsetting can be combined with assignment to edit values
```{r, eval=FALSE}
x <- c("Mary", "James", "Perdo")
x[3] <- "Pedro"
x
# [1] "Mary"  "James" "Pedro"
```

- The lengths of the subset and assignment vector should be the same to prevent recycling

- An element of a list can be converted to NULL this way
```{r, eval=FALSE}
x <- list(a = 1, b = 2)
x[["b"]] <- NULL
str(x)
# List of 1
#  $ a: num 1
```
---
- By subsetting with nothing, we can maintain the structure of the object
```{r, eval=FALSE}
mtcars[] <- lapply(mtcars, as.integer)
is.data.frame(mtcars)
# [1] TRUE

mtcars <- lapply(mtcars, as.integer)
is.data.frame(mtcars)
# [1] FALSE
```

---
## 4.5 Applications
### 4.5.1 Lookup tables (character subsetting)
```{r, eval=FALSE}
x <- c("m", "f", "u", "f", "f", "m", "m")
lookup <- c(m = "Male", f = "Female", u = NA)
lookup[x]
#        m        f        u        f        f        m        m 
#   "Male" "Female"       NA "Female" "Female"   "Male"   "Male"
```


---
### 4.5.2 Matching and merging by hand (integer subsetting)
- The `match()` function allows merging a vector with a table
```{r, eval=FALSE}
grades <- c("D", "A", "C", "B", "F")

info <- data.frame(
  grade = c("A", "B", "C", "D", "F"),
  desc = c("Excellent", "Very Good", "Average", "Fair", "Poor"),
  fail = c(F, F, F, F, T)
)

id <- match(grades, info$grade)
id
# [1] 3 2 2 1 3

info[id, ]
#   grade      desc  fail
# 4     D      Fair FALSE
# 1     A Excellent FALSE
# 3     C   Average FALSE
# 2     B Very Good FALSE
# 5     F      Poor  TRUE
```

---
### 4.5.3 Random samples and bootstrapping (integer subsetting)
```{r, eval=FALSE}
mtcars[sample(nrow(mtcars), 3), ] # use replace = TRUE to replace
#                     mpg cyl  disp  hp drat    wt  qsec vs am gear carb
# Lotus Europa       30.4   4  95.1 113 3.77 1.513 16.90  1  1    5    2
# Mazda RX4          21.0   6 160.0 110 3.90 2.620 16.46  0  1    4    4
# Cadillac Fleetwood 10.4   8 472.0 205 2.93 5.250 17.98  0  0    3    4
```


---
### 4.5.4 Ordering (integer subsetting)
```{r, eval=FALSE}
mtcars[order(mtcars$mpg), ]
#                      mpg cyl  disp  hp drat    wt  qsec vs am gear carb
# Cadillac Fleetwood  10.4   8 472.0 205 2.93 5.250 17.98  0  0    3    4
# Lincoln Continental 10.4   8 460.0 215 3.00 5.424 17.82  0  0    3    4
# Camaro Z28          13.3   8 350.0 245 3.73 3.840 15.41  0  0    3    4
# Duster 360          14.3   8 360.0 245 3.21 3.570 15.84  0  0    3    4
# Chrysler Imperial   14.7   8 440.0 230 3.23 5.345 17.42  0  0    3    4
# Maserati Bora       15.0   8 301.0 335 3.54 3.570 14.60  0  1    5    8
# ...
```

---
### 4.5.5 Expanding aggregated counts (integer subsetting)
- We can expand a count column by using `rep()`
```{r, eval=FALSE}
df <- tibble(x = c("Amy", "Julie", "Brian"), n = c(2, 1, 3))
df[rep(1:nrow(df), df$n), ]
# A tibble: 6 x 2
#   x         n
#   <chr> <dbl>
# 1 Amy       2
# 2 Amy       2
# 3 Julie     1
# 4 Brian     3
# 5 Brian     3
# 6 Brian     3
```


---
### 4.5.6 Removing columns from data frames (character)
- We can remove a column by subsetting, which does not change the object
```{r, eval=FALSE}
df[, 1]
# A tibble: 3 x 1
#   x    
#   <chr>
# 1 Amy  
# 2 Julie
# 3 Brian
```
- We can also delete the column using `NULL`
```{r, eval=FALSE}
df$n <- NULL
df
# A tibble: 3 x 1
#   x    
#   <chr>
# 1 Amy  
# 2 Julie
# 3 Brian
```


---
### 4.5.7 Selecting rows based on a condition (logical subsetting)

```{r, eval=FALSE}
mtcars[mtcars$gear == 5, ]
#                 mpg cyl  disp  hp drat    wt qsec vs am gear carb
# Porsche 914-2  26.0   4 120.3  91 4.43 2.140 16.7  0  1    5    2
# Lotus Europa   30.4   4  95.1 113 3.77 1.513 16.9  1  1    5    2
# Ford Pantera L 15.8   8 351.0 264 4.22 3.170 14.5  0  1    5    4
# Ferrari Dino   19.7   6 145.0 175 3.62 2.770 15.5  0  1    5    6
# Maserati Bora  15.0   8 301.0 335 3.54 3.570 14.6  0  1    5    8
```


---
### 4.5.8 Boolean algebra versus sets (logical and integer)
- `which()` gives the indices of a Boolean vector

```{r, eval=FALSE}
(x1 <- 1:10 %% 2 == 0) # 1-10 divisible by 2
#  [1] FALSE  TRUE FALSE  TRUE FALSE  TRUE FALSE  TRUE FALSE  TRUE
(x2 <- which(x1))
# [1]  2  4  6  8 10
(y1 <- 1:10 %% 5 == 0) # 1-10 divisible by 5
#  [1] FALSE FALSE FALSE FALSE  TRUE FALSE FALSE FALSE FALSE  TRUE
(y2 <- which(y1))
# [1]  5 10
x1 & y1
# [1] FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE  TRUE
```

