---
title: "AdvR Chapter 22: Debugging"
author: "Tan Ho (@_TanHo)"
date: "Last Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: libs/switch-themer.css
    chakra: libs/remark.js
    nature:
      highlightLines: true
    includes:
      after_body: libs/toggle.html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      comment = "#>")
xaringanExtra::use_tile_view()
xaringanExtra::use_panelset()
xaringanExtra::use_webcam()
xaringanExtra::use_editable()
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         
  mute_unhighlighted_code = TRUE  
)

```

## Preface

--

<center><img src = "eikifjb.png" alt= "Everything I Know Is From Jenny Bryan" width = 300> </center>

"Everything I Know Is From Jenny Bryan" 

\- Sharla Gelfand 

--

\- Tan Ho

---

## Preface

There are some excellent talks on debugging (some are better on debugging than Adv R itself!)

- Jenny Bryan's ["Object of type closure is not subsettable"](https://github.com/jennybc/debugging#readme) talk from rstudio::conf 2020
- Jenny Bryan and Jim Hester's book: ["What They Forgot to Teach You About R"](https://rstats.wtf/) Ch11
- Hadley's video on a [minimal reprex for a shiny app](https://www.youtube.com/watch?v=9w8ANOAlWy4) 
--
` <- insanely useful!`

--

and I suggest that you visit these resources later if you're looking for more of a talk or text-based content on debugging.

---

## Today's Plan

We've all "read" the chapter, so I'll (obviously) take the liberty to live-code some debugging. Along the way, we'll try to hit: 

- print/message debugging
- traceback/error (Tidyverse: `rlang::last_trace`/`rlang::last_error`)
- `browser()`: open at a specific point, or when a specific condition is met (`expr` argument)
- `debug`/`debugonce`/`undebug`: step through a function
- `options(error = utils::recover)`: open `utils::recover` to navigate the call stack to hone in on the error
- `options(warn = 2)`: throw an error instead of warning

Parallel Code:
 - `{catchr}`
 - Snaketail and stdout/stderr logging

by looking through some examples, a few contrived, a few less-contrived, and lastly (possibly?) a Shiny app?

---
## Examples

```{r eval = FALSE}

```

---

## Debug a Shiny app?

I can do this too if people are keen!

