---
title: "Advanced R"
subtitle: "Chapter 19: Quasiquotation"
author: "Jon Harmon"
output:
  xaringan::moon_reader:
    nature:
      highlightLines: true
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

```{r libraries, include = FALSE}
library(rlang)
library(knitr)
library(purrr)
library(flair)
```

# 19.1: Introduction

* **Quotation** prevents evaluation
* **Unquotation** re-allows evaluation
* **Quasiquotation** allows you to mix the two
  * Useful (?) for metaprogramming
  * Let's see if that's true!

---

# 19.2: Motivation

```{r cement, results='hide'}
cement <- function(...) {
  args <- rlang::ensyms(...)
  paste(
    purrr::map(args, rlang::as_string), 
    collapse = " "
  )
}
```
--
```{r cement-call, results='hide'}
canonical <- paste(
  letters[c(3, 15, 14, 20, 18, 9, 22, 5, 4)],
  collapse = ""
)
cement(
  This, example, is, !!canonical
)
```
--
```{r cement-print, echo=FALSE}
cement(
  This, example, is, !!canonical
)
```

---

# 19.2.1: Vocabulary

* **Evaluated** arguments obey R's normal rules.
* **Quoted** arguments are captured and processed in a special way.

--
```{r vocabulary}
# Evaluated
mean(1:5)
1:5
```
--
```{r vocabulary-2}
# Quoted
library(rlang)
```
```{r vocabulary-quoted, eval = FALSE}
rlang
#> Error: object 'rlang' not found
```

---

# 19.2.1: More Vocabulary

* **quoting function**
* "function that quotes 1 or more arguments"
* non-standard evaluation
* NSE

---

# 19.2.1: More Vocabulary

* **quoting function** `==` <br>
"function that quotes 1 or more arguments" `==` <br>
non-standard evaluation `==` <br>
NSE

--

**Kinda.** Technically NSE is about arguments only.

---

# 19.2.2: Exercises: Base

```{r motivation-examples-noflair, eval=FALSE, include=FALSE, results = 'hide'}
library(MASS)

mtcars2 <- subset(mtcars, cyl == 4)

with(mtcars2, sum(vs))
sum(mtcars2$am)
```
```{r motivation-examples-noflair-flaired, echo = FALSE}
flair::decorate("motivation-examples-noflair") %>% 
  flair::knit_print.with_flair()
```


---

# 19.2.2: Exercises: Base

```{r motivation-examples, echo = FALSE, eval = FALSE, results = 'hide'}
library(MASS)

mtcars2 <- subset(mtcars, cyl == 4)

with(mtcars2, sum(vs))
sum(mtcars2$am)
```
```{r motivation-examples-flair, echo = FALSE}
flair::decorate("motivation-examples") %>% 
  flair::flair("MASS") %>% 
  flair::flair("cyl == 4") %>% 
  flair::flair("sum(vs)") %>% 
  flair::flair("am") %>% 
  flair::knit_print.with_flair()
```

---

# 19.2.2: Exercises: Tidyverse

```{r motivation-examples-t-noflair, eval=FALSE, include = FALSE, fig.show='hide'}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)

byCyl <- mtcars %>%
  group_by(cyl) %>%
  summarise(mean = mean(mpg), .groups = "drop_last")

ggplot(byCyl, aes(cyl, mean)) + geom_point()
```
```{r motivation-examples-t-noflair-flaired, echo = FALSE}
flair::decorate("motivation-examples-t-noflair") %>% 
  flair::knit_print.with_flair()
```


---

# 19.2.2: Exercises: Tidyverse

```{r motivation-examples-t, include = FALSE, eval = FALSE, fig.show='hide'}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)

byCyl <- mtcars %>%
  group_by(cyl) %>%
  summarise(mean = mean(mpg), .groups = "drop_last")

ggplot(byCyl, aes(cyl, mean)) + geom_point()
```
```{r motivation-examples-t-flair, echo = FALSE}
flair::decorate("motivation-examples-t") %>% 
  flair::flair("dplyr") %>% 
  flair::flair("ggplot2") %>% 
  flair::flair("cyl") %>% 
  flair::flair("mean(mpg)") %>% 
  flair::flair("cyl") %>% 
  flair::flair("mean") %>% 
  flair::knit_print.with_flair()
```

Note: The first `mean` shouldn't be highlighted, but I'm struggling with {flair} `r emo::ji("frowning")`

---

# 19.3: Quoting

* "We'll need a pair of functions" (~6 pairs, including base)
  * **directly supplied** vs
  * **indirectly supplied**

---

# 19.3.1: Capturing expressions
  
* `{rlang}`
  * `expr` vs `enexpr`
  * `exprs` vs `enexprs`
  * `quote directly` vs `quote the thing in the calling environment`

---

# 19.3.1: Capturing expressions

```{r expr-vs-enexpr}
testing <- "foo"
testing2 <- testing
rlang::expr(testing)
rlang::expr(testing2)
```
--
```{r expr-vs-enexpr2}
rlang::enexpr(testing)
```

---

# 19.3.1: Capturing expressions

```{r expr-vs-enexpr3}
catch_it <- function(x) rlang::enexpr(x)
catch_it(testing)
catch_it(testing2)
```

---

# 19.3.1: Capturing expressions

* `exprs` and `enexprs` capture a list of `expr` or `enexpr`

```{r exprs-vs-enexprs}
rlang::exprs(x = testing, y = testing2, z = paste(testing, testing2))
# shorthand for
# list(
#  x = expr(testing), 
#  y = expr(testing2), 
#  z = expr(paste(testing, testing2))
# )
```

---

# 19.3.1: Capturing expressions

* The pairs are always **console_version** + ***en*console_version**
* Aka ***no-en*-function_version** + **function_version**

---

# 19.3.2: Capturing symbols

* `sym` and `syms` make sure the thing they're capturing is symbol or character
* `ensym` and `ensyms` in functions

---

# 19.3.3: With base R

* `expr` is to `quote`
* as `enexpr` is to `substitute` ("normal" usage)
* as `exprs` is to `alist`
* as `enexprs(...)` is to `as.list(substitute(...()))` ???
  * This is undocumented and kinda insane and I don't think I'll ever use it

---

# 19.3.3: With base R

* A `substitute` example: `library`

```{r library}
library_meat <- function(package) {
  as.character(substitute(package)) # Without as.character, it returns a name
}
library_meat(rlang)
```

---

# 19.3.4: Substitution

* `substitute` can be used to... substitute

```{r substitution}
f4 <- function(x) substitute(x * 2)
f4(a + b + c)
f4(whatever)
```

---

# 19.3.4: Substitution

* `substitute` can be used to... substitute
* It's confusing, so make it explicit.
* I tried to implement this and found what I really wanted was to unquote part of the expression...

```{r substitution2}
f4b <- function(x) substitute(x, list(a = 1, b = 2, c = 3))
f4b(a + b + c) # Not what I meant!
f4b(literally - anything)
```

---

# 19.4: Unquoting

* quasiquotation means you can unquote selectively
* Base R doesn't allow you to selectively unquote (mostly)
* Unquoting means "replace the object with the thing it represents"

---

# 19.4.1: Unquoting one argument

* The confusingly beautiful heart of quasiquotation: **!!**
* Use it to unquote one thing

```{r bang-bang}
x <- expr(-1)
y <- "a character"
rlang::expr(f(!!x, y))
rlang::expr(f(x, !!y))
rlang::expr(mean(1:3) + mean(4:6))
rlang::expr(!!mean(1:3) + !!mean(4:6))
```

---

# 19.4.2: Unquoting a function

* Now we start to move toward metaprogramming

```{r replace-function}
replace_f <- function(func) {
  f <- rlang::enexpr(func)
  rlang::expr((!!f)(x, y)) # Would be better if we used the actual formals...
}
replace_f(mean)
replace_f(rlang::enexpr)
```

---

# 19.4.5: Unquoting many arguments

* `!!!` is to `!!` as `exprs` is to `expr`
* **"unquote-splice"** because it **unquotes** the list then **splices** them in as if they were all separate arguments

```{r bang-bang-bang}
multi_arg <- list(a = 1, b = 2, c = "other")
rlang::expr(f(!!!multi_arg, another_arg))
```

---

# 19.4.6: The polite fiction of !!

* `!!` and `!!!` don't actually exist

```{r fictional-functions}
rlang::`!!`
rlang::`!!!`
```

---

# 19.6 ... (dot-dot-dot)

* This section is the leftovers, it isn't about `...` per se
* `:=` "colon-equals" or "digested is" (how I think of it)
  * Another fiction to trick R into working how we want
  * LHS of `=` can't be evaluated, so we trick R

```{r digested}
var <- "my_var_name"
val <- 1:3
tibble::tibble(!!var := val)
```

---

# 19.6.2: exec()

* This feels out-of-place in this chapter
* `rlang::exec` is similar to `base::do.call`
* Allows you to use unquoting to do fancy things

```{r exec}
params <- list(na.rm = TRUE, trim = 0.1)
func <- "mean"
rlang::exec(func, x = 1:10, !!!params)
```

---

# 19.7: Case studies

* I'd like a function like `stringr::str_replace(string, pattern, replacement)`
* But instead, `funky::fn_replace(function, pattern, replacement)`
* I don't want the user to have to use any `rlang` or `quote` or `substitute` (etc)