---
title: "<u>Advanced R</u><br>Chapter 11: Function Operators"
subtitle: "Stephen Holsenbeck"
author: "sholsen at alumni dot emory dot edu"
date: "last updated: `r Sys.Date()`"
institution: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    chakra: libs/remark.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.align="center", fig.width=5, fig.height=5, warning = FALSE, message = TRUE)
library(magrittr)
```

```{r xaringan-themer, include = FALSE}
xaringanthemer::style_duo_accent(
  primary_color = "#bada90",
  secondary_color = "#bada55",
  header_font_google = xaringanthemer::google_font("Roboto", "400"),
  header_background_content_padding_top = "3.5rem",
  text_font_google   = xaringanthemer::google_font("Lato", "300"),
  header_color = "skyblue",
  title_slide_text_color = "#c9d6de",
  padding = "12px 32px 12px 32px",
  code_font_google = xaringanthemer::google_font("Oxygen Mono", "200"),
  code_highlight_color = "rgba(232,239,250, 0.3)",
  code_font_size = "0.75rem",
  code_inline_font_size = "0.9rem",
  title_slide_background_image = "images/switchboard-operator.jpg"
  
)
xaringanthemer::style_extra_css(
  list(
    pre = list( 
    display = "block",
    `font-family` = "monospace",
    `white-space` = "pre-wrap",
    margin = "0px"
    ),
    `h1,h2,h3,h4,h5` = list(
      `padding-top` = ".25rem",
      `padding-bottom` = ".25rem",
      `margin-top` = ".25rem",
      `margin-bottom` = ".25rem"
    ),
    `.title-slide > h1, h2, h3` = list(
      `background-color` = "rgba(255,255,240, .5);",
      `text-shadow` = "2px 2px rgba(0,0,0,.4);"
      )
    )
, append = TRUE)
```



```{css, echo = F}
hide {
  display: none;
}
.remark-slide-content h1 {
  font-size: 45px;
}
h1 {
  font-size: 2em;
  margin-block-start: 0.67em;
  margin-block-end: 0.67em;
}
.remark-slide-content {
  font-size: 16px
}
.remark-code {
  font-size: 14px;
}
code.r {
  font-size: 14px;
}
pre {
  margin-top: 0px;
  margin-bottom: 0px;
}
.red {
  color: #FF0000;
}
.footnote {
  color: #800020;
  font-size: 9px;
}
```


## 11.1 Overview 


- > A function operator (FO) is a function that takes one (or more) functions as input and returns a function as output. 


- > FOs allow you to factor out complexity in order to make your code more readable and reusable.


```{r, 'fo table', eval=T, echo=F, include=T}
kableExtra::kable_minimal(knitr::kable(tibble::tibble(
    Term = c('Functionals', 'Function Factory', 'Function Operator'),
    `Required Input` = c('Function', '', 'Function'),
    `Optional Input` = c('Vector', 'Vector, Function', 'Vector'),
    Output = c('Vector', 'Function', 'Function')
  ))) 
```

--

Three types of FOs:

+ __Behavioral FO__: Change the behavior of a function, e.g. "memorizing" previous outputs

  + `memoise::memoise()`

+ __Output FO__: Wrap a function and change the output

  + `purrr::possibly()`, `purrr::safely()`, `purrr::quietly()`

+ __Input FO__: Pre-fill arguments of a function

  + `purrr::partial()`

---

### A general example


```{r 'general example'}
slowly <- function(f, n){
  force(f)
  force(n)
  function(...){
    stopifnot(is.numeric(n))
    message(
      glue::glue('Sleeping for {n} seconds.')
    )
    Sys.sleep(n)
    f(...)
  }
}

purrr::walk(
  c('hello', 'world'),
  slowly(cat, 0.1), #<<
  sep = '\n' # Passed to `f()` via `...`
)
```

.footnote[
https://gist.github.com/ColinFay/d32cf4c9c5fb8d849f12a4e98d6c0549
]

---

## 11.2 Existing FOs



Useful when `map`ping over unpredictable inputs or trying various attempts:

 - files
 - manually input data
 - survey responses
 - webpages
 - statistical modeling across imbalanced classes
 
---

### `purrr::safely` example

```{r 'safely ex', error = TRUE}
state_data <- tibble::rownames_to_column(as.data.frame(state.x77), "State")
get_state_stats <- rlang::as_function(~{
  .l <- state_data$State %in% .x
  if (purrr::some(.l, isTRUE)) {
    state_data[.l,]
  } else 
    stop("No match", call. = FALSE)
  })

purrr::map(c(
  "Georgia",
  "Pennsylvania",
  "Texas",
  "Puerto Rico"
), get_state_stats)
```

---
#### `purrr::safely` example cont.


```{r 'safely ex 2'}
safe_state_stats <- purrr::safely(get_state_stats)
state_stats <- purrr::map(c(
  "Georgia",
  "Pennsylvania",
  "Texas",
  "Puerto Rico"
), safe_state_stats)
str(state_stats, max.level = 2)
```

---
#### `purrr::safely` example cont.

```{r 'safely ex 3'}
state_stats <- purrr::transpose(state_stats)
str(state_stats, max.level = 2)
```


---

#### `purrr::safely` example cont.

```{r 'safely ex 4'}
knitr::kable(dplyr::bind_rows(state_stats$result))
```



--- 

### `purrr::possibly`
 
Similar to `purrr::safely` except:

- requires `otherwise` which specifies the outcome when an error occurs 
- only returns objects / alternatives and not errors.

---
### `purrr::quietly`
 
Similar to `purrr::safely` except:

- returns a list with 
  - result
  - output
  - messages
  - warnings


---

### Caching Computations with `memoise::memoise()`

Testing the speed of `memoise::memoise()`

```{r 'memoise 1'}
slow_function <- function(x) {
  Sys.sleep(0.2)
  x * runif(1)
}
fast_function <- memoise::memoise(slow_function)
```

<br />

```{r 'memoise 2'}
system.time(slow_function(1))
system.time(slow_function(1)) #<<
```

```{r 'memoise 3'}
system.time(fast_function(11))
system.time(fast_function(11)) #<<
```

---

### `memoise::memoise` continued

Even if you've changed the inputs since the most recent call, it will still be fast.

```{r 'memoise 4'}
system.time(fast_function(22))
system.time(fast_function(33))
system.time(fast_function(22)) #<<
```


In fact, it remembers everything from the same session (assuming you haven't used `memoise::forget()`).

```{r 'memoise 5'}
system.time(fast_function(11))
system.time(fast_function(22))
system.time(fast_function(33))
```

---


### `purrr::partial`

Using `purrr::partial()` to set `na.rm = TRUE`

```{r 'partial ex'}
purrr::map(list(
  rnorm(3, 1.5)
  , c(rnorm(3, 1.5), NA)
), mean)
# With partial
pmean <- purrr::partial(mean, na.rm = TRUE)
purrr::map(list(
  rnorm(3, 1.5)
  , c(rnorm(3, 1.5), NA)
), pmean)
```

---

### 11.3.1 #5
> Modify delay_by() so that instead of delaying by a fixed amount of time, it ensures that a certain amount of time has elapsed since the function was last called. That is, if you called `g <- delay_by(1, f); g(); Sys.sleep(2); g()` there shouldn’t be an extra delay.


```{r 'Exercise 11.3.1'}

delay_by <- function(f, delay, .t = Sys.time(), ...) {
  force(f)
  force(delay)
  amount <- .t + delay - Sys.time()
  timestamp()
  function (...) {
      Sys.sleep(amount)
    timestamp()
      f(...)
    }
}



delay_by(runif, as.difftime(4, units = "secs"))(10)
```

---
# Using `purrr::safely` demo

---

# Thanks to Cohort 1's Tony ElHabr
