---
title: "Advanced R"
subtitle: "Chapters 15 & 16, S4 and Trade-offs"
author: "Erick Knackstedt"
institute: "@Eknackstedt"
date: " `r Sys.Date()` "
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css","custom.css"]
    includes:
      after_body: insert-logo.html
---


```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

# Outline

--

+ Introducing S4
--

+ Classes, generics, & methods
--

+ Prototypes, constructors, helpers, & validators
--

+ Creating & accessing generics and methods
--

+ Method dispatch, multiple inheritance, & multiple dispatch
--

+ Interaction of S4 & S3
--

+ Trade-offs

---

## Introducing S4

--

> Why S4?

--

+ Like S3, uses `function overloading` to throttle the number of functions required to interact with objects
--

+ A more formal and rigorous update to S3
--

+ Much of the code is written by `John Chambers` himself
--

+ Used extensively in the `Bioconductor` project
--

+ Fairly independent of S3, separate systems
--

+ Provides multiple inheritance and multiple dispatch
--

+ Has the slot and a specialized subsetting operator `@`
--

+ Implemented using the `methods` package

---

## Classes, generics, & methods

--

> How do you create a class?

--

`setClass("myClass", slots= ...., contains =...., prototype =....)`

--

> How do you create a generic for the class?

--

`setGeneric(name, def)`
--

> How do you create a method for the generic?

--

`setMethod(f, signature, definition)`

---

## Prototypes, constructors, helpers, & validators

--
> If the prototype is optional, why should we use it?

--

This allows you to pass in default values for class slots (attributes)
--

> I want to build on top of another class. Is there a workaround to copy and pasting attributes from one class to another?

--

Why, yes there is! The `contains` argument in `setClass` allows you to specify the class to inherit from
--

> If I can use `new()` to construct an object, why do I bother to create a constructor?

--

Apparently programmers want other people to use their stuff. Constructors politely define the slots (attributes) that require data, which you then assign to `new()` in the body of the constructor

---

### Prototypes, constructors, helpers, & validators cont.

--

> I don't want this use of my class to break, how do I validate that an object has the correct attributes?

--

`setValidity(Class, method)`, where the method is a function that contains the validity checks. You define this once and it is associated with object construction going forward. It is not triggered if an object is modified

--

> Well, if I can't validate on modificaiton, what do I do?

--

This is where we build on the idea of helpers and wrap an accessor in a function that also validates the inputs to underlying attributes

---

## Creating & accessing generics and methods

--

> What was the point of a generic again?

--

Generics are functions that describe behavior found across multiple classes of objects. By decomposing functions into generics and methods we solve the issue of function overpopulation

--

> And the S4 syntax for creating one?

--

```{r, eval=FALSE}
setGeneric(name, def)

# Example
setGeneric("myGeneric", function(x) standardGeneric("myGeneric"))
```

---

### Creating & accessing generics and methods cont.

--

> What was the point of a method again?

--

A method is a function for a specific class of object. Which means as a programmer you are still writing up functions for every class but for the user the interface is standardized through the generic

--

> And the S4 syntax for creating one?

--

```{r, eval=FALSE}
setMethod(f, signature, definition)

#Example
setMethod("myGeneric", "Person", function(x) {
  # method implementation
})
```

---

## Method dispatch, multiple inheritance, & multiple dispatch

--

> I've overloaded generic functions with a bunch of class specific methods. Now what?

--

The generic function looks at the class and lookups up the method!

--

> What if I built up classes one on top of the other? You know, like evolution or something?

-- 

Well, the generic will search for the methods assigned to the classes and use the one that is "closest". If there there is not a closest, then it picks the method that comes earliest in the alphabet. 

---

## Interaction of S4 & S3

--

> How exactly do these two system allow interactions?

--

Well, you can assign an S3 object to a S4 `slot` or build on top of an S3 object by using `contains`.

--

> Sounds easy! Really glad there isn't a fancy function required to allow this to work!!

--

Silly, of course there is a fancy function required! Seems painless though

`setOldClass(Classes)`

```{r, eval=FALSE}
setClass("factor",
  contains = "integer",
  slots = c(
    levels = "character"
  ),
  prototype = structure(
    integer(),
    levels = character()
  )
)
setOldClass("factor", S4Class = "factor")
```

### Interaction of S4 & S3 cont.

--

> Ok, we've got a way to inherit attributes from S3 Objects. What about importing S3 generics to S4?

--

`setGeneric(name, def)`

Does this break its use for S3 methods?

---

## Trade-offs

> S4 vs S3

--

+ S3 and S4 solve similar problems with solutions pointed towards the entire R ecosystem
+ S4 has conventions by design, while S3 allows the same conventions, others, or none
+ S3 seems more like the self-service friendly tool-set
+ S3 is more generally used, S4 is used by large projects like `bioconductor`

--

> R6 vs S3

+ R6 seems to allow by nature the build up of a mini-ecosystem of functionality
+ R6 reference semantics seem cool
+ Something about using the `$` to pipe method calls
+ R6 seems to have some nice features to prevent breaking changes when improvements to classes or methods are made