# Environments

```{r, include=FALSE}
library(tidyverse)
library(kableExtra)
library(lobstr)
library(rlang)
library(openintro)
```


## 7.2.3 Parents {-}

:::question
What's the relationship between base, empty enviornment, and global enviornment?
:::

The global environment is your working enviornment, where all packages loaded are its parent. When you list the parents of the gloal environment the last environment is always `base` which is responsible for loading all other packages. The base's parent is the "ultimate ancestor", the empty enviornment.


:::question
Clarification: the global environment's parent is all the packages loaded within the environment, not the empty environment, right? How do we check this? `parent.env(global)`?
:::

Each package becomes a parent to the global environment. We can inspect local enviornments using `search_envs()`

```{r}
search_envs()
```

## 7.2.5 Getting and setting {-}

:::question
"But you can’t use `[[` with numeric indices, and you can’t use `[`:"

It makes sense to me that you can't use numeric indexes because objects in an environment aren't ordered, but why can't you use `[`?
:::

:::TODO
The solutions manual states: "The second option would return two objects at the same time. What data structure would they be contained inside? But can we elaborate on what this means?
:::

## 7.2.6 Advanced bindings {-}

:::question
Let's go over some source code that uses `env_bind_lazy()` - Hadley mentions its used when autoloading datasets with packages - can we find an example of this? How is this different from including `LazyData: true` in your description file?
:::

:::TODO
XXX
:::

## 7.3.1.2 {-}

:::question
I understood the recursion in the prior example, but what is inherits doing here? Can we go through this line for line and discuss what is happening in this function?

```{r}
fget <- function(name, env = caller_env(), inherits = TRUE) {
  # Base case
  if (env_has(env, name)) {
    obj <- env_get(env, name)

    if (is.function(obj)) {
      return(obj)
    }
  }

  if (identical(env, emptyenv()) || !inherits) {
    stop("Could not find function called \"", name, "\"", call. = FALSE)
  }

  # Recursive Case
  fget(name, env_parent(env))
}
```

```{r}
# Test
mean <- 10
fget("mean", inherits = TRUE)
```
:::

## 7.4.2 Function enviornment {-}

:::question
I found this section a little confusing. Can we go over the second figure in the section? Where does `x` live?  `g` points to x but `x` is in the global enviornment? Can we come up with our own example for a function being bound to the global environment but accessing variables from its own enviornment? (I think this is what the second figure in the section is trying to display)

```{r}
y <- 1
e <- env()
e$g <- function() 1
e$g
```
:::


:::TODO

Is this example trying to display that if we change the function to return `y` this does not return `1` because `y` inside the function cannot access `y` at the global scope?

```{r}
y <- 1
e <- env()
e$g <- function() y
e$g
```
:::

## 7.4.3 Namespaces {-}

:::question
"Every binding in the package environment is also found in the namespace environment; this ensures every function can use every other function in the package. But some bindings only occur in the namespace environment. These are known as internal or non-exported objects, which make it possible to hide internal implementation details from the user."

Can we find some simple-ish package source code as an example of this? 
:::

:::TODO
XXX
:::


## 7.6 Data Structures {-}

:::question
Can we build on this function to include `on.exit()`?

```{r}
my_env <- new.env(parent = emptyenv())
my_env$a <- 1

get_a <- function() {
  my_env$a
}
set_a <- function(value) {
  old <- my_env$a
  my_env$a <- value
  invisible(old)
}
```
:::

:::question
The bullet point mentions hashmaps but I'm still really unclear on what that is. What are hash tables and are they related?
:::

:::TODO
:::

