# Conditions

```{r, include=FALSE}
library(tidyverse)
library(rlang)
```

## 8.3 Ignoring conditions {-}

:::question
I can't quite think of a time where `try` is more appropraite than `tryCatch` - does anyone have an example?
:::

It seems that try is just a wrapper for `tryCatch` and you can use it whenever you'd use `try`

:::question
When would you actually use `suppressWarning`? Maybe when loading libraries? I created an example for `suppressMessages`, but does someone have a better, practical use case?
:::

As a function user it's common to supress warnings when loading in data and using `readr` and `dplyr` - but not so much as a function creator

## 8.4.2 Exiting handlers {-}

:::question
```{r, eval=FALSE}
tryCatch(
  message = function(cnd) "There",
  {
    message("Here")
    stop("This code is never run!")
  }
)
#> [1] "There"
```

> The protected code is evaluated in the environment of `tryCatch()`, but the handler code is not, because the handlers are functions.

Clarifcation question what is the "handler code" and "protected code"?
:::

The protected code is inside the `{}` and the handler code is `message = function(cnd) "There"`

:::question
In the tryCatch example where we have `finally` print "Thank God for Beer" I find it interesting that this is printed **before** the code inside the tryCatch. Can anyone explain why?
:::

The handler code is evaulated prior to the protected code, that is why we see it second!

## 8.4.3 Calling handlers {-}

:::question
Can we make up our own bubbling and muffled `withCallingHandlers` examples? The example code was a little unclear to me.
:::

:::TODO
XXX
:::

:::question
Why is this message executed once per message in the function?

```{r, eval=FALSE}
withCallingHandlers(
  message = function(cnd) cat("Caught a message!\n"), 
  {
    message("Someone there?")
    message("Why, yes!")
  }
)
```

```
#> Caught a message!
#> Someone there?
#> Caught a message!
#> Why, yes!
```
:::

:::TODO
XXX
:::

:::question
> The return value of a calling handler is ignored because the code continues to execute after the handler completes; where would the return value go? That means that calling handlers are only useful for their side-effects.

Can we come up with an example for this masking? I think seeing it will help me understand...
:::

:::TODO
XXX
:::

:::question
How does muffling differ from `suppressWarnings`?
:::

:::TODO
XXX
:::

## 8.4.5.2 Exercises {-}

:::question
In the example 

```{r, eval=FALSE}
show_condition <- function(code) {
  tryCatch(
    error = function(cnd) "error",
    warning = function(cnd) "warning",
    message = function(cnd) "message",
    {
      code
      NULL
    }
  )
}
```

The first three calls to `show_condition` make sense to me, and I even understand that the first time the code is evaluated inside a `tryCatch` it exists (that's why it returns `message`) but how are you suppling `3` arguments to a function that just takes on one argument, `code`? Is that what the `{}` are for?

```{r, eval=FALSE}
show_condition({
  10
  message("?")
  warning("?!")
})
```
:::

`{}` let us execute multiple lines of code!

## 8.4.5.3 Exercises {-}

:::question
I couldn't follow the manual's answer for what's happening here, can we come up with our own answer for what's happening here in words?
:::

```{r, eval=FALSE}
withCallingHandlers(
  message = function(cnd) message("b"),
  withCallingHandlers(
    message = function(cnd) message("a"),
    message("c")
  )
)
```

:::TODO
XXX
:::

## 8.6.3 Resignal {-}

:::question
Can we go over what is happening here?

```{r, eval=FALSE}
warning2error <- function(expr) {
  withCallingHandlers(
    warning = function(cnd) abort(conditionMessage(cnd)),
    expr
  )
}

warning2error({
  x <- 2 ^ 4
  warn("Hello")
})
```

```
Error: Hello
```

:::TODO
:::

## 8.6.4 Record {-}

:::question
Why are we using `cnd_muffle` here?

```{r, eval=FALSE}
catch_cnds <- function(expr) {
  conds <- list()
  add_cond <- function(cnd) {
    conds <<- append(conds, list(cnd))
    cnd_muffle(cnd)
  }
  
  withCallingHandlers(
    message = add_cond,
    warning = add_cond,
    expr
  )
  
  conds
}

catch_cnds({
  inform("a")
  warn("b")
  inform("c")
})
```
:::

If we remove `cnd_muffle` we see that `a`, `b`, and `c` are printed to the console prior to getting the `conds` output. 

:::question
Would it be possible in the second Record example to create a function that doesn't require us to put the `abort` statement at the end, just ignoring it? Or because `abort` is an exiting handler it needs to be last?
:::

```{r, eval=FALSE}
catch_cnds <- function(expr) {
  conds <- list()
  add_cond <- function(cnd) {
    conds <<- append(conds, list(cnd))
    cnd_muffle(cnd)
  }
  
  tryCatch(
    error = function(cnd) {
      conds <<- append(conds, list(cnd))
    },
    withCallingHandlers(
      message = add_cond,
      warning = add_cond,
      expr
    )
  )
  
  conds
}

catch_cnds({
  abort("a")
  inform("b")
  warn("c")
})
```

:::TODO
XXX
:::

:::question
What is `signal` and what is it doing here?

```{r, eval=FALSE}
log <- function(message, level = c("info", "error", "fatal")) {
  # if we remove match.arg and just use level
  # the signal returns infoerrorfatal as a single string
  # we need match.arg to find the selected level
  # and it defaults to the first if none are selected
  level <- match.arg(level)
  signal(message, "log", level = level)
}
```
:::

Signal is a general function that calls abort, inform  or warn. It has the same signature as those functions except for the class argument that is necessary for signal but NULL by default for each of the others

```{r, eval=FALSE}
signal <- function(message, class, ..., .subclass) {
  if (!missing(.subclass)) {
    deprecate_subclass(.subclass)
  }
  message <- collapse_cnd_message(message)
  cnd <- cnd(class, ..., message = message)
  cnd_signal(cnd)
}
```

:::question
> If you create a condition object by hand, and signal it with signalCondition(), cnd_muffle() will not work. Instead you need to call it with a muffle restart defined, like this: `withRestarts(signalCondition(cond), muffle = function() NULL)`

Where does this code go given the prior example? 
:::

:::TODO
XXX
:::

## 8.6.6.2 Exercises {-}

:::question
> Calling handlers are called in the context of the call that signalled the condition. Exiting handlers are called in the context of the call to tryCatch().

What exactly does this mean? tryCatch evaluates what we were calling the protected code first and calling handlers execute the handling code first? Can we make a simple example?
:::

:::TODO
XXX
:::

## 8.6.6.4 Exercises {-}

:::question
> There’s no way to break out of the function because we’re capturing the interrupt that you’d usually use!

What does this mean? You can't stop the function if you set an `interrupt` argument inside a `tryCatch`?
:::

That's exactly it - this was a warning not to use `inturrupt` in your functions!

## Slides {-}

:::question
Let's revist my example from the talk: can we build on this to use tidyeval so that the user can write `beer_states %>% beer_mean(state)` [I think we need to use `.data` and note how state is given as an object so we need to use tidyeval to suppress it's evaluation...]
:::

```{r, eval=FALSE}
beer_states <- readr::read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')

beer_mean <- function(.data, x) {
  
  column_name <- deparse(substitute(x))

  msg <- glue::glue("Can't calculate mean, {column_name} is not numeric")
  
  if (!is.numeric(.data[[column_name]])) {
    abort(
      message = msg,
      arg = column_name,
      data = .data
    )
  } else {
    mean(which(!is.na(.data[[column_name]])[.data[[column_name]]]))
  }
}

beer_states %>% beer_mean(barrels)
```

```
[1] 806.4551
```
