# Functionals

```{r, include=FALSE}
library(purrr)
```

## 9.2.2 Anonymous functions and shortcuts {-}

:::question
We saw you'll get an error if you try to map over elements that don't exist, and can use `.default` to override that. Is this related to `tryCatch` in some way? Can we look at the `map` source code for `.default?` And how would we overcome this error if we were to use base R's `lapply(x, 'two')`

```{r}
x <- list(
  list(one = "a"),
  list(one = "b"),
  list(one = "c")
)

map_chr(x, 'two', .default = NA)
```

```{r, eval=FALSE}
lapply(x, 'two')
```

```
Error in get(as.character(FUN), mode = "function", envir = envir) : object 'two' of mode 'function' was not found
```
:::

:::TODO
Within `purrr` [there's the function](https://github.com/tidyverse/purrr/blob/7104367bb6599f13e56c554bd07488b508a8a02b/R/as_mapper.R#L98)

```{r, eval=FALSE}
find_extract_default <- function(.null, .default) {
  if (!missing(.null)) {
    .null
  } else {
    .default
  }
}
```

So it doesn't seem to be a conditional, but rather a way to deal with missing errors
:::

:::TODO
XXX
:::

## 9.2.6.4 Exercise {-}

:::question
In order to extract p-values the solution manual suggests using `map_dbl` but could we use `pluck` to get these values? If so how?
:::

```{r}
trials <- map(1:100, ~ t.test(rpois(10, 10), rpois(10, 7)))
tibble(p_value = map_dbl(trials, "p.value"))
```

:::TODO
```{r, eval=FALSE}
pluck(trials, "p.value")
```
:::

## 9.2.6.5 Exercise {-}

:::question
Can we make this work?

```{r, eval=FALSE}
x <- list(
  list(1, c(3, 9)),
  list(c(3, 6), 7, c(4, 7, 6))
)

triple <- function(x) x * 3
map(x, map, .f = triple)
```
:::

:::TODO
XXX
:::

## 9.4.1 Same type of output as input: `modify()` {-}

When using modify we now get the warning:`Warning message: `modify()` is deprecated as of rlang 0.4.0. Vector tools are now out of scope for rlang to make it a more focused package.`

How should we rewrite this example?

```{r, eval=FALSE}
data.frame(
  x = 1:3,
  y = 6:4
) %>% 
  modify( ~ .x * 2)
```
:::

:::TODO
XXX
:::

## 9.4.5 Any number of inputs: pmap() and friends {-}

:::question
I want to use pmap to map over a vector, but the metadata for the functions other arguments is elsewhere - how would I combine these using `pmap`? The book says to name the metadata columns the same name as your function which I did but this still doesn't work?
:::

```{r, eval=FALSE}
my_data <- 1:3

metadata <- tribble(
  ~id, ~mult, ~adder,
  "one",   2,      5,
  "two",   3,      6,
  "three", 4,      7
)

the_function <- function(vec, id, mult, adder) {
  glue::glue("{id} is now {vec * mult + adder}")
}

# my_data doesn't change but we want to map over the metadata
# x = string
# y = multiplier
# z = adder
pmap(list(metadata), ~the_function(vec = my_data))
```

:::TODO
XXX
:::

## 9.4.6.2 Exercise {-}

:::question
I see how we can use `iwalk` and `walk2` for writing to multiple files, but the question asks about disadvantages to this - what would they be?
:::

:::TODO
```{r, eval=FALSE}
cyls <- split(mtcars, mtcars$cyl)
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
walk2(cyls, paths, write.csv)

mtcars %>% 
  split(mtcars$cyl) %>% 
  set_names(~ file.path(temp, paste0("cyl-", .x, ".csv"))) %>% 
  iwalk(~ write.csv(.x, .y))
```
:::


## 9.5.4 Multiple inputs {-}

:::question
Can we think of a simple example for `reduce2()`? I was playing with `accumulate(c(FALSE, FALSE, TRUE), `|`)`, maybe we can combine this with something?
:::

:::TODO
XXX
:::

## 9.5.5 Map-reduce {-}

:::question
I like the books Hadoop simple example - can we come up with one with code or is it solely for computing on multiple machines?
:::

:::TODO
XXX
:::

