# R6

## 14.2 Classes and methods {-}

:::question
> Any ... R6 method called for its side effects ... should return invisible(self).

WHY? What is `invisible(self)` doing in the add function below? I tried removing this line but it doesn't change the output...
:::

```{r, eval=FALSE}
Accumulator <- R6Class("Accumulator", list(
  sum = 0,
  add = function(x = 1) {
    self$sum <- self$sum + x 
    invisible(self)
  })
)

x <- Accumulator$new()
```

:::TODO
:::

## 14.2.2 Important methods {-}

:::question
Why don't we need to specifically call `haldey2$print` to see the output below)? Is the print method just a built in output of an `R6` object?
:::
 
```{r, eval=FALSE}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) {
    self$name <- name
    self$age <- age
  },
  print = function(...) {
    cat("Person: \n")
    cat("  Name: ", self$name, "\n", sep = "")
    cat("  Age:  ", self$age, "\n", sep = "")
    #invisible(self)
  }
))

hadley2 <- Person$new("Hadley")
hadley2
```

```
Person: 
  Name: Hadley
  Age:  NA
```

:::TODO
:::

## Exercises 14.2.6.4 {-} 

:::question
Why can't we use method chaining to access the current time zone?
:::

```{r, eval=FALSE}
Timezone <- R6Class(
  classname = "Timezone", 
  public = list(
    get = function() {
      Sys.timezone()
    },
    set = function(value) {
      stopifnot(value %in% OlsonNames())
      
      old <- self$get()
      Sys.setenv(TZ = value)
      invisible(old)
    })
)

tz <- Timezone$new()

old <- tz$set("Antarctica/South_Pole")
tz$get()
```

```
[1] "America/Los_Angeles"
```

```{r, eval=FALSE}
tz$set("Antarctica/South_Pole")$get()
```

```
Error in tz$set("Antarctica/South_Pole")$get : 
  $ operator is invalid for atomic vectors
```

:::TODO
```{r, eval=FALSE}
Timezone <- R6Class(
  classname = "Timezone", 
  public = list(
    current_zone = Sys.timezone(),
    get = function() {
      Sys.timezone()
      self$current_zone <- Sys.timezone()
    },
    set = function(value) {
      stopifnot(value %in% OlsonNames())
      
      old <- self$get()
      Sys.setenv(TZ = value)
      self$current_zone <- Sys.setenv(TZ = value)
      invisible(old)
    })
)

tz <- Timezone$new()

tz$set("Antarctica/South_Pole")$current_zone
```
:::

## 14.3.2 Active fields {-}

:::question
Why can we write over the `$name` but not `$age` methods when both are part of the `active` list?
:::

```{r, eval=FALSE}
Person <- R6Class("Person", 
  private = list(
    .age = NA,
    .name = NULL
  ),
  active = list(
    age = function(value) {
      if (missing(value)) {
        private$.age
      } else {
        stop("`$age` is read only", call. = FALSE)
      }
    },
    name = function(value) {
      if (missing(value)) {
        private$.name
      } else {
        stopifnot(is.character(value), length(value) == 1)
        private$.name <- value
        self
      }
    }
  ),
  public = list(
    initialize = function(name, age = NA) {
      private$.name <- name
      private$.age <- age
    }
  )
)
```

```{r, eval=FALSE}
hadley4 <- Person$new("Hadley", age = 38)
hadley4$name <- "Maya"
hadley4$name
```

```
[1] "Maya"
```

```{r, eval=FALSE}
hadley4$age <- 20
```

```
Error: `$age` is read only
```

:::TODO
:::

## 14.4 Reference semantics {-}

:::question
> $clone() does not recursively clone nested R6 objects. If you want that, you’ll need to use $clone(deep = TRUE).

Can we see this in action using a subclass?
:::

:::TODO
```{r, eval=FALSE}
A <- R6Class(
  classname = "A",
   public = list(
    field = "foo",
    method = function() {
      "bar"
    }
  )
)

B <- A$new()
C <- B$clone(deep = FALSE)
```
:::


:::question
When or why would I set the clone argument to FALSE?
:::

Clone has a large memory footprint so if you're going to create a lot of R6 methods you may want to exclude this!

:::question
Is R6 "public" related to private/public methods/classes in traditional OO?
:::

