# Quasiquotation

```{r, include=FALSE}
library(rlang)
library(purrr)
```

## 19.2 Motivation {-}

:::question
How does `cement` compare to `glue`? Can we look at the source code for quasiquotation and using `{}` for interpolation?
:::

:::TODO
They are opposite; while `glue` evaluates expressions inside a string, `cement` quotes expressions.  
:::

## 19.3.4 Substitution {-}

:::question

> Is it supplied by the developer of the code or the user of the code? In other words, is it fixed (supplied in the body of the function) or varying (supplied via an argument)?

What are two examples (from the chapter or our own) of the former vs latter
:::

This destinction can be put as: do you want exactly what the developer/coder typed or do you want what was put into the function? This will tell you if you want to use `expr` or `enexpr`

:::question
Hadley creates a table for quoting expressions using base R and tidy eval, but there isn't really a one to one relationship here. For instance `substitute` is compared to `enexprs` but that fails here?

```{r, eval=FALSE}
f4 <- function(x) substitute(x * 2)
f4(a + b + c)

f5 <- function(x) enexprs(x*2)
f5(a + b + c)
```
:::

:::TODO
That is because they are not a 1-1 relationship, `substitute` has `5` different functions under the hood (but they are not `S3` methods!!), these are only related somewhat. 
:::

## Exercises 19.3.6.5 {-}

:::question
What does the argument `.named` do in `exprs`? What does "to ensure all dots are named" mean?
:::

You can pass extra arguments to your function as dots, and this will ensure that those extra arguments are named


## 19.4.6 Polite fiction of !! {-}

:::question
What exactly is happening here? Why is `y` evaluating to `6`?!

```{r}
df <- data.frame(x = 1:5)
y <- 100
with(df, x + !!y)
```
:::

:::TODO
Any number that isn't `0` is `TRUE` in R:

```{r}
as.logical(1.2123)
```

So because `y==TRUE`, `with` is taking every element in `x` and adding `1` to it because once you introduce a `+` inside `with` you're essentially saying `as.numeric(TRUE)` which is `1`.

```{r}
with(df, x + TRUE) == with(df, x + 1)
```
:::

## 19.4.7 Nonstandard ASTs {-}

:::question
When talking about non-standard ASTs Hadley says:

> These are valid, and occasionally useful, but their correct use is beyond the scope of this book.

What is their correct use?
:::

A use for this could potentially be if you need to manipulate an AST and subsitute an environment within the AST? In the question below we are creating a non-standard AST (not that we should be!)

## 19.4.8.2 Exercises {-}

:::question
What exactly is going on here? What is `<inline integer>`?

```{r}
lobstr::ast(mean(1:10))
lobstr::ast(mean(!!(1:10)))
```
:::

The inline integer is the actual integers being inserted into the AST rather than as expressions

## Non-quoting {-}

:::question
What exactly is the difference between "turn quoting off" and "using unquoting" -- maybe to explain this we can we come up with an example for "turning quoting off" when we expect "unquoting" and therefore the operation fails?
:::

In base R functions it's all or nothing: you can either quote everything or nothing, you cannot selectively quote expressions.


## 19.6 (...) {-}

:::question
Based on [this thread]() it looks like `...` should be the first line of defense before using tidyeval - why?

```{r, eval=FALSE}
my_groups <- function(df, ...) {
  df %>%
    group_by(...) %>%
    summarise(n = n())
}

my_groups <- function(df, col) {
  df %>%
    group_by(!!enquo(col)) %>%
    summarise(n = n())
}
```
:::

The first option lets the robust function `group_by` figure out any quoting. Furthermore, the second option only allows the user to supply a single column to group by and assumes the user is passing in a valid column.

:::question
Can we go over the "spatting" example, I have no idea what this means.
:::

I think "splatting" is the correct term*. It means taking a single argument in a list and interpreting it as separate arguments. Hadley is referring to the specific use of `!!!` to unpack a list into function arguments.

* at least googling for "splatting Ruby" gives lots of relevant hits vs "spatting Ruby"

## 19.6.2 `exec()` {-}

:::question
Can we use this function instead of `do.call` in our Expressions shiny app? Do we need to change `argumentlist()` to use `list2` rather than `list`?

```{r, eval=FALSE}
do.call(call2, argumentlist(), quote = TRUE)
```
:::

```{r, eval=FALSE}
exec(call2, !!!argumentlist())
```

## 19.6.3 `dots_list()` {-}

:::question
This function seems pretty rad, what's a use case for it in the wild?
:::

We can use this function anytime there's a use for `list2`!

## 19.6.4 With base R {-}

:::question
What is going on with this RCurl example and how does it avoid using `do.call`? Can we find some Hadley source code that uses it since he said he liked it at one point? Why do you think he moved away from it? 

```{r, eval=FALSE}
f <- function(..., .dots) {
  dots <- c(list(...), .dots)
  # Do something
}
```
:::

:::TODO
The `dots` object combines all the arguments passed as dots and the arguments not passed as dots. This gives a lot (perhaps too much) leeway to the user of the function.
:::

