# Evaluation

```{r, include=FALSE}
library(rlang)
library(purrr)
library(dplyr)
library(palmerpenguins)
```

## 20.1 Introduction {-}

:::question
There's an immediate distinction between unquotation (user) and evaluation (developer). What are they?
:::

Unquoting seems to be "evaluate this one part and then return the expression"
evaluate means "give me the result of this whole expression"

## 20.2.1 `local` {-}

:::question
This is such a cool function - where is it used in the wild? Maybe we can come up with a TidyModels example or something where you'd want to throw away your intermediates? 
:::

:::TODO
:::

:::question
Can we go over again what is happening in `local2`? Why isn't the environment included as a function argument? Furthermore I'm a little confused by what the function `caller_env()` does.

```{r, eval=FALSE}
local2 <- function(expr) {
  env <- env(caller_env())
  eval(enexpr(expr), env)
}

foo <- local2({
  x <- 10
  y <- 200
  x + y
})
```
:::

:::TODO
:::

## 20.2.2 `source` {-}

:::question
Expression vectors get a shout out again (they are an aside in the Expression chapter) and while Hadley advises against introducing another data structure, I wanted to make sure my high level understanding was correct: eval can be vectorized if the input is of type `expression vector` [but not if it's of type `list`]?
:::


```{r, eval=FALSE}
exp_vec <- expression(x <- 4, x)
eval(exp_vec)
```

```{r, eval=FALSE}
exp_list <- list2(
  rlang::expr(x <- 4),
  rlang::expr(x)
)
eval(exp_list)
```

## 20.2.3 `function` {-}

:::question
I think I understand this issue based on the concrete example but can we summarize the "gotcha" here? Is it that the developer could assume that using `eval` with a quasiquoted function will result in returning the function, but really it evaluates because `srcref` is sneaky?
:::

It's basically that the printing of the function is a lie (because base R doesn't get it). And that's dangerous and confusing, so use rlang to make it not a lie.

## Exercises 20.2.4.4 {-}

:::question
Can we go over what exactly is happening here in the solutions manual prior to applying a `map`? I don't really understand what this is returning...

```{r, eval=FALSE}
source2 <- function(path, env = caller_env()) {
  file <- paste(readLines(path, warn = FALSE), collapse = "\n")
  exprs <- parse_exprs(file)

  res <- vector(mode = "list", length = length(exprs))      
  for (i in seq_along(exprs)) {
    res[[i]] <- eval(exprs[[i]], env)
  }

  invisible(res)
}

tmp_file <- tempfile()
writeLines(
  "x <- 1
   x
   y <- 2
   y  # some comment",
  tmp_file
)

(source2(tmp_file))
```
:::

:::TODO
:::

## Excercises 20.2.4.5 {-}

:::question
Can we also go over what is happening in this question?

```{r, eval=FALSE}
local3 <- function(expr, envir = new.env()) {
  call <- substitute(eval(quote(expr), envir))
  eval(call, envir = parent.frame())
}
```
:::

It creates a new environment with the calling environment as its parent and then evaluates the expression in that environment.

## 20.3.3 Dots {-}

:::question
I have no idea what's happening in this section. Can we go over the point/what this code is showing us? 

```{r, eval=FALSE}
f <- function(...) {
  x <- 1
  g(..., f = x)
}

g <- function(...) {
  enquos(...)
}

x <- 0
qs <- f(global = x)
qs
```
:::

:::TODO
As a preliminary, I note that the following code also illustrates the point:

```{r, eval=FALSE}
f <- function(...) {
  x <- 1
  g(..., f = x)
}
g <- function(...) {
  enquos(...)
}
x <- 0
qs <- f(x)
qs
```

I think what he's trying to say is that if you didn't have quosures you would have to create a list of environments to match the list of expressions if you wanted your expressions to contain the same names but not necessarily have those names have the same meaning.

But here we have a magical situation where each quosure has the same expression but each environment is different and thus each `x` will be evaluated differently when the time comes.

I think the real barrier to truly getting it is figuring out a plausible example of wanting the same name to have different meanings in your list of expressions.. and I'm at a loss for that at the moment.
:::

## 20.3.4 Under the hood {-}

:::question
> Unfortunately, however, there is no clean way to make ~ a quasiquoting function.

Can we come up with a (broken) example of trying to quasiquote an object of type `formula`? I think seeing this fail will help me to understand its limitations
:::

:::TODO
:::

## Exercises 20.3.6.2 

:::question
What is going on here?? 

```{r eval=FALSE}
enenv <- function(x){
  get_env(enquo(x))
}

# Test
capture_env <- function(x){
  enenv(x)
}

enenv(x)
#> <environment: R_GlobalEnv>
capture_env(x)  # functions execution environment is captured
#> <environment: 0x60a2538>
```

:::

The `enquo` is capturing the environment of the `x` being passed into enenv, so it returns a different environment (the one inside `capture_env`) than running `enenv(x)` on its own

## 20.4.1 Basics {-}

:::question
I am clearly missing something here. What is so special about multiplying across a vector? Is it that data masks allow us to just write `y` instead of `df$y`?
:::

I think the simplicity of the multiplication may be what he's going for: the reader can focus on the language feature provided without the distraction of a complex computation.

## 20.4.3 `subset` {-}

:::question
What is the `eval_tidy` doing here?

```{r, eval=FALSE}
subset2 <- function(data, rows) {
  rows <- enquo(rows)
  rows_val <- eval_tidy(rows, data)
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}
```
:::

:::TODO
:::

## 20.4.5 `transform` {-}

:::question
Can we say in words what this one is doing or comment all the lines? 
:::

:::TODO
```{r, eval=FALSE}
transform2 <- function(.data, ...) {
  
  # create quosures for all the supplied arguments and values
  dots <- enquos(...)

  # for all argument value pairs
  for (i in seq_along(dots)) {
    # get the names of the arguments
    name <- names(dots)[[i]]
    # get their values
    dot <- dots[[i]]

    # add column with names equal to argument names
    # eval_tidy and values (single column) equal to the evaluated quosure (dot)"
    .data[[name]] <- eval_tidy(dot, .data)
  }

  .data
}
```
:::

## 20.5.2 Handling ambiguity {-}

:::question
> If you unquote, val will be early evaluated by enquo(); if you use a pronoun, val will be lazily evaluated by eval_tidy().

What is the case where this subtile distinction between `.env$val` vs `!!val` matters? 
:::

:::TODO
:::

## 20.6.1 substitute() {-}

:::question
Why exactly can't we use `subset` with `map`?
:::

:::TODO
:::

:::question
> Calling subset() from another function requires some care: you have to use substitute() to capture a call to subset() complete expression

Why?!
:::

No reason to expect this to work:

```{r, eval=FALSE}
test <- function(df, expr) {
  subset(df, expr)
}

test(mtcars, cyl > 4)
```

```
#> Error in eval(e, x, parent.frame()): object 'cyl' not found
```

substitute only looks at the expression from which it was called. this doesn't work because subset sees `substitute(expr)`, not `cyl > 4`

```{r, eval=FALSE}
test <- function(df, expr) {
  subset(df, substitute(expr))
}
test(mtcars, cyl > 4)
```

```
#> Error in subset.data.frame(df, substitute(expr)): 'subset' must be logical
```

so instead, we are going to build the call to `subset`, substituting in what was passed as arguments, and then eval the call.

```{r, eval=FALSE}
test <- function(df, expr) {
  subset_call <- substitute(subset(df, expr))
  print(subset_call)
  eval(subset_call)
}
test(mtcars, cyl > 4)
```

```
#> subset(mtcars, cyl > 4)
```


## 20.6.2 `match.call` {-}

:::question
Can we speak more about the differences in these two functions and their pros and cons?
:::

```{r, eval=FALSE}
resample_lm2 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_env <- env(env, resample_data = resample_data)
  lm_call <- expr(lm(!!formula, data = resample_data))
  expr_print(lm_call)
  eval(lm_call, lm_env)
}
resample_lm2(y ~ x, data = df)
```


```{r, eval=FALSE}
resample_lm0 <- function(
  formula, data,
  resample_data = data[sample(nrow(data), replace = TRUE), ,
                       drop = FALSE],
  env = current_env()
) {
  formula <- enexpr(formula)

  lm_call <- expr(lm(!!formula, data = resample_data))
  expr_print(lm_call)
  eval(lm_call, env)
}

df <- data.frame(x = 1:10, y = 5 + 3 * (1:10) + round(rnorm(10), 2))
(resamp_lm1 <- resample_lm0(y ~ x, data = df))
```

I think these two functions should work identically assuming formula and data are the only two inputs that are set by the user. (The results will differ unless a seed is set.)

I suppose the `resample_lm0()` here provides more flexibility to the user (compared to `resample_lm2()`, i.e. the user can specify `resample_data` (since it's an argument) instead of being "forced" to use the custom function `resample()`, which is embedded in the body of `resample_lm2()`. 

Downside to `resample_lm0` is that there might be too much flexibility. Yes, `resample_data` is customizable, but so is `env`. What if the user changes `env = current_env()` to `env = caller_env()`? Then the function breaks.

:::question
Is the closest equivalent to `deparse(substitute(x))` with rlang `expr_text(enexpr(x))` (assuming this is in a function, hence the `en` in `enexpr`)?
:::

```{r}
penguins <- palmerpenguins::penguins
add_bleh <- function(df, nm = deparse(substitute(df))) {
  col_out <- sym(sprintf('new_%s_col', nm))
  df %>% 
    mutate(!!col_out := 'bleh')
}
add_bleh(penguins) %>% slice(1) %>% glimpse()
```

:::TODO
:::

:::question
In [this thread](https://community.rstudio.com/t/quasiquotation-inside-a-formula/14929/11) Lionel says "`enexpr` should almost never be used. So when should it?
:::

Seems like when wrapping a base NSE function like `lm` or trying to `deparse(substitute(x))` using `rlang`

