---
title: "R4DS Questions From Advanced R"
output:
  html_document:
    toc: true
    toc_float: true
    css: styles.css
---

A companion to [Advanced R](https://adv-r.hadley.nz/rcpp.html#acknowledgments) and suppliment to [Advanced R Solutions](https://advanced-r-solutions.rbind.io/names-and-values.html#copy-on-modify)

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("r4ds.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; 
                        right:0;
                        padding:0;
                        width: 25%;
               height 25%')
```

# Chapter 2 

## 2.3 Copy-on-modify 

:::question
copy-on-modify vs copy-in-place: is one more preferable in certain situations?
:::

modify in place only happens when objects with a single binding get a special performance optimisation and to enviornments.

## 2.3.2 Function calls

:::question
Can we go over and break down figure in 2.3.2
:::

When you create this function:

```{r, eval=FALSE}
crazyfunction <- function(eh) {eh}
```

`eh` doesn't exist in memory at this point.

```{r, eval=FALSE}
x <- c(1,2,3)
```

x exists in memory.

```{r, eval=FALSE}
z <- crazyfunction(x) 
```

`z` now points at `x`, and `eh` still doesn't exist (except metaphorically in Canada). `eh` was created and exists WHILE `crazyfunction()` was being run, but doesn't get saved to the global environment, so after the function is run you can't see its memory reference. 

The round brackets `(eh)` list the arguments, the curly brackets `{eh}` define the operation that it's doing - and you're assigning it to `crazyfunction`. 

**R functions automatically return the result of the last expression** so when you call that object (the argument `eh`) it returns the value of that argument. This is called **implicit returns**

## 2.3.3 Lists

:::question
When would you prefer a deep copy of a list to a shallow copy? Is this something to consider when writing functions or package development or is this more something that's optimized behind the scenes?
:::

Automagical!

## 2.3.5 Character vectors


:::question
Is there a way to clear the "global string pool"?
:::

## 2.4.1 Object size


:::question
I understand that `ALTREP` makes the size of `obj_size(1:3) == obj_size(1:100)` but do I need to have a better grasp than that? [Should we go over this](https://blog.revolutionanalytics.com/2017/09/altrep-preview.html)
:::

It's kind of cool to know but... Probably not.

## 2.5.1 Modify-in-place

:::question
"When it comes to bindings, R can currently only count 0, 1, or many. That means that if an object has two bindings, and one goes away, the reference count does not go back to 1: one less than many is still many. In turn, this means that R will make copies when it sometimes doesn’t need to."

Can we come up with an example of this? It seems really theoretical right now.
:::

XXX

:::question

Can we break down this code a bit more? I'd like to really understand when and how it's copying three times.

```{r}
# dataframe of 5 columns of numbers
x <- data.frame(matrix(runif(5 * 1e4), ncol = 5))
# median number for each column
medians <- vapply(x, median, numeric(1))

# subtract the median of each column from each value in the column
for (i in seq_along(medians)) {
  x[[i]] <- x[[i]] - medians[[i]]
}
```
:::

If we run the following chunk, `dubugging` our loop

```{r, eval=FALSE}
cat(tracemem(x), "\n")
```

```
<0x7fdc99a6f9a8> 
```

```{r, eval=FALSE}
for (i in 1:5) {
  x[[i]] <- x[[i]] - medians[[i]]
}
```

```
tracemem[0x7fdc99a6f9a8 -> 0x7fdc9de83e38]: 
tracemem[0x7fdc9de83e38 -> 0x7fdc9de83ea8]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de83ea8 -> 0x7fdc9de83f18]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de83f18 -> 0x7fdc9de83f88]: 
tracemem[0x7fdc9de83f88 -> 0x7fdc9de83ff8]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de83ff8 -> 0x7fdc9de84068]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de84068 -> 0x7fdc9de840d8]: 
tracemem[0x7fdc9de840d8 -> 0x7fdc9de84148]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de84148 -> 0x7fdc9de841b8]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de841b8 -> 0x7fdc9de84228]: 
tracemem[0x7fdc9de84228 -> 0x7fdc9de84298]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de84298 -> 0x7fdc9de84308]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de84308 -> 0x7fdc9de84378]: 
tracemem[0x7fdc9de84378 -> 0x7fdc9de843e8]: [[<-.data.frame [[<- 
tracemem[0x7fdc9de843e8 -> 0x7fdc9de84458]: [[<-.data.frame [[<- 
```

When we run `tracemem` on the for loop above we see each column is copied twice followed by the `[[<-.data.frame [[<- `, the stack trace showing exactly where the duplication occured.

When we write `x[[i]] <- value`, we're really calling the function `[[<-.data.frame` with inputs `x`, `i`, and `value`. We can step into the call of this function by running `debug(``[[<-.data.frame``)`:

```{r, eval=FALSE}
debug(`[[<-.data.frame`)
```

and once inside, use `tracemem()` to find where the new values are assigned to the column:

```{r eval=FALSE}
function (x, i, j, value) 
{
  if (!all(names(sys.call()) %in% c("", "value"))) 
    warning("named arguments are discouraged")
  cl <- oldClass(x)
  # this is where another copy of x is made!
  class(x) <- NULL
```

```
 # tracemem[0x7fdc9d852a18 -> 0x7fdc9c99cc08]: 
```

```{r, eval = FALSE}
nrows <- .row_names_info(x, 2L)
  if (is.atomic(value) && !is.null(names(value))) 
    names(value) <- NULL
  if (nargs() < 4L) {
    nc <- length(x)
    if (!is.null(value)) {
      N <- NROW(value)
      if (N > nrows) 
        stop(sprintf(ngettext(N, "replacement has %d row, data has %d", 
          "replacement has %d rows, data has %d"), N, 
          nrows), domain = NA)
      if (N < nrows) 
        if (N > 0L && (nrows%%N == 0L) && length(dim(value)) <= 
          1L) 
          value <- rep(value, length.out = nrows)
        else stop(sprintf(ngettext(N, "replacement has %d row, data has %d", 
          "replacement has %d rows, data has %d"), N, 
          nrows), domain = NA)
    }
    x[[i]] <- value
    if (length(x) > nc) {
      nc <- length(x)
      if (names(x)[nc] == "") 
        names(x)[nc] <- paste0("V", nc)
      names(x) <- make.unique(names(x))
    }
    class(x) <- cl
    return(x)
  }
  if (missing(i) || missing(j)) 
    stop("only valid calls are x[[j]] <- value or x[[i,j]] <- value")
  rows <- attr(x, "row.names")
  nvars <- length(x)
  if (n <- is.character(i)) {
    ii <- match(i, rows)
    n <- sum(new.rows <- is.na(ii))
    if (n > 0L) {
      ii[new.rows] <- seq.int(from = nrows + 1L, length.out = n)
      new.rows <- i[new.rows]
    }
    i <- ii
  }
  if (all(i >= 0L) && (nn <- max(i)) > nrows) {
    if (n == 0L) {
      nrr <- (nrows + 1L):nn
      if (inherits(value, "data.frame") && (dim(value)[1L]) >= 
        length(nrr)) {
        new.rows <- attr(value, "row.names")[seq_len(nrr)]
        repl <- duplicated(new.rows) | match(new.rows, 
          rows, 0L)
        if (any(repl)) 
          new.rows[repl] <- nrr[repl]
      }
      else new.rows <- nrr
    }
    x <- xpdrows.data.frame(x, rows, new.rows)
    rows <- attr(x, "row.names")
    nrows <- length(rows)
  }
  iseq <- seq_len(nrows)[i]
  if (anyNA(iseq)) 
    stop("non-existent rows not allowed")
  if (is.character(j)) {
    if ("" %in% j) 
      stop("column name \"\" cannot match any column")
    jseq <- match(j, names(x))
    if (anyNA(jseq)) 
      stop(gettextf("replacing element in non-existent column: %s", 
        j[is.na(jseq)]), domain = NA)
  }
  else if (is.logical(j) || min(j) < 0L) 
    jseq <- seq_along(x)[j]
  else {
    jseq <- j
    if (max(jseq) > nvars) 
      stop(gettextf("replacing element in non-existent column: %s", 
        jseq[jseq > nvars]), domain = NA)
  }
  if (length(iseq) > 1L || length(jseq) > 1L) 
    stop("only a single element should be replaced")
  x[[jseq]][[iseq]] <- value
  # here is where x is copied again!
  class(x) <- cl
```

```
# tracemem[0x7fdc992ae9d8 -> 0x7fdc9be55258]: 
```
```{r, eval=FALSE} 
  x
}
```

Thus seeing exactly where the three copies are happening!
