---
title: "Advanced R"
subtitle: "Chapter 25: Rewriting R code in C++"
author: "Daryn Ramsden"
output:
  xaringan::moon_reader:
    nature:
      highlightLines: true
    lib_dir: libs
    css: xaringan-themer.css
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#3092FF",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono"),
)
```

```{r libraries, include = FALSE}
library(rlang)
library(knitr)
library(purrr)
library(flair)
library(Rcpp)
```

### Sad reality: sometimes R won't cut it

#### When can C++ help the R programmer?


  * Loops that can’t be easily vectorised because subsequent iterations depend on previous ones.

  * Recursive functions, or problems which involve calling functions millions of times. The overhead of calling a function in C++ is much lower than in R.

  * Problems that require advanced data structures and algorithms that R doesn’t provide. 
  


  
---
### Rcpp: The best way to use C++ 



```{r}
library(Rcpp)
```

Key functions in `Rcpp`:
  * `cppFunction`
  
  * `sourceCpp`
  
  
---
### Detour: If you really want an `Rcpp` tutorial


Dirk Eddelbuettel's `Rcpp` tutorial at useR! 2020 (h/t: Pavitra)

<iframe width="560" height="315" src="https://www.youtube.com/embed/57H34Njrns4" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>


  
---
### Using `cppFunction`

Just put the whole C++ function in a string and pass it to `cppFunction`

```{r}
cppFunction('int add(int x, int y, int z) {
  int sum = x + y + z;
  return sum;
}')

add(1729, 99, 14)
```


---
### An example R function vs C++ implementation 

```{r}
fibonacci_r <- function(n){
  
  if(n < 2) return(n)
  
  return (fibonacci_r(n-1) + fibonacci_r(n-2))
  
}
```

```{r}
cppFunction("int fibonacci_cpp(int n){
  
  if (n < 2) return n;
  
  return fibonacci_cpp(n-1) + fibonacci_cpp(n-2);
  
}")
```

```{r}
bench::mark(fibonacci_r(10), fibonacci_cpp(10))
```

---
### Using `sourceCPP`

`sourceCpp` reads a C++ file and exports functions for use in R 

To use `sourceCpp`, start your standalone C++ file with:

```{Rcpp eval = FALSE}
#include <Rcpp.h>
using namespace Rcpp;
```

And before each function you want to export:
```{Rcpp eval = FALSE}
// [[Rcpp::export]]
```

---
### Really useful: the vector classes

`Rcpp` provides C++ vector types corresponding to the main vector types in R 

  * `IntegerVector` 
  * `NumericVector` 
  * `LogicalVector` 
  * `CharacterVector`



---
### Example using `NumericVector`

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double meanC(NumericVector x) {
  int n = x.size();
  double total = 0;

  for(int i = 0; i < n; ++i) {
    total += x[i];
  }
  return total / n;
}

/*** R
x <- runif(1e5)
bench::mark(
  mean(x),
  meanC(x)
)
*/
```

---
### Benchmarking against `mean`


```{r}
x <- runif(1e5)
bench::mark(
  mean(x),
  meanC(x)
)
```

---
### Other types

`Rcpp` also has types for the following R entities
  
  * `List`  
  
  * `DataFrame`
  
  * `Function`
  
  
  
---
### Attributes


The attributes of an R object can be queried using the `.attr()` method of the corresponding Rcpp object.


An example of using `.attr()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector attribs() {
  NumericVector out = NumericVector::create(1, 2, 3);

  out.names() = CharacterVector::create("a", "b", "c");
  out.attr("my-attr") = "my-value";
  out.attr("class") = "my-class";

  return out;
}
```





---
### Missing values: scalars

`NA` corresponds to a different C++ constant for each underlying scalar type:

```{r}
cppFunction("
List scalar_missings() {
  int int_s = NA_INTEGER;
  String chr_s = NA_STRING;
  bool lgl_s = NA_LOGICAL;
  double num_s = NA_REAL;

  return List::create(int_s, chr_s, lgl_s, num_s);
}
")
```

```{r}
str(scalar_missings())
```


Looks mostly good. But there are some pesky details in section 25.4 (I'm going to pretend they don't exist.)


---
### Missing values: vectors

With vectors, you need to use a missing value specific to the type of vector, `NA_REAL`, `NA_INTEGER`, `NA_LOGICAL`, `NA_STRING`:


```{r}
cppFunction("List missing_sampler() {
  return List::create(
    NumericVector::create(NA_REAL),
    IntegerVector::create(NA_INTEGER),
    LogicalVector::create(NA_LOGICAL),
    CharacterVector::create(NA_STRING)
  );
}")
```


```{r}
str(missing_sampler())
```


---
### Standard Template Library 

The STL:

  * a really extensive C++ software library
  
  * has 4 components:
    * algorithms
    
    * containers
    
    * functions 
    
    * iterators
    
Note: this is not exactly the same thing as the **C++ Standard Library**


---
### Detour: if you really want to learn the STL 

You can learn the STL from STL (video 1 of n):

<iframe src="https://channel9.msdn.com/Series/C9-Lectures-Stephan-T-Lavavej-Standard-Template-Library-STL-/C9-Lectures-Introduction-to-STL-with-Stephan-T-Lavavej/player" width="720" height="405" allowFullScreen frameBorder="0" title="C9 Lectures: Stephan T. Lavavej - Standard Template Library (STL), 1 of n - Microsoft Channel 9 Video"></iframe>

You will truly get the STL if you get to the end of this series.

---
### Iterators 

C++ has many iterator types. 

Key features of iterators are:

  * Advance with `++`.
  
  * Get the value they refer to, or dereference, with `*`.

  * Compare with `==` 
  
  


---
### Example using iterator features


```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double sum3(NumericVector x) {
  double total = 0;
  
  NumericVector::iterator it;
  for(it = x.begin(); it != x.end(); ++it) {
    total += *it;
  }
  return total;
}
```

```{r}
sum3(c(1,12,201,2001))
```

---
### Algorithms

The STL also has a lot of efficiently implemented algorithms. 

The following code uses the `std::upper_bound` algorithm from the STL to create a function  that takes two arguments a vector of values and a vector of breaks, and locates the bin that each x falls into.

```{Rcpp}
#include <algorithm>
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
IntegerVector findInterval2(NumericVector x, NumericVector breaks) {
  IntegerVector out(x.size());

  NumericVector::iterator it, pos;
  IntegerVector::iterator out_it;

  for(it = x.begin(), out_it = out.begin(); it != x.end(); 
      ++it, ++out_it) {
    pos = std::upper_bound(breaks.begin(), breaks.end(), *it);
    *out_it = std::distance(breaks.begin(), pos);
  }

  return out;
}
```


---
### Particularly useful STL data structures


  * `vector`: similar to an R vector. But more efficient. Templated.
  
  * `set`: maintain a set of unique values. Good when you need to identify if you have seen a particular value already
    * `std::set`
    
    * `std::unordered_set`
  
  * `map`: data structure containing (key,value) pairs aka dictionaries. 

  


---
### Using `Rcpp` in a package 

Two simple steps:

  * In DESCRIPTION add

```{r eval = FALSE}
LinkingTo: Rcpp
Imports: Rcpp
```


  * Make sure your NAMESPACE includes:

```{r eval = FALSE}
useDynLib(mypackage)
importFrom(Rcpp, sourceCpp)
```

The only thing left is the actual code. 


---
### The end 

I have tried to read this book on multiple occasions. And failed. 


So I am glad that I saw this tweet:

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">neRds! i joined a bookclub going through Hadley Wickham&#39;s Advanced R book every thursday at 6pm. im presenting chapter 2 this week. if u wanna join us DM me</p>&mdash; Asmae Toumi (@asmae_toumi) <a href="https://twitter.com/asmae_toumi/status/1247203272496209923?ref_src=twsrc%5Etfw">April 6, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

